### 2021-12-10 reports/01-introduction.Rmd

---
output:
  pdf_document:
    includes:
      in_header: "~/headRs/StatHead.sty"
fontsize: 11pt
bibliography: "../../causal.bib"
# csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
---

# Introduction

The United Auto Workers-General Motors (UAW-GM) Cohort Study is a longitudinal occupational cohort study established in the early 1980s to study the health effects of metalworking fluids [@Eisen_1992;@Eisen_2001]. Metalworking fluids (MWF) are complex mixtures of fluids used in industrial metalworking operations to lubricate and cool machinery and parts. The three major classes of MWF are straight, soluble, and synthetic metalworking fluids [@Byers_2006]. Possible routes of human exposure include absorption through skin, inhalation or aerosols, and ingestion of droplets.

A central concern in the analysis of occupational cohorts is the potential for the healthy worker survivor effect (HWSE), the phenomenon by which healthy individuals remain at work, while less healthy individuals leave work -- possibly in response to exposure-related health decline. In the presence of the HWSE, those with the highest cumulative occupational exposures are also those who are less at risk of disease. Thus, standard measures of association would show an inverse relationship between occupational exposure and poor health outcomes [@Arrighi_1994]. The HWSE is an example of time-varying confounding affected by past exposure. Previous studies have attempted to assess the presence of the HWSE in observed data by assessing so-called path-specific associations using Cox proportional hazards modeling [@Naimi_2013;@Garcia_2017]. However, these measures of associations are themselves subject to the confounding they seek to quantify.

If sequential ignorability of exposure status at each point in follow-up and positivity can be attained conditional on covariates, then causal methods can be applied to account for the HWSE. Past studies have applied causal methods capable of accounting for time-varying confounding affected by past exposure to the study of MWF exposures and cancer mortality outcomes in the UAW-GM Cohort Study [@Garcia_2018;@Izano_2019], but the study of cancer incidence outcomes is further problematized by incomplete observation of cancer incidence outcomes over the study period. We wish to make inferences about the carcinogenicity of MWF exposure over an individual's lifetime starting three years after they enter the workforce. The UAW-GM Cohort includes those hired roughly between 1938 and 1982. However, cancer incidence reporting at the Michigan Cancer Registry did not begin until 1985. In particular, we have the presence of _left filtering_ in the UAW-GM Cohort Study when cancer incidence is the outcome of interest: before 1985, both cancer incidence status and time of cancer incidence are unknown. Observation of the complete cancer incidence outcome vector over the study period is conditional on an individual surviving to 1985 cancer-free.

In the presence of the HWSE, left filtering implies outcome misclassification that is informative of true cancer status. As part of her dissertation research, @Izano_2017_thesis conducted a quantitative bias analysis for the estimation of survival curves in the presence of left filtering and the HWSE. She simulated data compatible with the HWSE and estimated cancer-free survival curves using an inverse probability of treatment and censoring weighted Kaplan-Meier (WKM) estimator and the WKM with an Aalen filter for left-filtering (AWKM) [@Andersen_1993;@Xie_2005]. Data were simulated under five different scenarios. The causal estimands of interest were the survival curves under interventions: (1) always exposed at work with no censoring due to death and (2) never exposed at work with no censoring due to death. For each intervention, three survival curves were produced: (1) the true survival curve, (2) the WKM survival curve, and (3) AWKM survival curve. The bias in mean survival difference calculated by AWKM was negligible when the true 20-year risk of cancer was under 0.25.

Here, we replicate the simulation and bias analysis presented in Chapter 3 of @Izano_2017_thesis, embed the problem in the non-parametric structural causal approach of @Pearl_1995, comment on the results, and apply the WKM and AWKM estimators to the UAW-GM Cohort.