### 2021-12-11 paper.tex
Interim results; running on GM data now

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Estimating survival in left filtered data},
  pdfauthor={Kevin Chen},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=2.54cm]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\PassOptionsToPackage{no-math}{fontspec}

%---------------------------------------------------------------------------
% Kevin's style file
%---------------------------------------------------------------------------
\usepackage{ifxetex,ifluatex}

%---------------------------------------------------------------------------
% General
%---------------------------------------------------------------------------

% When testing
% \usepackage{lipsum}

% Line numbering
\usepackage{lineno}

% Usage
	% \linenumbers
	% \modulolinenumbers[2]
	% \renewcommand\thelinenumber{\color{red}\Roman{linenumber}}
	% \begin{linenumbers}  \end{linenumbers}

\usepackage{setspace}
%\doublespacing

\usepackage{outlines}
% \usepackage{enumerate}
\usepackage{enumitem} % more powerful than enumerate
% \setenumerate[4]{label=\alph*).}

\setlength\parindent{0pt} % Removes all indentation from paragraphs

\usepackage{sectsty} % Allows customizing section commands
% \renewcommand{\thesection}{\Alph{section}} % Lettered, not numbered sections
% \allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps
%\numberwithin{equation}{section} % Number equations within sections
%\numberwithin{figure}{section} % Number figures within sections
%\numberwithin{table}{section} % Number tables within sections

\usepackage{multicol}
\setlength{\columnsep}{15pt}
\newlength\Colsep

%---------------------------------------------------------------------------
% Tables, figures, images
%---------------------------------------------------------------------------
\usepackage{color}
% \usepackage[usenames,dvipsnames]{xcolor}
\usepackage{xparse}
\usepackage{xhfill}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0
% Requires pdfTeX
\usepackage{linegoal}
\fi

\usepackage{longtable}
\setlength{\LTcapwidth}{\textwidth}
\usepackage{xtab, booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable} % Suzanne's

\usepackage{float} % Check code chunks section for floatrow

\usepackage{graphicx}
% \usepackage{grffile}
% \usepackage{subcaption}
% For tables and figures in parts %%%%
\usepackage{caption}
% \DeclareCaptionLabelFormat{cont}{#1~#2\alph{ContinuedFloat}}
% \captionsetup[ContinuedFloat]{labelformat=cont}
%\graphicspath{ {file.path.here} }

% Set column widths in tabular environment!
\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

% STATA output for `estout` package
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}

%---------------------------------------------------------------------------
% Header/footer
%---------------------------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancyplain}
\fancyhead[L]{}
\fancyhead[C]{}
\fancyhead[R]{}
% \fancyfoot[L]{}
% \fancyfoot[C]{}
% \fancyfoot[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}
% \renewcommand{\footrulewidth}{0pt}
\setlength{\headheight}{14.5pt}

%---------------------------------------------------------------------------
% Supplement (compatible for Beamer and Article)
%---------------------------------------------------------------------------
\input{\string~/HeadRs/common_supplement.tex}

%---------------------------------------------------------------------------
% Bibliography
%---------------------------------------------------------------------------
% \usepackage[backend=bibtex]{biblatex}

%---------------------------------------------------------------------------
% Usage
%---------------------------------------------------------------------------

% For use in LaTex or Sweave:
% \input{/Users/kevinchen/Documents/computing/HeadRs/StatHead.tex}

% An example YAML header for Rmd
% ---
% title: "Stat 2"
% author: "Kevin Chen"
% date: \today
% output:
%   pdf_document:
%      # keep_tex: true
%     #  pandoc_args: [
%     #   "-V", "classoption=twocolumn"
%     # ]
%      includes:
%         in_header: /Users/kevinchen/Documents/computing/HeadRs/StatHead.tex
% bibliography: /Users/kevinchen/Documents/2017 Fa/245/MV_FinalProj/mv_final.bib
% csl: /Users/kevinchen/Documents/computing/HeadRs/AMA.csl
% geometry: margin=2cm, bottom=2.5cm
% ---

%---------------------------------------------------------------------------
% Fonts
%---------------------------------------------------------------------------

%\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
%\usepackage[utf8]{inputenc}
% Latin Modern Roman
% \usepackage{lmodern}
%\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default

\ifluatex
%%%%%%%%%%%%%%%%%%%%%%%%%%%
% For Arial (with lualatex)
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[no-math]{fontspec}
% \setmainfont{Arial}[
% % Options below necessary for Windows
% 	% Extension = .ttf,
% 	% UprightFont = *,
% 	% BoldFont = *bd,
% 	% ItalicFont = *i,
% 	% BoldItalicFont = *bi
% 	]
% In YAML
% ---
% mainfont: Arial
% ---
% \usepackage{lualatex-math}

% Math as text, but numerals do not mpa well becuase
% mathastext clahes with unicode-math
% \usepackage[italic]{mathastext}
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Estimating survival in left filtered data}
\author{Kevin Chen}
\date{Stat 256: Causal Inference (Fall 2021)}

\begin{document}
\maketitle

\fancyhead[R]{Kevin Chen}
\fancyhead[L]{}
\renewcommand{\headrulewidth}{0pt}

\onehalfspacing
\renewcommand{\arraystretch}{1.1}

\thispagestyle{empty}

\setlength{\columnseprule}{0pt}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The United Auto Workers-General Motors (UAW-GM) Cohort Study is a
longitudinal occupational cohort study established in the early 1980s to
study the health effects of metalworking fluids (Eisen et al. 1992,
2001). Metalworking fluids (MWF) are complex mixtures of fluids used in
industrial metalworking operations to lubricate and cool machinery and
parts. The three major classes of MWF are straight, soluble, and
synthetic metalworking fluids (Byers 2006). Possible routes of human
exposure include absorption through skin, inhalation or aerosols, and
ingestion of droplets.

A central concern in the analysis of occupational cohorts is the
potential for the healthy worker survivor effect (HWSE), the phenomenon
by which healthy individuals remain at work, while less healthy
individuals leave work -- possibly in response to exposure-related
health decline. In the presence of the HWSE, those with the highest
cumulative occupational exposures are also those who are less at risk of
disease. Thus, standard measures of association would show an inverse
relationship between occupational exposure and poor health outcomes
(Arrighi and Hertz-Picciotto 1994). The HWSE is an example of
time-varying confounding affected by past exposure. Previous studies
have attempted to assess the presence of the HWSE in observed data by
assessing so-called path-specific associations using Cox proportional
hazards modeling (Naimi et al. 2013; Garcia et al. 2017). However, these
measures of associations were themselves subject to the confounding
structures they sought to characterize.

If conditional sequential ignorability of exposure and censoring status
at each point in follow-up and positivity can be attained given
covariate history, then causal methods can be applied to account for the
HWSE. Past studies have applied causal methods capable of accounting for
time-varying confounding affected by past exposure to the study of MWF
exposures and cancer mortality outcomes in the UAW-GM Cohort Study
(Garcia et al. 2018; Izano et al. 2019), but the study of cancer
incidence outcomes is further problematized by the incomplete
observation of cancer incidence outcomes at every point of follow-up
over the study period. Nonetheless, we wish to make inferences about the
carcinogenicity of MWF exposure over an individual's lifetime starting
upon entry into the workforce. The UAW-GM Cohort included those hired
roughly between 1938 and 1985. However, cancer incidence reporting at
the Michigan Cancer Registry did not begin until 1985. Hence, our
observed cohort data is subject to \emph{left filtering}: cancer
incidence is the outcome of interest, but before 1985, both cancer
incidence status and time of cancer incidence are unknown. Observation
of the complete cancer incidence outcome vector over the study period
was conditional on an individual surviving to 1985 cancer-free.

In the presence of the HWSE, left filtering implies outcome
misclassification that is informative of true cancer status. As part of
her dissertation research, Izano (2017) conducted a quantitative bias
analysis for the estimation of survival curves using left filtered data
in the presence of HWSE. She simulated data compatible with the HWSE and
estimated cancer-free survival curves using an inverse probability of
treatment and censoring weighted Kaplan-Meier (WKM) estimator and the
WKM with an Aalen filter for left-filtering (AWKM) (Andersen et al.
1993; Xie and Liu 2005). Data were simulated under five different
scenarios. The survival curves under the following interventions were
computed or estimated in 250 replicates: (1) always exposed at work with
no censoring due to death and (2) never exposed at work with no
censoring due to death. For each intervention, three survival curves
were produced: (1) the true survival curve, (2) the WKM survival curve,
and (3) AWKM survival curve. Estimator bias was evaluated by comparing
the average of 250 estimates of to the average difference in cancer-free
survival under the two rules to the true average survival difference
when exposure and censoring were controlled deterministically.

The present project replicates the simulation and bias analyses
presented in Chapter 3 of Izano (2017), embeds the problem in the
non-parametric structural causal approach of Pearl (1995), and applies
the WKM and AWKM estimators to the estimation of colon cancer survival
under exposure/censoring rules in the UAW-GM Cohort.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{causal-model}{%
\subsection{Causal model}\label{causal-model}}

The UAW-GM Cohort data included person-year level exposure, outcome, and
covariate data starting at hire. To emulate the shape of the data for
this longitudinal cohort, we considered 20 years of data over time
indexed by years since hire. Notation representing the variables of
interest are presented in Table \ref{tab:variables}. The causal model
represents hypothetical relationships between variables over time
compatible with the theory underlying the HWSE in longitudinal
occupational cohort studies. At each time point, the effect of
cumulative exposure \(\bar A(t)\) on cancer incidence \(Y^(t)\) is
confounded by a path through employment status \(N(t)\) and underlying
health \(H(t)\) as well as the path through past exposure
\(\bar A(t - 1)\) and vital status \(D(t)\). These paths follow
straightforward logic: occupational exposure depends upon employment
status and past exposure; mortality status is affected by past exposure
and cancer history. Confounding by baseline covariates \(W\) is assumed
throughout.

\begin{table}
\caption{Descriptions of variables.}\label{tab:variables}
\begin{center}
\begin{tabular}{lp{0.5\linewidth}}
\toprule
 Variable  &  Description                                 \\ \midrule
 $R$       & Time until start of registry \\
 $W$       & Baseline covariates \\
 $S$       & Susceptibility to effects of metalworking fluid exposure \\
 $H(t)$    & Adverse health status at time $t$ \\
 $N(t)$    & Employment status at time $t$ \\
 $A(t)$    & Metalworking fluid exposure at time $t$ \\
 $D(t)$    & Mortality status at time $t$ \\
 $Y^*(t)$  & Cancer status at time $t$ \\
 $Y(t)$    & Observed Cancer status at time $t$ \\
 $t = \{1, 2, \ldots, 20\}$ & Time, indexed in years after hire \\
 \bottomrule
\end{tabular}
\end{center}\end{table}

Assume we have \(n = 50\,000\) iid units in \(X\) with
\[X_i(t) = \left(R_i = 0, W_i, S_i, \bar H_i(t), \bar N_i(t), \bar A_i(t), \bar a_i(t), \bar Y^*_i(t) = \bar Y_i(t) \right).\]
The bar notation to indicates variable history where
\(\bar X_i(t) = \left(X_i(k)\right)_{k = 1}^t\). In the case of
exposure, \(\bar A_i(t)\) is the indicator of whether unit \(i\) was
ever exposed through time \(t\). Note that true cancer status \(Y^*(t)\)
not observed until \(t \ge R\), after the start of the registry. Call
\(X\) the full data, where we have \(R = 0\) for all. In the observed
data \(X^\text{obs}\), we cannot assume \(R = 0\) for all;
susceptibility \(S\) and underlying health status \(H\) are not known:
\[X^\text{obs}_i(t) = \left(R_i, W_i, \bar N_i(t), \bar A_i(t), \bar a_i(t), \bar Y_i(t) \right).\]
Under the causal model, we assume the following non-parametric
structural equations: \[\begin{aligned}
R     & = f_R \left( U_R \right) \\
W     & = f_W \left( U_W \right) \\
S     & = f_S \left( U_S \right) \\
H(t)  & = f_{H(t)} \left( H(t - 1), U_{H(t)} \right) \\
N(t)  & = f_{N(t)} \left( W, N(t - 1), H(t), A(t - 1), U_{N(t)} \right) \\
A(t)  & = f_{A(t)} \left( W, \bar A(t - 1), N(t), U_{A(t)} \right) \\
D(t)  & = f_{D(t)} \left( W, \bar A(t - 1), D(t - 1), Y^*(t - 1), U_{D(t)} \right) \\
Y^*(t)& = f_{Y^*(t)} \left(W, S, H(t), \bar A (t), D(t), Y^*(t - 1), U_{Y^*(t)} \right) \\
Y(t) & = Y^*(t) \times \Ind{Y^*(\lfloor R \rfloor) = 0} \times \Ind{D(t) = 0} \ .
\end{aligned}\] The exogenous variables (errors)
\(U = \left( U_R, U_W, U_S, U_{H(t)}, U_{N(t)}, U_{A(t)}, U_{D(t)}, U_{Y^*(t)} \right)_{t = 1}^{T}\)
are mutually independent. Exposure status is a time-varying indicator;
let exposure history \(\bar A(t)\) be summarized as being ever-exposed
or not with \(\bar A(t) = \Ind{\sum_{k = 1}^t\Ind{A(k) = 1} > 0}\). The
outcome of interest is a survival outcome, so
\(Y^*(t - 1) = 1 \Rightarrow Y^*(t) = 1\). The observed outcome \(Y(t)\)
at time \(t\) is a function of true outcome status and the time points
delimiting left and right censoring. An abbreviated directed acyclic
graph (DAG) representing the causal relationships encoded in the
equations above is presented in Figure \ref{fig:dag}.

\begin{figure}
\caption{Directed acyclic graph representing the causal relationships encoded in the non-parametric structural equation model at time $t$.}
\label{fig:dag}
\begin{center}
\begin{tikzpicture}[>= stealth, auto, node distance = 2.25cm, semithick, fill=white, inner sep=0pt]
\tikzstyle{every state}=[shape = circle, align = center, draw = none]
\node[state] (H0) {$H(t - 1)$};
\node[state] (N0) [below right of=H0] {$N(t - 1)$};
\node[state] (A0) [below right of=N0] {$\bar A(t - 1)$};
\node[state] (D0) [below right of=A0] {$D(t - 1)$};
\node[state] (Y0) [below right of=D0] {$Y^*(t - 1)$};
\node[state] (W) [above right of=H0] {$W$};
\node[state] (H) [right of=H0, node distance = 4cm] {$H(t)$};
\node[state] (N) [below right of=H, node distance = 4cm] {$N(t)$};
\node[state] (A) [below right of=N, node distance = 4cm] {$\bar A(t)$};
\node[state] (D) [above right of=N, node distance = 3cm] {$D(t)$};
\node[state] (Y) [above right of=A, node distance = 3cm] {$Y^*(t)$};
\node[state] (S) [above right of=D, node distance = 3cm] {$S$};
%
\path[->] (H0) edge (N0);
\path[->] (N0) edge (A0);
\path[->] (A0) edge [bend right=20] (Y0);
\path[->] (H0) edge (H);
\path[->] (W) edge [bend right=10] (N);
\path[->] (N0) edge (N);
\path[->] (H) edge (N);
\path[->] (A0) edge (N);
\path[->] (W) edge [bend right=15] (A);
\path[->] (A0) edge (A);
\path[->] (N) edge (A);
\path[->] (W) edge (D);
\path[->] (A0) edge [bend left=15] (D);
\path[->] (D0) edge [bend right=15] (D);
\path[->] (Y0) edge [bend right=10] (D);
\path[->] (W) edge [bend left=5] (Y);
\path[->] (S) edge [bend left=30] (Y);
\path[->] (S) edge [bend left=15] (Y0);
\path[->] (H) edge (Y);
\path[->] (A) edge (Y);
\path[->] (D) edge (Y);
\path[->] (Y0) edge (Y);
\end{tikzpicture}
\end{center}
\end{figure}

Figure \ref{fig:dag} clarifies our conceptualization of the HWSE. At
each time point \(t\), the cumulative effect of exposure \(\bar A(t)\)
on cancer incidence \(Y^*(t)\) is confounded by employment status
\(N(t)\) through the backdoor path
\(\bar A(t) \leftarrow N(t) \leftarrow H(t) \rightarrow Y^*(t)\). In the
absence of observed data on health status \(H(t)\), an analyst may be
tempted to condition on employment status \(N(t)\), but doing so would
introduce collider bias while blocking the causal path between past
exposure \(\bar A(t - 1)\) and cancer \(Y^*(t)\). Furthermore, an
analysis starting at an arbitrary time point after the time origin
(hire) would be tantamount to conditioning on those still alive at that
time, which would also result in both collider bias and the conditioning
on nodes on the causal path between the exposure and outcome of
interest.

\hypertarget{simulation}{%
\subsection{Simulation}\label{simulation}}

To generate data compatible with our structural causal model, we imposed
parametric relationships between the variables. For the \(n = 50\,000\)
units over \(T = 20\) years, we have:

\begin{itemize}
\tightlist
\item
  \(U_j \overset{\text{iid}}{\sim} \text{uniform}\,[0,1]\) for all \(j\)
\item
  In full data \(R = 0\) otherwise \(R \sim \text{uniform}\,[0, 30]\)
\item
  \(W = \Ind{U_W \le p_W} \sim \text{Bernoulli}\left( p_W \right)\)
\item
  \(S = \Ind{U_S \le p_S} \sim \text{Bernoulli}\,(p_S)\)
\item
  If \(H(t - 1) = 1\), then \(H(t) = 1\) otherwise
  \(H(t) = \Ind{U_{H(t)} \le p_H} \sim \text{Bernoulli}\,(p_H)\)
\item
  if \(N(t - 1) = 0\) then \(N(t) = 0\) otherwise
  \[N(t) \sim \text{Bernoulli}\left\{
    \text{logit}\,\left(
        \beta_0^N
        + \beta_W^N W
        + \beta_H^N H(t)
        + \beta_A^N A(t - 1) \times \Ind{t > 1}
        + U_{N(t)}
    \right)
    \right\}\]
\item
  If \(N(t) = 0\) then \(A(t) = 0\) otherwise
  \[A(t) \sim \text{Bernoulli}\left\{
    \text{logit}\,\left(
        \left( \beta_0^A + \beta_W^A W\right) \times \Ind{t = 1}
        + \beta_A^A A(t - 1) \times \Ind{t > 1}
        + U_{A(t)}
    \right)
    \right\}\]
\item
  If \(D(t - 1) = 1\) then \(D(t) = 1\) otherwise
  \[D(t) \sim \text{Bernoulli}\left\{
    \text{logit}\,\left(\begin{aligned}
        \beta_0^D
        + \beta_W^D W
        + \beta_{\bar A}^D \bar A(t - 1) \times \Ind{t > 1} \\
        + \beta_{\bar Y}^D \sum_{k = 1}^{t - 1} Y^*(k) \times \Ind{t > 1}
        + U_{D(t)}
    \end{aligned}\right)
    \right\}\]
\item
  If \(Y^*(t - 1) = 1\) then \(Y^*(t) = 1\) otherwise
  \[Y^*(t) \sim \text{Bernoulli}\left\{
    \text{logit}\,\left(\begin{aligned}
        \beta_0^Y
        + \beta_W^Y W
        + \beta_{A}^Y A(t)
        + \beta_{\bar A}^Y \bar A(t - 1) \times \Ind{t > 1} \\
        + \beta_{S}^Y S \times \bar A(t)
        + \beta_{H}^Y H(t)
        + U_{Y^*(t)}
    \end{aligned}\right)
    \right\}\]
\item
  If \(t < R\) then \(Y(t) = 0\)
\item
  If \(t \ge R\) then
  \(Y(t) = Y^*(t) \times \Ind{Y^*(\lfloor R \rfloor) = 0} \times \Ind{D(t) = 0}\).
\end{itemize}

Five sets of data were generated using these equations to represent five
scenarios. Scenario 1 represents the base case where 10\% of workers are
susceptible to exposure-related effects, the odds ratio of mortality
each additional year following cancer diagnosis is about 1.6, and there
is moderate time-varying confounding by health status. In scenario 2, we
have greater cancer-related mortality by increasing
\(\beta_{\bar Y}^D\). In scenario 3, we increase \(p_S\), the proportion
of the study population susceptible to the carcinogenic effects of MWF
exposure. In scenario 4, we consider greater time-varying confounding by
health status by increasing \(\beta_H^N\) and \(\beta_H^Y\). In the last
scenario, we have greater background cancer incidence by increasing
\(\beta_0^Y\). The sets of parameters used in the five scenarios are
presented in Table \ref{tab:params}.

\begin{table}[h]
\caption{Simulation parameters.}
\label{tab:params}
\begin{center}
\begin{tabular}{rrrrrr}
  \toprule
Parameter & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
$p_S$ & 0.10 & 0.10 & \textbf{0.20} & 0.10 & 0.10 \\ 
  $p_W$ & 0.20 & 0.20 & 0.20 & 0.20 & 0.20 \\ 
  $p_H$ & 0.30 & 0.30 & 0.30 & 0.30 & 0.30 \\ 
  $\beta_0^N$ & 3.00 & 3.00 & 3.00 & 3.00 & 3.00 \\ 
  $\beta_W^N$ & -0.10 & -0.10 & -0.10 & -0.10 & -0.10 \\ 
  $\beta_H^N$ & -0.50 & -0.50 & -0.50 & \textbf{-1.50} & -0.50 \\ 
  $\beta_A^N$ & -1.50 & -1.50 & -1.50 & -1.50 & -1.50 \\ 
  $\beta_0^A$ & -1.50 & -1.50 & -1.50 & -1.50 & -1.50 \\ 
  $\beta_W^A$ & -0.50 & -0.50 & -0.50 & -0.50 & -0.50 \\ 
  $\beta_A^A$ & 2.50 & 2.50 & 2.50 & 2.50 & 2.50 \\ 
  $\beta_0^D$ & -5.50 & -5.50 & -5.50 & -5.50 & -5.50 \\ 
  $\beta_W^D$ & 1.00 & 1.00 & 1.00 & 1.00 & 1.00 \\ 
  $\beta_{\bar A}^D$ & 0.50 & 0.50 & 0.50 & 0.50 & 0.50 \\ 
  $\beta_{\bar Y}^D$ & 0.50 & \textbf{2.00} & 0.50 & 0.50 & 0.50 \\ 
  $\beta_0^Y$ & -7.00 & -7.00 & -7.00 & -7.00 & \textbf{-6.00} \\ 
  $\beta_W^Y$ & 2.00 & 2.00 & 2.00 & 2.00 & 2.00 \\ 
  $\beta_A^Y$ & 0.25 & 0.25 & 0.25 & 0.25 & 0.25 \\ 
  $\beta_{\bar A}^Y$ & 0.20 & 0.20 & 0.20 & 0.20 & 0.20 \\ 
  $\beta_H^Y$ & 0.70 & 0.70 & 0.70 & \textbf{1.70} & 0.70 \\ 
  $\beta_S^Y$ & 0.30 & 0.30 & 0.30 & 0.30 & 0.30 \\ 
   \bottomrule
\end{tabular}
\end{center}
\end{table}

\hypertarget{interventions-potential-outcomes-target-parameters-and-estimation}{%
\subsection{Interventions, potential outcomes, target parameters, and
estimation}\label{interventions-potential-outcomes-target-parameters-and-estimation}}

The substantive question of interest was the causal effect of
occupational exposure to MWF on cancer incidence risk. Since
occupational MWF exposure occurs only when individuals are at work, we
defined dynamic exposure regimes that depend on employment status. Under
rule \(a_0\), set \(D(t) = 0\), and set \(A(t) = 0\) while \(N(t) = 1\).
Under rule \(a_1\), set \(D(t) = 0\), and set \(A(t) = 1\) while
\(N(t) = 1\). Under both rules, we prevented censoring by death as if it
were intervenable. The causal effect was defined by contrasting the
survival function \(S_{a_1}(t) = 1 - \E{Y_{a_1}(t)}\) under rule \(a_1\)
to \(S_{a_0}(t) = 1 - \E{Y_{a_0}(t)}\) that under rule \(a_0\). Note
that this causal estimand was defined over \emph{a priori
counterfactuals} not observable in the real world (Frangakis and Rubin
2002). This approach is standard in epidemiologic studies, however.

The survival function expresses the probability that a person following
rule \(a\) is cancer-free at the end of time point \(t\). The expected
time until cancer under rule \(a\) is
\(\mu_{a} = \sum_0^K S_{a}(t) \, dt\). The parameter used in the
estimation of bias was the summary measure
\(\psi = \mu_{a_1} - \mu_{a_0}\), the difference in expected time until
event under two different interventions over 20 years of follow-up under
five different data generating scenarios. Bias was evaluated by
comparing estimates of \(\psi\) to its true value in 250 simulations per
scenario (the original analysis performed 500). The true value was
calculated by simulating the full data for five hundred thousand
individuals (the original analysis used one million) with rules \(a_0\)
and \(a_1\) applied deterministically. Estimates of \(\psi\) were
obtained by first estimating the survival curves \(S_a(t)\) using two
estimators: the inverse probability weighted Kaplan-Meier estimator
(WKM) and the Aalen-filtered WKM (AWKM). These survival estimators are
detailed in the following section.

\hypertarget{kaplan-meier-estimator-and-extensions}{%
\subsection{Kaplan-Meier estimator and
extensions}\label{kaplan-meier-estimator-and-extensions}}

To estimate survival, we applied extensions of the widely-known
Kaplan-Meier (KM) estimator for survival (Kaplan and Meier 1958). First,
we review the estimator of Xie and Liu (2005), an extension of the KM
estimator where units are weighted by the inverse probability of
treatment. The standard KM estimator requires counting up the number of
cases \(c^0_{a}(t)\) that occurred in interval \((t - 1, t]\) and the
number of units at risk \(R^0_{a}(t)\) in that interval at all event
times \(t\). Assuming cancer status was assessed at the end of regular
intervals \(t = 1, \ldots, K\), we have: \[\begin{aligned}
c^0_{a} (t)
& = \sum_i^n \Ind{Y_i(t) = 1} \times \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)} \\
R^0_{a} (t)
& = \sum_i^n \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)}.
\end{aligned}\] The standard survival estimator is \[\hat S^0_{a}(t) =
\begin{cases}
1 & \text{if } t < t_1 \\
\prod_{j \le t} \left(1 - \frac{c^0_{a}(j)}{R^0_{a}(j)}\right) & \text{if } t \ge t_1 \\
\end{cases}\] where \(t_1\) is the first event time.

In observational studies, survival contrasts estimated using the
standard KM estimator are biased for the true causal survival contrast.
However, if conditional ignorability and positivity are attained, the
inverse probability weighted KM (WKM) estimator of Xie and Liu (2005)
yields unbiased estimates of the true causal survival curve. The WKM
estimator augments the standard KM estimator by weighting units at time
\(t\) by \(w_{i, a}(t)\) the inverse probability of treatment:
\[\begin{aligned}
c^w_{a} (t)
& = \sum_i^n w_{i, a}(t) \times \Ind{Y_i(t) = 1} \times \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)}\\
R^w_{a} (t)
& = \sum_i^n w_{i, a}(t) \times \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)}
\end{aligned}\] The WKM survival estimator for rule \(a\) is
\[\hat S^w_{a}(t) =
\begin{cases}
1 & \text{if } t < t_1 \\
\prod_{j \le t} \left(1 - \frac{c^w_{a}(j)}{R^w_{a}(j)}\right) & \text{if } t \ge t_1 \\
\end{cases}\] where \(t_1\) is the first event time.

Finally, to account for (uninformative) left filtering, we applied the
Aalen filter, which considers only the units at time \(t\) for which the
outcome is observed: \[\begin{aligned}
c_{a} (t)
& = \sum_i^n w_{i, a}(t) \times \Ind{Y_i(t) = 1} \times \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)} \times \Ind{t \ge R_i} \\
R_{a} (t)
& = \sum_i^n w_{i, a}(t) \times \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)} \times \Ind{t \ge R_i}
\end{aligned}\] The Aalen-filtered WKM (AWKM) estimator for rule \(a\)
is \[\hat S_{a}(t) =
\begin{cases}
1 & \text{if } t < t_1 \\
\prod_{j \le t} \left(1 - \frac{c_{a}(j)}{R_{a}(j)}\right) & \text{if } t \ge t_1 \\
\end{cases}\] where \(t_1\) is the first event time.

In the full data, the WKM and AWKM estimators are equivalent, and
identification is achieved under overlap (positivity) and sequential
ignorability (randomization) assumptions: \[\begin{aligned}
Y^*_{a, \bar d = 0} (t')   & \indep A(t) \mid W,\ \bar A(t - 1) = \bar a(t - 1),\ D(t - 1) = 0,\ N(t) = 1\\
Y^*_{a, \bar d = 0} (t') & \indep D(t) \mid W,\ D(t - 1) = 0,\ Y^*(t - 1) = 0,\ \bar A(t - 1) = \bar a(t - 1)
\end{aligned}\] for all times \(t'\ge t\), and
\[0 < \Prob{ A(t) = 1 \mid W,\ \bar A(t - 1) = \bar a(t - 1),\ D(t - 1) = 0,\ N(t) = 1} < 1\]
\[0 < \Prob{D(t) = 0 \mid W,\ D(t - 1) = 0,\ Y^*(t - 1) = 0,\ \bar A(t - 1) = \bar a(t - 1)} < 1.\]

Graphical representations of the first and second components of the
ignorability assumption are presented in Figures
\ref{fig:ignorability_a} and \ref{fig:ignorability_d} where conditioning
on boxed variables are represented by the removal of edges pointing away
from those variables. The resulting graphs show the fulfillment of
Pearl's backdoor criterion for the estimation of the causal effects of
\(\bar A(t)\) on \(Y^*(t)\) and \(D(t)\) on \(Y^*(t)\), respectively.
Thus, the causal effect of the joint intervention on
\(\big(\bar A(t), D(t)\big)\) at each time \(t\) is identified. Causal
identification is not attained when true cancer status \(Y^*(t)\) is not
known.

\begin{figure}
\caption{Directed acyclic graph representing the causal relationships encoded in the non-parametric structural equation model at time $t$ after conditioning on $\{W,\ \bar A(t - 1),\ D(t - 1),\ N(t)\}$.}
\label{fig:ignorability_a}
\begin{center}
\begin{tikzpicture}[>= stealth, auto, node distance = 2.25cm, semithick, fill=white, inner sep=0pt]
\tikzstyle{every state}=[shape = circle, align = center, draw = none]
\node[state] (H0) {$H(t - 1)$};
\node[state] (N0) [below right of=H0] {$N(t - 1)$};
\node[state] (A0) [shape=rectangle, draw=black, inner sep=2.5pt, below right of=N0] {$\bar A(t - 1)$};
\node[state] (D0) [shape=rectangle, draw=black, inner sep=2.5pt, below right of=A0] {$D(t - 1)$};
\node[state] (Y0) [below right of=D0] {$Y^*(t - 1)$};
\node[state] (W) [shape=rectangle, draw=black, inner sep=2.5pt, above right of=H0] {$W$};
\node[state] (H) [right of=H0, node distance = 4cm] {$H(t)$};
\node[state] (N) [shape=rectangle, draw=black, inner sep=2.5pt, below right of=H, node distance = 4cm] {$N(t)$};
\node[state] (A) [below right of=N, node distance = 4cm] {$\bar A(t)$};
\node[state] (D) [above right of=N, node distance = 3cm] {$D(t)$};
\node[state] (Y) [above right of=A, node distance = 3cm] {$Y^*(t)$};
\node[state] (S) [above right of=D, node distance = 3cm] {$S$};
%
\path[->] (H0) edge (N0);
\path[->] (N0) edge (A0);
\path[->] (H0) edge (H);
\path[->] (N0) edge (N);
\path[->] (H) edge (N);
\path[->] (S) edge [bend left=15] (Y0);
\path[->] (S) edge [bend left=30] (Y);
\path[->] (H) edge (Y);
\path[->] (A) edge (Y);
\path[->] (D) edge (Y);
\path[->] (Y0) edge [bend right=10] (D);
\path[->] (Y0) edge (Y);
\end{tikzpicture}
\end{center}
\end{figure}

\begin{figure}
\caption{Directed acyclic graph representing the causal relationships encoded in the non-parametric structural equation model at time $t$ after conditioning on $\{W,\ \bar A(t - 1),\ D(t - 1),\ Y^*(t - 1)\}$.}
\label{fig:ignorability_d}
\begin{center}
\begin{tikzpicture}[>= stealth, auto, node distance = 2.25cm, semithick, fill=white, inner sep=0pt]
\tikzstyle{every state}=[shape = circle, align = center, draw = none]
\node[state] (H0) {$H(t - 1)$};
\node[state] (N0) [below right of=H0] {$N(t - 1)$};
\node[state] (A0) [shape=rectangle, draw=black, inner sep=2pt, below right of=N0] {$\bar A(t - 1)$};
\node[state] (D0) [shape=rectangle, draw=black, inner sep=2pt, below right of=A0] {$D(t - 1)$};
\node[state] (Y0) [shape=rectangle, draw=black, inner sep=2pt, below right of=D0] {$Y^*(t - 1)$};
\node[state] (W) [shape=rectangle, draw=black, inner sep=2pt, above right of=H0] {$W$};
\node[state] (H) [right of=H0, node distance = 4cm] {$H(t)$};
\node[state] (N) [below right of=H, node distance = 4cm] {$N(t)$};
\node[state] (A) [below right of=N, node distance = 4cm] {$\bar A(t)$};
\node[state] (D) [above right of=N, node distance = 3cm] {$D(t)$};
\node[state] (Y) [above right of=A, node distance = 3cm] {$Y^*(t)$};
\node[state] (S) [above right of=D, node distance = 3cm] {$S$};
%
\path[->] (H0) edge (N0);
\path[->] (N0) edge (A0);
\path[->] (H0) edge (H);
%\path[->] (W) edge [bend right=10] (N);
\path[->] (N0) edge (N);
\path[->] (H) edge (N);
%\path[->] (A0) edge (N);
%\path[->] (W) edge [bend right=15] (A);
%\path[->] (A0) edge (A);
\path[->] (N) edge (A);
%\path[->] (W) edge (D);
%\path[->] (A0) edge [bend left=15] (D);
%\path[->] (D0) edge [bend right=15] (D);
%\path[->] (Y0) edge [bend right=10] (D);
%\path[->] (W) edge [bend left=5] (Y);
\path[->] (S) edge [bend left=30] (Y);
\path[->] (S) edge [bend left=15] (Y0);
\path[->] (H) edge (Y);
\path[->] (A) edge (Y);
\path[->] (D) edge (Y);
%\path[->] (Y0) edge (Y);
\end{tikzpicture}
\end{center}
\end{figure}

\hypertarget{estimation-of-weights}{%
\subsection{Estimation of weights}\label{estimation-of-weights}}

To estimate the weights for the WKM and AWKM estimators, we fit two
logistic regressions at each time point \(t = 1, \ldots 20\):
\[\begin{aligned}
\text{logit} \left(\Prob{
    A(t) = 1
    \mid W,\ \bar A(t - 1),\ D(t - 1) = 0,\ N(t) = 1
    }\right)
& = \alpha_0 + W \alpha_1 + A(t - 1) \alpha_2 \\
\text{logit} \left(\Prob{
    D(t) = 1
    \mid W,\ D(t - 1) = 0,\ Y(t - 1) = 0,\ \bar A(t - 1)}\right)
    & = \beta_0 + W \beta_1 + \bar A(t - 1) \beta_2
\end{aligned}\] The first was fit on data for those alive and at work at
time \(t\). The second was among those alive and (observed to be)
cancer-free. For each unit at time \(t\), the weight was calculated by
taking the inverse of the cumulative probability of following the
exposure rule and remaining uncensored: \[\hat w_{a} (t) =
    \left[
    \prod^t_{j = 1}
    \begin{aligned}
    \widehat {\mathbb P} \left\{
    A(j) = a(j) \mid W,\ \bar A(j - 1) = \bar a (j - 1),\ D(j - 1) = 0,\ N(j) = 1
    \right\} \times \\
    \widehat {\mathbb P} \left\{D(j) = 0
    \mid W,\ D(j - 1) = 0,\ Y(j - 1) = 0,\ \bar A(j) = \bar a (j) \right\}
    \end{aligned}\right]^{-1}.\]

\hypertarget{results}{%
\section{Results}\label{results}}

Figure \ref{fig:survival} presents the true survival curves as well as
the WKM and AWKM survival curves averaged over 250 replications for each
intervention rule and scenario. Qualitatively, the WKM estimator
consistently over-estimated survival whereas the the AWKM survival curve
was much closer to the truth. The bias of the AWKM survival estimator
appeared to be larger in earlier follow-up and smaller for later
follow-up time points. The bias of both the the WKM and the AWKM
estimators appeared largest in Scenario 5.

\begin{figure}
\caption{Cancer-free survival over time since hire in five simulation scenarios. The true (discrete) survival curve is represented by the solid lines. The average inverse probability weighted Kaplan-Meier (WKM) survival curve is represented by the dashed-line with short dashes. The average Aalen-filtered inverse probability weighted Kaplan-Meier (AWKM) survival curve is represented by the dashed-line with long dashes. Estimated survival curves were averaged over 250 replicates. Salmon color indicates survival and survival estimates under rule $a_0$ when workers are always unexposed. Cyan color indicates those under rule $a_1$ when workers are always exposed while employed.}
\label{fig:survival}
\begin{center}
\includegraphics{/Users/kevinchen/Documents/2021 Fall/256 - Causal Inference/256-final/reports/resources/results.pdf}
\end{center}
\end{figure}

Table \ref{tab:survival} presents true and estimated average cancer-free
survival times under each intervention rule and scenario. Table
\ref{tab:differences} presents differences in survival time contrasting
rule \(a_1\) to rule \(a_0\). Table \ref{tab:bias} presents estimates of
the bias of the WKM and AWKM estimators for \(\psi\), the difference in
average cancer-free survival time over 20 years of follow-up. These
numeric results were consistent with the qualitative interpretations of
Figure \ref{fig:survival}. The WKM estimator over-estimated the
difference in cancer-free survival time, resulting in bias toward the
null, whereas the AWKM estimator under-estimates the cancer-free
survival, resulting in bias away from the null. In every scenario, the
bias of the WKM-derived contrast was several times larger in magnitude
than that of the AWKM-derived contrast. The bias of both estimators was
greatest for Scenario 5.

The qualitative results here were consistent with those of Izano (2017).
However, true and estimated survival in the present analysis was larger
than those found previously. Furthermore, the true and estimator average
mean differences in survival were smaller in magnitude in the present
case. The magnitudes of the bias estimates were also smaller.

\begin{table}[H]
\centering
\caption{True cancer-free survival time $\mu_{a}$ over 20-year follow-up and estimator averages over 250 replicates.} 
\label{tab:survival}
\begin{tabular}{llrrrrr}
  \toprule
Rule & Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
$a_0$ & Truth & 19.08 & 19.10 & 19.10 & 18.02 & 17.83 \\ 
   & WKM & 19.52 & 19.51 & 19.52 & 18.91 & 18.90 \\ 
   & AWKM & 19.01 & 19.00 & 19.01 & 17.80 & 17.71 \\ 
   \midrule
$a_1$ & Truth & 18.80 & 18.79 & 18.76 & 17.54 & 17.29 \\ 
   & WKM & 19.41 & 19.39 & 19.39 & 18.68 & 18.67 \\ 
   & AWKM & 18.76 & 18.76 & 18.74 & 17.30 & 17.22 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Difference in average cancer-free survival time over 20-year follow-up comparing rule $a_1$ always exposed to rule $a_0$ never exposed at work: true value $\psi$ and estimator averages over 250 replicates.} 
\label{tab:differences}
\begin{tabular}{lrrrrr}
  \toprule
Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
Truth & -0.28 & -0.31 & -0.34 & -0.48 & -0.54 \\ 
  WKM & -0.11 & -0.12 & -0.13 & -0.23 & -0.23 \\ 
  AWKM & -0.25 & -0.24 & -0.27 & -0.50 & -0.50 \\ 
   \bottomrule
\end{tabular}
\end{table}
\begin{table}[H]
\centering
\caption{Bias estimates of estimators for $\psi$, the difference in average cancer-free survival time over 20 years of follow-up.} 
\label{tab:bias}
\begin{tabular}{lrrrrr}
  \toprule
Estimator & Scenario 1 & Scenario 2 & Scenario 3 & Scenario 4 & Scenario 5 \\ 
  \midrule
WKM & 0.17 & 0.19 & 0.21 & 0.25 & 0.31 \\ 
  AWKM & 0.03 & 0.07 & 0.07 & -0.02 & 0.04 \\ 
   \bottomrule
\end{tabular}
\end{table}

\hypertarget{application-to-the-uaw-gm-cohort}{%
\section{Application to the UAW-GM
Cohort}\label{application-to-the-uaw-gm-cohort}}

In the simulation study, we showed that under several scenarios
compatible with our hypothetical causal structure, the AWKM survival
estimator had small bias compared to the WKM estimator. The bias was
smaller when the cumulative incidence of the outcome was low and at
later follow-up time points. Next, we estimated cancer-free survival in
a real-world context. Using data from the UAW-GM Cohort study, we
followed \(38\,553\) individuals starting from hire to 40 years after
hire for incidence of colon cancer. This duration of follow-up spans the
vast majority of individuals' working lifetimes. As in the simulation,
the UAW-GM data were longitudinal data with baseline covariates,
time-varying covariates, and a survival outcome.

The exposures of interest were MWF of three types: straight, soluble,
and synthetic (Byers 2006; F. Mirer 2003; F. E. Mirer 2010). Straight
MWFs are hydrocarbon-based fluids whose use became widespread by the
1920s. They continue to be widely used because of their simple
formulation. In straight MWFs, hydrocarbons of different lengths are
mixed together with other additives to attain different properties.
Straight MWFs contain polycyclic aromatic hydrocarbons, long known to be
carcinogenic (IARC 1973). Soluble oils are water-based oil emulsions
first introduced in response to rising oil prices. They now make up the
largest market share of MWFs (Childers 2006). Soluble oils are
vulnerable to microbial contamination, so the addition of biocides is
needed. However, their high lubricity makes soluble MWFs the most
popular fluid type. Synthetic MWFs have the best toxicological profile,
have no oil, and have a higher resistance to microbial growth. They were
introduced into the MWF market in the second half of the 20th C., but
fail to out-perform soluble MWFs in metalworking applications. Soluble
and synthetic MWFs contain biocides, corrosion inhibitors, and
chlorinated compounds, some classified as carcinogenic by the IARC (IARC
1987).

The outcome of interest was colon cancer incidence. There is little past
research linking colon cancer to MWF exposures, but there is some
evidence suggesting that straight MWFs cause colon cancer (Izano et al.
2019). We obtained cancer incidence data via linkage to the Michigan
Cancer Registry, which recorded cancer incidence cases starting on
January 1, 1985. The cohort is comprised of individuals hired between
1938 and 1985. Cancer-free survival to the start of the registry was a
left-filtering process possibly in the presence of the HWSE as
investigated in the simulations. Over the 40 year follow-up period,
vital status was obtained through the Social Security Administration,
the National Death Index, as well as records provided by the UAW.
Baseline confounders included race, sex, plant, and year of hire.
Time-varying confounders included age, cumulative time off, and exposure
to the metalworking fluids not under study. These terms were included in
the estimation of both the treatment and censoring mechanisms, which
were estimated with stratification on every year of follow-up. The
exposure rules of interest were identical to those in the simulation
study: \(a_0\) never exposed and \(a_1\) always exposed while at work.
We assumed that the effect of exposure on health status, mortality,
cancer incidence, etc occurs after a lag of 15 years. Similarly, we
assumed that the effects of cumulative time off occur after a 15 year
lag. Counterfactual survival under rules \(a_0\) and \(a_1\) were
estimated using the WKM and AWKM estimators. Summary statistics for the
full study population and the colon cancer cases are presented in Table
\ref{tab:popchar}.

\begin{table}[h]
\caption{Study population characteristics.}
\label{tab:popchar}
\begin{center}\begin{tabular}{lrlrl}
\toprule
&       \multicolumn{2}{c}{Full cohort}       & \multicolumn{2}{c}{Colon cancer cases}      \\
\midrule
$n$ (person-years) 
            &     38 553 & (1 402 372)       & 267 & (9 635)        \\
Race (\%) \\
\hspace{1.5em} Black   &      7 133 & (18.5)                &  83 & (31.1)          \\
\hspace{1.5em} White   &     31 420 & (81.5)                & 184 & (68.9)          \\
Sex (\%) \\
\hspace{1.5em} Female  &      4 757 & (12.3)                &  49 & (18.4)      \\
\hspace{1.5em} Male    &     33 796 & (87.7)                & 218 & (81.6)      \\
Plant (\%) \\
\hspace{1.5em} Plant 1 &      9 092 & (23.6)                &  66 & (24.7)      \\
\hspace{1.5em} Plant 2 &     17 090 & (44.3)                & 107 & (40.1)      \\
\hspace{1.5em} Plant 3 &     12 371 & (32.1)                &  94 & (35.2)          \\
\midrule
Ever exposed to MWF (\%)
                        &     22193 & (57.6)                & 104 & (39.0)          \\
Year of hire (mean (\textsc{sd}))
                        &     1963 & (11.62)            & 1968 & (8.79)         \\
Age (mean (\textsc{sd}))
                        &     63.38 & (9.08)                & 67.78 & (9.27)        \\
Cumulative time off (mean (\textsc{sd}))
                        &     0.04 & (0.14)               & 0.05 & (0.15)       \\
\bottomrule
\end{tabular}\end{center}
\end{table}

\hypertarget{assumptions}{%
\subsection{Assumptions}\label{assumptions}}

Since we are working with observational data, the evaluation of the
no-interference, causal consistency, ignorability, and positivity
assumptions are critical for causal inference. The stability of our
estimation depends on positivity, which we assessed qualitatively by
examining the distribution of the weights. The no-interference
assumption may be problematized by the fact that there were a finite
number of job types in the factory setting. If one worker operates a
particular metalworking machine, then the the other workers are not able
to operate that machine at that time. Instead, they may be assigned to
assembly tasks, which have lower MWF exposure opportunities. That said,
since these factories were quite large, there may be approximate
independence. The consistency assumption is also problematic. The MWFs
of interest are complex chemical mixtures whose composition changes by
design and by nature. Over the last several decades, the formulation of
MWFs has changed significantly in reaction to performance needs and
toxicity concerns (F. Mirer 2003; Byers 2006). The composition of MWFs
also undergoes unintentional changes because of the nature of their use:
MWFs are often applied in contexts where contamination by other
substances and microbes is possible and chemical changes due to heat and
pressure are likely. Indeed, there are substantial concerns over the
carcinogenicity of the chemical species formed in the MWF mixtures that
were not originally added (Hidajat et al. 2020). Concerns regarding the
consistency assumption may be abrogated in part by adequate adjustment
for secular and factory-level characteristics.

Another key assumption meriting discussion is that of conditional
ignorability. In order to achieve identification, even in the absence of
left filtering, we need to have conditionally ignorable future exposure
status and ignorable future censoring status at each time point given
past data. In occupational cohorts, employment status and health history
are strong predictors of future death (Häfner 1987; Halliday 2014;
Laliotis and Stavropoulou 2018). Logically, major causes of death first
act through employment status before they precipitate death. This
dynamic is actually a key component in the setup for HWSE. We are
therefore relatively confident that conditional ignorability of
censoring due to death is attained given covariate, exposure, and cancer
history. Our confidence in the conditional ignorability of exposure
given history was not as strong. In particular, workers may be assigned
to certain tasks based on their specific skills and knowledge, which may
be associated with structural privileges that confer a lower risk of
deleterious health outcomes. The potential magnitude of this
uncontrolled confounding is likely to be limited, however. The education
level of the workers in the cohort was homogeneous, and all cohort
members were members of the UAW union, which had uniform procedures in
place for equitable access to training, wages, and career advancement
(Harbison 1950; Barnard 2005). The presence of UAW policies support the
assertion that given time since hire, job types (and therefore
exposures) were randomly allocated.

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

\begin{itemize}
\tightlist
\item
  Foo
\end{itemize}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\begin{itemize}
\tightlist
\item
  FU in GM is much longer than 20 years
\item
  Exclusion restriction on \(H\) so that its parents do not include
  \(A\) is not necessary for identification in full data
\item
  Exclusion restriction on \(D\) so that its parents do not include
  \(H\) not needed if \(H\) is observed and conditioned upon
\item
  Not clear if these were for convenience or what
\item
  What if we increase FU? Lags?
\item
  Greater fineness in exposure definition? Different interventions?
\item
  No way to verify to what extent the simulation informs the real data
  generating system
\item
  Model specification, double robustness
\end{itemize}

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Andersen_1993}{}%
Andersen, P. K., Ø. Borgan, R. D. Gill, and N. Keiding. 1993.
\emph{Statistical Models Based on Counting Processes}. Springer Series
in Statistics. Springer, New York, NY.
\url{https://books.google.com/books?id=kBnvAAAAMAAJ}.

\leavevmode\hypertarget{ref-Arrighi_1994}{}%
Arrighi, H. Michael, and Irva Hertz-Picciotto. 1994. {``The Evolving
Concept of the Healthy Worker Survivor Effect.''} \emph{Epidemiology} 5
(2): 189--96. \url{http://www.jstor.org/stable/3702361}.

\leavevmode\hypertarget{ref-Barnard_2005}{}%
Barnard, John. 2005. \emph{American Vanguard: The United Auto Workers
During the Reuther Years, 1935-1970}. Wayne State University Press.

\leavevmode\hypertarget{ref-Byers_2006}{}%
Byers, Jerry P. 2006. \emph{Metalworking Fluids}. CRC Press.

\leavevmode\hypertarget{ref-Childers_2006}{}%
Childers, Jean. 2006. {``The Chemistry of Metalworking Fluids.''} In
\emph{Metalworking Fluids}. CRC Press.

\leavevmode\hypertarget{ref-Eisen_2001}{}%
Eisen, Ellen A, Judith Bardin, Rebecca Gore, Susan R Woskie, Marilyn F
Hallock, and Richard R Monson. 2001. {``Exposure-Response Models Based
on Extended Follow-up of a Cohort Mortality Study in the Automobile
Industry.''} \emph{Scandinavian Journal of Work, Environment \& Health}
27 (4): 240--49.

\leavevmode\hypertarget{ref-Eisen_1992}{}%
Eisen, Ellen A, Paige E Tolbert, Richard R Monson, and Thomas J Smith.
1992. {``Mortality Studies of Machining Fluid Exposure in the Automobile
Industry {I}: A Standardized Mortality Ratio Analysis.''} \emph{American
Journal of Industrial Medicine} 22 (6): 809--24.

\leavevmode\hypertarget{ref-Frangakis_2002}{}%
Frangakis, Constantine E, and Donald B Rubin. 2002. {``Principal
Stratification in Causal Inference.''} \emph{Biometrics} 58 (1): 21--29.

\leavevmode\hypertarget{ref-Garcia_2017}{}%
Garcia, Erika, Sally Picciotto, Sadie Costello, Patrick T Bradshaw, and
Ellen A Eisen. 2017. {``Assessment of the Healthy Worker Survivor Effect
in Cancer Studies of the United Autoworkers-General Motors Cohort.''}
\emph{Occupational and Environmental Medicine} 74 (4): 294--300.

\leavevmode\hypertarget{ref-Garcia_2018}{}%
Garcia, Erika, Sally Picciotto, Andreas M Neophytou, Patrick T Bradshaw,
John R Balmes, and Ellen A Eisen. 2018. {``Lung Cancer Mortality and
Exposure to Synthetic Metalworking Fluid and Biocides: Controlling for
the Healthy Worker Survivor Effect.''} \emph{Occupational and
Environmental Medicine} 75 (10): 730--35.

\leavevmode\hypertarget{ref-Halliday_2014}{}%
Halliday, Timothy J. 2014. {``Unemployment and Mortality: Evidence from
the PSID.''} \emph{Soc Sci Med} 113 (July): 15--22.
\url{https://doi.org/10.1016/j.socscimed.2014.04.038}.

\leavevmode\hypertarget{ref-Harbison_1950}{}%
Harbison, Frederick H. 1950. {``The General Motors-United Auto Workers
Agreement of 1950.''} \emph{Journal of Political Economy} 58 (5):
397--411.

\leavevmode\hypertarget{ref-Hafner_1987}{}%
Häfner, H. 1987. {``Unemployment and Health.''} \emph{Dtsch Med
Wochenschr} 112 (37): 1428--32.
\url{https://doi.org/10.1055/s-2008-1068265}.

\leavevmode\hypertarget{ref-Hidajat_2020}{}%
Hidajat, Mira, Damien Martin McElvenny, Peter Ritchie, Andrew Darnton,
William Mueller, Raymond M Agius, John W Cherrie, and Frank de Vocht.
2020. {``Lifetime Cumulative Exposure to Rubber Dust, Fumes and
n-Nitrosamines and Non-Cancer Mortality: A 49-Year Follow-up of UK
Rubber Factory Workers.''} \emph{Occup Environ Med} 77 (5): 316--23.
\url{https://doi.org/10.1136/oemed-2019-106269}.

\leavevmode\hypertarget{ref-IARC_1973}{}%
IARC. 1973. \emph{IARC Monographs on the Evaluation of Carcinogenic Risk
of the Chemical to Man: Certain Polycyclic Aromatic Hydrocarbons and
Heterocyclic Compounds}. Vol. 3. World Health Organization International
Agency for Research on Cancer.

\leavevmode\hypertarget{ref-IARC_1987}{}%
---------. 1987. \emph{IARC Monographs on the Evaluation of the
Carcinogenic Risk of Chemicals to Humans. Overall Evaluations of
Carcinogenicity: An Updating of IARC Monographs, Volumes 1 to 42}.
\emph{World Health Organization}. Vol. 7. World Health Organization
International Agency for Research on Cancer.

\leavevmode\hypertarget{ref-Izano_2017_thesis}{}%
Izano, Monika A. 2017. {``Estimating Causal Effects of Occupational
Exposures.''} PhD thesis, University of California, Berkeley; University
of California, Berkeley.

\leavevmode\hypertarget{ref-Izano_2019}{}%
Izano, Monika A, Oleg A Sofrygin, Sally Picciotto, Patrick T Bradshaw,
and Ellen A Eisen. 2019. {``Metalworking Fluids and Colon Cancer Risk:
Longitudinal Targeted Minimum Loss-Based Estimation.''}
\emph{Environmental Epidemiology} 3 (1): e035.

\leavevmode\hypertarget{ref-Kaplan_1958}{}%
Kaplan, Edward L, and Paul Meier. 1958. {``Nonparametric Estimation from
Incomplete Observations.''} \emph{Journal of the American Statistical
Association} 53 (282): 457--81.

\leavevmode\hypertarget{ref-Laliotis_2018}{}%
Laliotis, Ioannis, and Charitini Stavropoulou. 2018. {``Crises and
Mortality: Does the Level of Unemployment Matter?''} \emph{Soc Sci Med}
214 (October): 99--109.
\url{https://doi.org/10.1016/j.socscimed.2018.08.016}.

\leavevmode\hypertarget{ref-Mirer_2003}{}%
Mirer, Franklin. 2003. {``Updated Epidemiology of Workers Exposed to
Metalworking Fluids Provides Sufficient Evidence for Carcinogenicity.''}
\emph{Applied Occupational and Environmental Hygiene} 18 (11): 902--12.

\leavevmode\hypertarget{ref-Mirer_2010}{}%
Mirer, Franklin E. 2010. {``New Evidence on the Health Hazards and
Control of Metalworking Fluids Since Completion of the OSHA Advisory
Committee Report.''} \emph{American Journal of Industrial Medicine} 53
(8): 792--801.

\leavevmode\hypertarget{ref-Naimi_2013}{}%
Naimi, Ashley I, Stephen R Cole, Michael G Hudgens, M Alan Brookhart,
and David B Richardson. 2013. {``Assessing the Component Associations of
the Healthy Worker Survivor Bias: Occupational Asbestos Exposure and
Lung Cancer Mortality.''} \emph{Annals of Epidemiology} 23 (6): 334--41.

\leavevmode\hypertarget{ref-Pearl_1995}{}%
Pearl, Judea. 1995. {``Causal Diagrams for Empirical Research.''}
\emph{Biometrika} 82 (4): 669--88.

\leavevmode\hypertarget{ref-Xie_2005}{}%
Xie, Jun, and Chaofeng Liu. 2005. {``Adjusted Kaplan--Meier Estimator
and Log-Rank Test with Inverse Probability of Treatment Weighting for
Survival Data.''} \emph{Statistics in Medicine} 24 (20): 3089--3110.

\end{CSLReferences}

\end{document}