### 2021-12-11 reports/05-discussion.Rmd

---
output:
  pdf_document:
    includes:
      in_header: "~/headRs/StatHead.sty"
fontsize: 11pt
bibliography: "../../causal.bib"
# csl: "`r path.expand('~/HeadRs/csl/ama.csl')`"
---

# Discussion

- FU in GM is much longer than 20 years
- Exclusion restriction on $H$ so that its parents do not include $A$ is not necessary for identification in full data
- Exclusion restriction on $D$ so that its parents do not include $H$ not needed if $H$ is observed and conditioned upon
- Not clear if these were for convenience or what
- What if we increase FU? Lags?
- Greater fineness in exposure definition? Different interventions?
- No way to verify to what extent the simulation informs the real data generating system
- Model specification, double robustness