### 2021-12-11 reports/resources/results.tex

% Created by tikzDevice version 0.12.3.1 on 2021-12-10 17:23:19
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,542.02);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (325.21,542.02);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,446.55) rectangle (306.77,536.52);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,446.55) rectangle (306.77,536.52);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,465.69) --
	(306.77,465.69);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,492.39) --
	(306.77,492.39);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,519.09) --
	(306.77,519.09);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,446.55) --
	( 63.65,536.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,446.55) --
	(129.54,536.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,446.55) --
	(195.42,536.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,446.55) --
	(261.31,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,452.35) --
	(306.77,452.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,479.04) --
	(306.77,479.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,505.74) --
	(306.77,505.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,532.44) --
	(306.77,532.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,446.55) --
	( 96.59,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,446.55) --
	(162.48,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,446.55) --
	(228.36,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,446.55) --
	(294.25,536.52);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,532.44) --
	( 70.24,532.44) --
	( 70.24,532.44) --
	( 83.42,532.44) --
	( 83.42,531.30) --
	( 96.59,531.30) --
	( 96.59,529.92) --
	(109.77,529.92) --
	(109.77,528.40) --
	(122.95,528.40) --
	(122.95,526.64) --
	(136.12,526.64) --
	(136.12,525.13) --
	(149.30,525.13) --
	(149.30,523.36) --
	(162.48,523.36) --
	(162.48,521.51) --
	(175.66,521.51) --
	(175.66,519.69) --
	(188.83,519.69) --
	(188.83,518.12) --
	(202.01,518.12) --
	(202.01,516.55) --
	(215.19,516.55) --
	(215.19,514.77) --
	(228.36,514.77) --
	(228.36,513.09) --
	(241.54,513.09) --
	(241.54,511.55) --
	(254.72,511.55) --
	(254.72,509.95) --
	(267.89,509.95) --
	(267.89,508.39) --
	(281.07,508.39) --
	(281.07,506.83) --
	(294.25,506.83) --
	(294.25,501.31);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,532.44) --
	( 70.24,532.44) --
	( 70.24,532.44) --
	( 83.42,532.44) --
	( 83.42,530.68) --
	( 96.59,530.68) --
	( 96.59,528.65) --
	(109.77,528.65) --
	(109.77,526.44) --
	(122.95,526.44) --
	(122.95,524.39) --
	(136.12,524.39) --
	(136.12,522.13) --
	(149.30,522.13) --
	(149.30,519.74) --
	(162.48,519.74) --
	(162.48,517.64) --
	(175.66,517.64) --
	(175.66,515.55) --
	(188.83,515.55) --
	(188.83,513.42) --
	(202.01,513.42) --
	(202.01,511.30) --
	(215.19,511.30) --
	(215.19,509.43) --
	(228.36,509.43) --
	(228.36,507.50) --
	(241.54,507.50) --
	(241.54,505.67) --
	(254.72,505.67) --
	(254.72,503.67) --
	(267.89,503.67) --
	(267.89,501.87) --
	(281.07,501.87) --
	(281.07,500.13) --
	(294.25,500.13) --
	(294.25,493.53);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,532.30) --
	( 70.24,532.30) --
	( 70.24,532.07) --
	( 83.42,532.07) --
	( 83.42,531.73) --
	( 96.59,531.73) --
	( 96.59,531.32) --
	(109.77,531.32) --
	(109.77,530.81) --
	(122.95,530.81) --
	(122.95,530.21) --
	(136.12,530.21) --
	(136.12,529.54) --
	(149.30,529.54) --
	(149.30,528.76) --
	(162.48,528.76) --
	(162.48,527.90) --
	(175.66,527.90) --
	(175.66,526.95) --
	(188.83,526.95) --
	(188.83,525.91) --
	(202.01,525.91) --
	(202.01,524.81) --
	(215.19,524.81) --
	(215.19,523.61) --
	(228.36,523.61) --
	(228.36,522.37) --
	(241.54,522.37) --
	(241.54,521.03) --
	(254.72,521.03) --
	(254.72,519.61) --
	(267.89,519.61) --
	(267.89,518.13) --
	(281.07,518.13) --
	(281.07,516.58) --
	(294.25,516.58) --
	(294.25,514.97);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,532.24) --
	( 70.24,532.24) --
	( 70.24,531.95) --
	( 83.42,531.95) --
	( 83.42,531.52) --
	( 96.59,531.52) --
	( 96.59,530.99) --
	(109.77,530.99) --
	(109.77,530.36) --
	(122.95,530.36) --
	(122.95,529.62) --
	(136.12,529.62) --
	(136.12,528.78) --
	(149.30,528.78) --
	(149.30,527.83) --
	(162.48,527.83) --
	(162.48,526.76) --
	(175.66,526.76) --
	(175.66,525.59) --
	(188.83,525.59) --
	(188.83,524.34) --
	(202.01,524.34) --
	(202.01,523.00) --
	(215.19,523.00) --
	(215.19,521.57) --
	(228.36,521.57) --
	(228.36,520.03) --
	(241.54,520.03) --
	(241.54,518.38) --
	(254.72,518.38) --
	(254.72,516.64) --
	(267.89,516.64) --
	(267.89,514.79) --
	(281.07,514.79) --
	(281.07,512.92) --
	(294.25,512.92) --
	(294.25,511.00);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,531.51) --
	( 70.24,531.51) --
	( 70.24,530.38) --
	( 83.42,530.38) --
	( 83.42,529.04) --
	( 96.59,529.04) --
	( 96.59,527.67) --
	(109.77,527.67) --
	(109.77,526.22) --
	(122.95,526.22) --
	(122.95,524.74) --
	(136.12,524.74) --
	(136.12,523.26) --
	(149.30,523.26) --
	(149.30,521.73) --
	(162.48,521.73) --
	(162.48,520.19) --
	(175.66,520.19) --
	(175.66,518.63) --
	(188.83,518.63) --
	(188.83,517.08) --
	(202.01,517.08) --
	(202.01,515.54) --
	(215.19,515.54) --
	(215.19,514.00) --
	(228.36,514.00) --
	(228.36,512.49) --
	(241.54,512.49) --
	(241.54,510.97) --
	(254.72,510.97) --
	(254.72,509.45) --
	(267.89,509.45) --
	(267.89,507.96) --
	(281.07,507.96) --
	(281.07,506.46) --
	(294.25,506.46) --
	(294.25,504.92);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,531.12) --
	( 70.24,531.12) --
	( 70.24,529.67) --
	( 83.42,529.67) --
	( 83.42,527.97) --
	( 96.59,527.97) --
	( 96.59,526.22) --
	(109.77,526.22) --
	(109.77,524.43) --
	(122.95,524.43) --
	(122.95,522.60) --
	(136.12,522.60) --
	(136.12,520.76) --
	(149.30,520.76) --
	(149.30,518.89) --
	(162.48,518.89) --
	(162.48,517.00) --
	(175.66,517.00) --
	(175.66,515.09) --
	(188.83,515.09) --
	(188.83,513.22) --
	(202.01,513.22) --
	(202.01,511.38) --
	(215.19,511.38) --
	(215.19,509.55) --
	(228.36,509.55) --
	(228.36,507.70) --
	(241.54,507.70) --
	(241.54,505.84) --
	(254.72,505.84) --
	(254.72,504.00) --
	(267.89,504.00) --
	(267.89,502.15) --
	(281.07,502.15) --
	(281.07,500.38) --
	(294.25,500.38) --
	(294.25,498.55);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,446.55) rectangle (306.77,536.52);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,351.08) rectangle (306.77,441.05);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,351.08) rectangle (306.77,441.05);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,370.22) --
	(306.77,370.22);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,396.92) --
	(306.77,396.92);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,423.62) --
	(306.77,423.62);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,351.08) --
	( 63.65,441.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,351.08) --
	(129.54,441.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,351.08) --
	(195.42,441.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,351.08) --
	(261.31,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,356.87) --
	(306.77,356.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,383.57) --
	(306.77,383.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,410.27) --
	(306.77,410.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,436.96) --
	(306.77,436.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,351.08) --
	( 96.59,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,351.08) --
	(162.48,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,351.08) --
	(228.36,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,351.08) --
	(294.25,441.05);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.96) --
	( 70.24,436.96) --
	( 70.24,436.96) --
	( 83.42,436.96) --
	( 83.42,435.74) --
	( 96.59,435.74) --
	( 96.59,434.43) --
	(109.77,434.43) --
	(109.77,432.98) --
	(122.95,432.98) --
	(122.95,431.41) --
	(136.12,431.41) --
	(136.12,429.75) --
	(149.30,429.75) --
	(149.30,427.98) --
	(162.48,427.98) --
	(162.48,426.30) --
	(175.66,426.30) --
	(175.66,424.60) --
	(188.83,424.60) --
	(188.83,422.98) --
	(202.01,422.98) --
	(202.01,421.24) --
	(215.19,421.24) --
	(215.19,419.76) --
	(228.36,419.76) --
	(228.36,418.14) --
	(241.54,418.14) --
	(241.54,416.55) --
	(254.72,416.55) --
	(254.72,414.90) --
	(267.89,414.90) --
	(267.89,413.35) --
	(281.07,413.35) --
	(281.07,411.82) --
	(294.25,411.82) --
	(294.25,406.04);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.96) --
	( 70.24,436.96) --
	( 70.24,436.96) --
	( 83.42,436.96) --
	( 83.42,435.30) --
	( 96.59,435.30) --
	( 96.59,433.15) --
	(109.77,433.15) --
	(109.77,430.80) --
	(122.95,430.80) --
	(122.95,428.76) --
	(136.12,428.76) --
	(136.12,426.39) --
	(149.30,426.39) --
	(149.30,424.20) --
	(162.48,424.20) --
	(162.48,422.13) --
	(175.66,422.13) --
	(175.66,420.01) --
	(188.83,420.01) --
	(188.83,417.89) --
	(202.01,417.89) --
	(202.01,415.73) --
	(215.19,415.73) --
	(215.19,413.87) --
	(228.36,413.87) --
	(228.36,412.02) --
	(241.54,412.02) --
	(241.54,410.11) --
	(254.72,410.11) --
	(254.72,408.09) --
	(267.89,408.09) --
	(267.89,406.26) --
	(281.07,406.26) --
	(281.07,404.36) --
	(294.25,404.36) --
	(294.25,397.39);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.82) --
	( 70.24,436.82) --
	( 70.24,436.59) --
	( 83.42,436.59) --
	( 83.42,436.27) --
	( 96.59,436.27) --
	( 96.59,435.85) --
	(109.77,435.85) --
	(109.77,435.33) --
	(122.95,435.33) --
	(122.95,434.71) --
	(136.12,434.71) --
	(136.12,434.01) --
	(149.30,434.01) --
	(149.30,433.22) --
	(162.48,433.22) --
	(162.48,432.33) --
	(175.66,432.33) --
	(175.66,431.35) --
	(188.83,431.35) --
	(188.83,430.30) --
	(202.01,430.30) --
	(202.01,429.17) --
	(215.19,429.17) --
	(215.19,427.95) --
	(228.36,427.95) --
	(228.36,426.66) --
	(241.54,426.66) --
	(241.54,425.30) --
	(254.72,425.30) --
	(254.72,423.86) --
	(267.89,423.86) --
	(267.89,422.34) --
	(281.07,422.34) --
	(281.07,420.76) --
	(294.25,420.76) --
	(294.25,419.13);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.79) --
	( 70.24,436.79) --
	( 70.24,436.50) --
	( 83.42,436.50) --
	( 83.42,436.07) --
	( 96.59,436.07) --
	( 96.59,435.53) --
	(109.77,435.53) --
	(109.77,434.89) --
	(122.95,434.89) --
	(122.95,434.16) --
	(136.12,434.16) --
	(136.12,433.29) --
	(149.30,433.29) --
	(149.30,432.32) --
	(162.48,432.32) --
	(162.48,431.20) --
	(175.66,431.20) --
	(175.66,429.97) --
	(188.83,429.97) --
	(188.83,428.71) --
	(202.01,428.71) --
	(202.01,427.33) --
	(215.19,427.33) --
	(215.19,425.87) --
	(228.36,425.87) --
	(228.36,424.24) --
	(241.54,424.24) --
	(241.54,422.48) --
	(254.72,422.48) --
	(254.72,420.67) --
	(267.89,420.67) --
	(267.89,418.79) --
	(281.07,418.79) --
	(281.07,416.82) --
	(294.25,416.82) --
	(294.25,414.88);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.03) --
	( 70.24,436.03) --
	( 70.24,434.88) --
	( 83.42,434.88) --
	( 83.42,433.58) --
	( 96.59,433.58) --
	( 96.59,432.19) --
	(109.77,432.19) --
	(109.77,430.73) --
	(122.95,430.73) --
	(122.95,429.19) --
	(136.12,429.19) --
	(136.12,427.66) --
	(149.30,427.66) --
	(149.30,426.11) --
	(162.48,426.11) --
	(162.48,424.54) --
	(175.66,424.54) --
	(175.66,422.95) --
	(188.83,422.95) --
	(188.83,421.38) --
	(202.01,421.38) --
	(202.01,419.81) --
	(215.19,419.81) --
	(215.19,418.25) --
	(228.36,418.25) --
	(228.36,416.70) --
	(241.54,416.70) --
	(241.54,415.15) --
	(254.72,415.15) --
	(254.72,413.61) --
	(267.89,413.61) --
	(267.89,412.08) --
	(281.07,412.08) --
	(281.07,410.56) --
	(294.25,410.56) --
	(294.25,409.00);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,435.82) --
	( 70.24,435.82) --
	( 70.24,434.36) --
	( 83.42,434.36) --
	( 83.42,432.67) --
	( 96.59,432.67) --
	( 96.59,430.87) --
	(109.77,430.87) --
	(109.77,429.08) --
	(122.95,429.08) --
	(122.95,427.28) --
	(136.12,427.28) --
	(136.12,425.38) --
	(149.30,425.38) --
	(149.30,423.49) --
	(162.48,423.49) --
	(162.48,421.50) --
	(175.66,421.50) --
	(175.66,419.53) --
	(188.83,419.53) --
	(188.83,417.65) --
	(202.01,417.65) --
	(202.01,415.76) --
	(215.19,415.76) --
	(215.19,413.91) --
	(228.36,413.91) --
	(228.36,411.95) --
	(241.54,411.95) --
	(241.54,409.99) --
	(254.72,409.99) --
	(254.72,408.06) --
	(267.89,408.06) --
	(267.89,406.19) --
	(281.07,406.19) --
	(281.07,404.32) --
	(294.25,404.32) --
	(294.25,402.48);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,351.08) rectangle (306.77,441.05);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,255.61) rectangle (306.77,345.58);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,255.61) rectangle (306.77,345.58);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,274.75) --
	(306.77,274.75);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,301.45) --
	(306.77,301.45);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,328.14) --
	(306.77,328.14);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,255.61) --
	( 63.65,345.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,255.61) --
	(129.54,345.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,255.61) --
	(195.42,345.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,255.61) --
	(261.31,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,261.40) --
	(306.77,261.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,288.10) --
	(306.77,288.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,314.80) --
	(306.77,314.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,341.49) --
	(306.77,341.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,255.61) --
	( 96.59,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,255.61) --
	(162.48,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,255.61) --
	(228.36,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,255.61) --
	(294.25,345.58);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,341.49) --
	( 70.24,341.49) --
	( 70.24,341.49) --
	( 83.42,341.49) --
	( 83.42,340.45) --
	( 96.59,340.45) --
	( 96.59,339.00) --
	(109.77,339.00) --
	(109.77,337.47) --
	(122.95,337.47) --
	(122.95,335.92) --
	(136.12,335.92) --
	(136.12,334.17) --
	(149.30,334.17) --
	(149.30,332.43) --
	(162.48,332.43) --
	(162.48,330.88) --
	(175.66,330.88) --
	(175.66,329.12) --
	(188.83,329.12) --
	(188.83,327.55) --
	(202.01,327.55) --
	(202.01,325.87) --
	(215.19,325.87) --
	(215.19,324.26) --
	(228.36,324.26) --
	(228.36,322.63) --
	(241.54,322.63) --
	(241.54,321.13) --
	(254.72,321.13) --
	(254.72,319.53) --
	(267.89,319.53) --
	(267.89,317.83) --
	(281.07,317.83) --
	(281.07,316.18) --
	(294.25,316.18) --
	(294.25,310.63);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,341.49) --
	( 70.24,341.49) --
	( 70.24,341.49) --
	( 83.42,341.49) --
	( 83.42,339.77) --
	( 96.59,339.77) --
	( 96.59,337.60) --
	(109.77,337.60) --
	(109.77,335.22) --
	(122.95,335.22) --
	(122.95,333.15) --
	(136.12,333.15) --
	(136.12,330.77) --
	(149.30,330.77) --
	(149.30,328.58) --
	(162.48,328.58) --
	(162.48,326.38) --
	(175.66,326.38) --
	(175.66,324.09) --
	(188.83,324.09) --
	(188.83,321.88) --
	(202.01,321.88) --
	(202.01,319.52) --
	(215.19,319.52) --
	(215.19,317.54) --
	(228.36,317.54) --
	(228.36,315.57) --
	(241.54,315.57) --
	(241.54,313.64) --
	(254.72,313.64) --
	(254.72,311.82) --
	(267.89,311.82) --
	(267.89,309.92) --
	(281.07,309.92) --
	(281.07,308.01) --
	(294.25,308.01) --
	(294.25,300.71);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,341.35) --
	( 70.24,341.35) --
	( 70.24,341.12) --
	( 83.42,341.12) --
	( 83.42,340.80) --
	( 96.59,340.80) --
	( 96.59,340.38) --
	(109.77,340.38) --
	(109.77,339.87) --
	(122.95,339.87) --
	(122.95,339.28) --
	(136.12,339.28) --
	(136.12,338.61) --
	(149.30,338.61) --
	(149.30,337.84) --
	(162.48,337.84) --
	(162.48,336.98) --
	(175.66,336.98) --
	(175.66,336.03) --
	(188.83,336.03) --
	(188.83,335.00) --
	(202.01,335.00) --
	(202.01,333.89) --
	(215.19,333.89) --
	(215.19,332.70) --
	(228.36,332.70) --
	(228.36,331.45) --
	(241.54,331.45) --
	(241.54,330.12) --
	(254.72,330.12) --
	(254.72,328.70) --
	(267.89,328.70) --
	(267.89,327.22) --
	(281.07,327.22) --
	(281.07,325.67) --
	(294.25,325.67) --
	(294.25,324.06);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,341.30) --
	( 70.24,341.30) --
	( 70.24,341.00) --
	( 83.42,341.00) --
	( 83.42,340.57) --
	( 96.59,340.57) --
	( 96.59,340.03) --
	(109.77,340.03) --
	(109.77,339.38) --
	(122.95,339.38) --
	(122.95,338.61) --
	(136.12,338.61) --
	(136.12,337.76) --
	(149.30,337.76) --
	(149.30,336.78) --
	(162.48,336.78) --
	(162.48,335.71) --
	(175.66,335.71) --
	(175.66,334.51) --
	(188.83,334.51) --
	(188.83,333.23) --
	(202.01,333.23) --
	(202.01,331.82) --
	(215.19,331.82) --
	(215.19,330.32) --
	(228.36,330.32) --
	(228.36,328.73) --
	(241.54,328.73) --
	(241.54,327.07) --
	(254.72,327.07) --
	(254.72,325.34) --
	(267.89,325.34) --
	(267.89,323.47) --
	(281.07,323.47) --
	(281.07,321.56) --
	(294.25,321.56) --
	(294.25,319.55);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,340.55) --
	( 70.24,340.55) --
	( 70.24,339.42) --
	( 83.42,339.42) --
	( 83.42,338.11) --
	( 96.59,338.11) --
	( 96.59,336.75) --
	(109.77,336.75) --
	(109.77,335.30) --
	(122.95,335.30) --
	(122.95,333.82) --
	(136.12,333.82) --
	(136.12,332.35) --
	(149.30,332.35) --
	(149.30,330.83) --
	(162.48,330.83) --
	(162.48,329.30) --
	(175.66,329.30) --
	(175.66,327.75) --
	(188.83,327.75) --
	(188.83,326.20) --
	(202.01,326.20) --
	(202.01,324.66) --
	(215.19,324.66) --
	(215.19,323.12) --
	(228.36,323.12) --
	(228.36,321.61) --
	(241.54,321.61) --
	(241.54,320.09) --
	(254.72,320.09) --
	(254.72,318.58) --
	(267.89,318.58) --
	(267.89,317.08) --
	(281.07,317.08) --
	(281.07,315.59) --
	(294.25,315.59) --
	(294.25,314.04);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,340.23) --
	( 70.24,340.23) --
	( 70.24,338.73) --
	( 83.42,338.73) --
	( 83.42,337.01) --
	( 96.59,337.01) --
	( 96.59,335.22) --
	(109.77,335.22) --
	(109.77,333.37) --
	(122.95,333.37) --
	(122.95,331.48) --
	(136.12,331.48) --
	(136.12,329.61) --
	(149.30,329.61) --
	(149.30,327.68) --
	(162.48,327.68) --
	(162.48,325.78) --
	(175.66,325.78) --
	(175.66,323.83) --
	(188.83,323.83) --
	(188.83,321.93) --
	(202.01,321.93) --
	(202.01,319.98) --
	(215.19,319.98) --
	(215.19,318.06) --
	(228.36,318.06) --
	(228.36,316.16) --
	(241.54,316.16) --
	(241.54,314.28) --
	(254.72,314.28) --
	(254.72,312.45) --
	(267.89,312.45) --
	(267.89,310.59) --
	(281.07,310.59) --
	(281.07,308.77) --
	(294.25,308.77) --
	(294.25,306.87);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,255.61) rectangle (306.77,345.58);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,160.14) rectangle (306.77,250.11);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,160.14) rectangle (306.77,250.11);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,179.28) --
	(306.77,179.28);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,205.98) --
	(306.77,205.98);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,232.67) --
	(306.77,232.67);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,160.14) --
	( 63.65,250.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,160.14) --
	(129.54,250.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,160.14) --
	(195.42,250.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,160.14) --
	(261.31,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,165.93) --
	(306.77,165.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,192.63) --
	(306.77,192.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,219.32) --
	(306.77,219.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,246.02) --
	(306.77,246.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,160.14) --
	( 96.59,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,160.14) --
	(162.48,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,160.14) --
	(228.36,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,160.14) --
	(294.25,250.11);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,246.02) --
	( 70.24,246.02) --
	( 70.24,246.02) --
	( 83.42,246.02) --
	( 83.42,243.83) --
	( 96.59,243.83) --
	( 96.59,240.92) --
	(109.77,240.92) --
	(109.77,237.63) --
	(122.95,237.63) --
	(122.95,233.74) --
	(136.12,233.74) --
	(136.12,229.93) --
	(149.30,229.93) --
	(149.30,226.16) --
	(162.48,226.16) --
	(162.48,222.32) --
	(175.66,222.32) --
	(175.66,218.61) --
	(188.83,218.61) --
	(188.83,214.89) --
	(202.01,214.89) --
	(202.01,211.14) --
	(215.19,211.14) --
	(215.19,207.61) --
	(228.36,207.61) --
	(228.36,204.27) --
	(241.54,204.27) --
	(241.54,200.76) --
	(254.72,200.76) --
	(254.72,197.48) --
	(267.89,197.48) --
	(267.89,194.34) --
	(281.07,194.34) --
	(281.07,191.11) --
	(294.25,191.11) --
	(294.25,180.15);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,246.02) --
	( 70.24,246.02) --
	( 70.24,246.02) --
	( 83.42,246.02) --
	( 83.42,243.24) --
	( 96.59,243.24) --
	( 96.59,239.25) --
	(109.77,239.25) --
	(109.77,234.65) --
	(122.95,234.65) --
	(122.95,229.74) --
	(136.12,229.74) --
	(136.12,224.94) --
	(149.30,224.94) --
	(149.30,220.15) --
	(162.48,220.15) --
	(162.48,215.44) --
	(175.66,215.44) --
	(175.66,210.69) --
	(188.83,210.69) --
	(188.83,206.27) --
	(202.01,206.27) --
	(202.01,202.11) --
	(215.19,202.11) --
	(215.19,198.14) --
	(228.36,198.14) --
	(228.36,194.18) --
	(241.54,194.18) --
	(241.54,190.34) --
	(254.72,190.34) --
	(254.72,186.73) --
	(267.89,186.73) --
	(267.89,183.18) --
	(281.07,183.18) --
	(281.07,179.87) --
	(294.25,179.87) --
	(294.25,167.81);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,245.72) --
	( 70.24,245.72) --
	( 70.24,245.22) --
	( 83.42,245.22) --
	( 83.42,244.49) --
	( 96.59,244.49) --
	( 96.59,243.55) --
	(109.77,243.55) --
	(109.77,242.36) --
	(122.95,242.36) --
	(122.95,240.98) --
	(136.12,240.98) --
	(136.12,239.40) --
	(149.30,239.40) --
	(149.30,237.63) --
	(162.48,237.63) --
	(162.48,235.66) --
	(175.66,235.66) --
	(175.66,233.45) --
	(188.83,233.45) --
	(188.83,231.09) --
	(202.01,231.09) --
	(202.01,228.56) --
	(215.19,228.56) --
	(215.19,225.87) --
	(228.36,225.87) --
	(228.36,223.01) --
	(241.54,223.01) --
	(241.54,219.97) --
	(254.72,219.97) --
	(254.72,216.86) --
	(267.89,216.86) --
	(267.89,213.57) --
	(281.07,213.57) --
	(281.07,210.22) --
	(294.25,210.22) --
	(294.25,206.69);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,245.59) --
	( 70.24,245.59) --
	( 70.24,244.91) --
	( 83.42,244.91) --
	( 83.42,243.97) --
	( 96.59,243.97) --
	( 96.59,242.78) --
	(109.77,242.78) --
	(109.77,241.36) --
	(122.95,241.36) --
	(122.95,239.66) --
	(136.12,239.66) --
	(136.12,237.77) --
	(149.30,237.77) --
	(149.30,235.63) --
	(162.48,235.63) --
	(162.48,233.29) --
	(175.66,233.29) --
	(175.66,230.69) --
	(188.83,230.69) --
	(188.83,227.85) --
	(202.01,227.85) --
	(202.01,224.84) --
	(215.19,224.84) --
	(215.19,221.67) --
	(228.36,221.67) --
	(228.36,218.22) --
	(241.54,218.22) --
	(241.54,214.58) --
	(254.72,214.58) --
	(254.72,210.76) --
	(267.89,210.76) --
	(267.89,206.94) --
	(281.07,206.94) --
	(281.07,202.93) --
	(294.25,202.93) --
	(294.25,199.00);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,244.05) --
	( 70.24,244.05) --
	( 70.24,241.55) --
	( 83.42,241.55) --
	( 83.42,238.62) --
	( 96.59,238.62) --
	( 96.59,235.52) --
	(109.77,235.52) --
	(109.77,232.18) --
	(122.95,232.18) --
	(122.95,228.78) --
	(136.12,228.78) --
	(136.12,225.36) --
	(149.30,225.36) --
	(149.30,221.94) --
	(162.48,221.94) --
	(162.48,218.49) --
	(175.66,218.49) --
	(175.66,214.98) --
	(188.83,214.98) --
	(188.83,211.54) --
	(202.01,211.54) --
	(202.01,208.13) --
	(215.19,208.13) --
	(215.19,204.78) --
	(228.36,204.78) --
	(228.36,201.47) --
	(241.54,201.47) --
	(241.54,198.16) --
	(254.72,198.16) --
	(254.72,194.99) --
	(267.89,194.99) --
	(267.89,191.84) --
	(281.07,191.84) --
	(281.07,188.80) --
	(294.25,188.80) --
	(294.25,185.60);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,243.15) --
	( 70.24,243.15) --
	( 70.24,239.77) --
	( 83.42,239.77) --
	( 83.42,236.02) --
	( 96.59,236.02) --
	( 96.59,232.11) --
	(109.77,232.11) --
	(109.77,228.13) --
	(122.95,228.13) --
	(122.95,224.01) --
	(136.12,224.01) --
	(136.12,219.94) --
	(149.30,219.94) --
	(149.30,215.85) --
	(162.48,215.85) --
	(162.48,211.80) --
	(175.66,211.80) --
	(175.66,207.73) --
	(188.83,207.73) --
	(188.83,203.66) --
	(202.01,203.66) --
	(202.01,199.70) --
	(215.19,199.70) --
	(215.19,195.82) --
	(228.36,195.82) --
	(228.36,191.90) --
	(241.54,191.90) --
	(241.54,188.04) --
	(254.72,188.04) --
	(254.72,184.24) --
	(267.89,184.24) --
	(267.89,180.67) --
	(281.07,180.67) --
	(281.07,177.12) --
	(294.25,177.12) --
	(294.25,173.64);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,160.14) rectangle (306.77,250.11);
\end{scope}
\begin{scope}
\path[clip] ( 31.37, 64.67) rectangle (306.77,154.64);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37, 64.67) rectangle (306.77,154.64);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37, 83.81) --
	(306.77, 83.81);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,110.50) --
	(306.77,110.50);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,137.20) --
	(306.77,137.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65, 64.67) --
	( 63.65,154.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54, 64.67) --
	(129.54,154.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42, 64.67) --
	(195.42,154.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31, 64.67) --
	(261.31,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 70.46) --
	(306.77, 70.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 97.16) --
	(306.77, 97.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,123.85) --
	(306.77,123.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,150.55) --
	(306.77,150.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 64.67) --
	( 96.59,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 64.67) --
	(162.48,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 64.67) --
	(228.36,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 64.67) --
	(294.25,154.64);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,150.55) --
	( 70.24,150.55) --
	( 70.24,150.55) --
	( 83.42,150.55) --
	( 83.42,147.28) --
	( 96.59,147.28) --
	( 96.59,143.70) --
	(109.77,143.70) --
	(109.77,139.54) --
	(122.95,139.54) --
	(122.95,135.49) --
	(136.12,135.49) --
	(136.12,131.43) --
	(149.30,131.43) --
	(149.30,127.27) --
	(162.48,127.27) --
	(162.48,123.40) --
	(175.66,123.40) --
	(175.66,119.60) --
	(188.83,119.60) --
	(188.83,115.97) --
	(202.01,115.97) --
	(202.01,112.27) --
	(215.19,112.27) --
	(215.19,108.52) --
	(228.36,108.52) --
	(228.36,105.14) --
	(241.54,105.14) --
	(241.54,101.84) --
	(254.72,101.84) --
	(254.72, 98.62) --
	(267.89, 98.62) --
	(267.89, 95.61) --
	(281.07, 95.61) --
	(281.07, 92.76) --
	(294.25, 92.76) --
	(294.25, 81.64);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,150.55) --
	( 70.24,150.55) --
	( 70.24,150.55) --
	( 83.42,150.55) --
	( 83.42,146.40) --
	( 96.59,146.40) --
	( 96.59,140.92) --
	(109.77,140.92) --
	(109.77,135.34) --
	(122.95,135.34) --
	(122.95,130.16) --
	(136.12,130.16) --
	(136.12,124.95) --
	(149.30,124.95) --
	(149.30,120.21) --
	(162.48,120.21) --
	(162.48,115.52) --
	(175.66,115.52) --
	(175.66,110.59) --
	(188.83,110.59) --
	(188.83,106.11) --
	(202.01,106.11) --
	(202.01,102.07) --
	(215.19,102.07) --
	(215.19, 98.13) --
	(228.36, 98.13) --
	(228.36, 94.49) --
	(241.54, 94.49) --
	(241.54, 90.89) --
	(254.72, 90.89) --
	(254.72, 87.28) --
	(267.89, 87.28) --
	(267.89, 83.67) --
	(281.07, 83.67) --
	(281.07, 80.56) --
	(294.25, 80.56) --
	(294.25, 68.76);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,150.18) --
	( 70.24,150.18) --
	( 70.24,149.60) --
	( 83.42,149.60) --
	( 83.42,148.79) --
	( 96.59,148.79) --
	( 96.59,147.79) --
	(109.77,147.79) --
	(109.77,146.60) --
	(122.95,146.60) --
	(122.95,145.20) --
	(136.12,145.20) --
	(136.12,143.61) --
	(149.30,143.61) --
	(149.30,141.83) --
	(162.48,141.83) --
	(162.48,139.86) --
	(175.66,139.86) --
	(175.66,137.70) --
	(188.83,137.70) --
	(188.83,135.37) --
	(202.01,135.37) --
	(202.01,132.84) --
	(215.19,132.84) --
	(215.19,130.16) --
	(228.36,130.16) --
	(228.36,127.31) --
	(241.54,127.31) --
	(241.54,124.33) --
	(254.72,124.33) --
	(254.72,121.21) --
	(267.89,121.21) --
	(267.89,117.94) --
	(281.07,117.94) --
	(281.07,114.58) --
	(294.25,114.58) --
	(294.25,111.11);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,150.07) --
	( 70.24,150.07) --
	( 70.24,149.32) --
	( 83.42,149.32) --
	( 83.42,148.30) --
	( 96.59,148.30) --
	( 96.59,147.04) --
	(109.77,147.04) --
	(109.77,145.50) --
	(122.95,145.50) --
	(122.95,143.83) --
	(136.12,143.83) --
	(136.12,141.94) --
	(149.30,141.94) --
	(149.30,139.74) --
	(162.48,139.74) --
	(162.48,137.40) --
	(175.66,137.40) --
	(175.66,134.81) --
	(188.83,134.81) --
	(188.83,132.03) --
	(202.01,132.03) --
	(202.01,129.02) --
	(215.19,129.02) --
	(215.19,125.87) --
	(228.36,125.87) --
	(228.36,122.58) --
	(241.54,122.58) --
	(241.54,119.15) --
	(254.72,119.15) --
	(254.72,115.52) --
	(267.89,115.52) --
	(267.89,111.71) --
	(281.07,111.71) --
	(281.07,107.78) --
	(294.25,107.78) --
	(294.25,103.68);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,148.08) --
	( 70.24,148.08) --
	( 70.24,145.17) --
	( 83.42,145.17) --
	( 83.42,141.98) --
	( 96.59,141.98) --
	( 96.59,138.67) --
	(109.77,138.67) --
	(109.77,135.32) --
	(122.95,135.32) --
	(122.95,131.90) --
	(136.12,131.90) --
	(136.12,128.47) --
	(149.30,128.47) --
	(149.30,125.04) --
	(162.48,125.04) --
	(162.48,121.61) --
	(175.66,121.61) --
	(175.66,118.18) --
	(188.83,118.18) --
	(188.83,114.80) --
	(202.01,114.80) --
	(202.01,111.42) --
	(215.19,111.42) --
	(215.19,108.10) --
	(228.36,108.10) --
	(228.36,104.80) --
	(241.54,104.80) --
	(241.54,101.58) --
	(254.72,101.58) --
	(254.72, 98.42) --
	(267.89, 98.42) --
	(267.89, 95.30) --
	(281.07, 95.30) --
	(281.07, 92.26) --
	(294.25, 92.26) --
	(294.25, 89.13);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,147.32) --
	( 70.24,147.32) --
	( 70.24,143.59) --
	( 83.42,143.59) --
	( 83.42,139.53) --
	( 96.59,139.53) --
	( 96.59,135.42) --
	(109.77,135.42) --
	(109.77,131.09) --
	(122.95,131.09) --
	(122.95,127.04) --
	(136.12,127.04) --
	(136.12,122.99) --
	(149.30,122.99) --
	(149.30,118.80) --
	(162.48,118.80) --
	(162.48,114.77) --
	(175.66,114.77) --
	(175.66,110.73) --
	(188.83,110.73) --
	(188.83,106.75) --
	(202.01,106.75) --
	(202.01,102.80) --
	(215.19,102.80) --
	(215.19, 98.97) --
	(228.36, 98.97) --
	(228.36, 95.25) --
	(241.54, 95.25) --
	(241.54, 91.63) --
	(254.72, 91.63) --
	(254.72, 88.04) --
	(267.89, 88.04) --
	(267.89, 84.49) --
	(281.07, 84.49) --
	(281.07, 81.02) --
	(294.25, 81.02) --
	(294.25, 77.41);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37, 64.67) rectangle (306.77,154.64);
\end{scope}
\begin{scope}
\path[clip] (306.77,446.55) rectangle (319.71,536.52);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,446.55) rectangle (319.71,536.52);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,491.54) {Scenario 1};
\end{scope}
\begin{scope}
\path[clip] (306.77,351.08) rectangle (319.71,441.05);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,351.08) rectangle (319.71,441.05);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,396.07) {Scenario 2};
\end{scope}
\begin{scope}
\path[clip] (306.77,255.61) rectangle (319.71,345.58);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,255.61) rectangle (319.71,345.58);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,300.60) {Scenario 3};
\end{scope}
\begin{scope}
\path[clip] (306.77,160.14) rectangle (319.71,250.11);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,160.14) rectangle (319.71,250.11);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,205.12) {Scenario 4};
\end{scope}
\begin{scope}
\path[clip] (306.77, 64.67) rectangle (319.71,154.64);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77, 64.67) rectangle (319.71,154.64);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,109.65) {Scenario 5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 61.92) --
	( 96.59, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 61.92) --
	(162.48, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 61.92) --
	(228.36, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 61.92) --
	(294.25, 64.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 96.59, 54.21) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (162.48, 54.21) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (228.36, 54.21) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (294.25, 54.21) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,449.59) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,476.29) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,502.98) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,529.68) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,452.35) --
	( 31.37,452.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,479.04) --
	( 31.37,479.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,505.74) --
	( 31.37,505.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,532.44) --
	( 31.37,532.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,354.12) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,380.82) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,407.51) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,434.21) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,356.87) --
	( 31.37,356.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,383.57) --
	( 31.37,383.57);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,410.27) --
	( 31.37,410.27);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,436.96) --
	( 31.37,436.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,258.65) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,285.34) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,312.04) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,338.74) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,261.40) --
	( 31.37,261.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,288.10) --
	( 31.37,288.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,314.80) --
	( 31.37,314.80);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,341.49) --
	( 31.37,341.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,163.18) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,189.87) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,216.57) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,243.27) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,165.93) --
	( 31.37,165.93);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,192.63) --
	( 31.37,192.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,219.32) --
	( 31.37,219.32);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,246.02) --
	( 31.37,246.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 67.70) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 94.40) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,121.10) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,147.79) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 70.46) --
	( 31.37, 70.46);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 97.16) --
	( 31.37, 97.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,123.85) --
	( 31.37,123.85);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,150.55) --
	( 31.37,150.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (169.07, 43.70) {Time since hire};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,300.60) {Cancer-free survival};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 19.60,  5.50) rectangle (142.67, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 30.60, 11.00) rectangle ( 45.05, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.05, 18.23) -- ( 43.61, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 86.63, 11.00) rectangle (101.09, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.08, 18.23) -- ( 99.64, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 50.55, 15.13) {Rule $a_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (106.59, 15.13) {Rule $a_1$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (153.67,  5.50) rectangle (318.53, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (164.67, 11.00) rectangle (179.12, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.11, 18.23) -- (177.68, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.89, 11.00) rectangle (227.34, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (214.34, 18.23) -- (225.90, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (262.84, 11.00) rectangle (277.29, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (264.28, 18.23) -- (275.85, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (184.62, 15.13) {Truth};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (232.84, 15.13) {WKM};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (282.79, 15.13) {AWKM};
\end{scope}
\end{tikzpicture}

\end{document}