### 2022-03-04 reports/resources/gm_results_1-2.tex

% Created by tikzDevice version 0.12.3.1 on 2022-01-19 23:30:07
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,186.91);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (325.21,186.91);
\end{scope}
\begin{scope}
\path[clip] ( 35.37,125.79) rectangle (306.77,181.41);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37,125.79) rectangle (306.77,181.41);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,136.24) --
	(306.77,136.24);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,153.30) --
	(306.77,153.30);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,170.35) --
	(306.77,170.35);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 67.18,125.79) --
	( 67.18,181.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (132.11,125.79) --
	(132.11,181.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (197.04,125.79) --
	(197.04,181.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.97,125.79) --
	(261.97,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,127.71) --
	(306.77,127.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,144.77) --
	(306.77,144.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,161.82) --
	(306.77,161.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,178.88) --
	(306.77,178.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.65,125.79) --
	( 99.65,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (164.57,125.79) --
	(164.57,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.50,125.79) --
	(229.50,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.43,125.79) --
	(294.43,181.41);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 47.70,178.88) --
	( 60.69,178.88) --
	( 60.69,178.88) --
	( 73.67,178.88) --
	( 73.67,178.88) --
	( 86.66,178.88) --
	( 86.66,177.42) --
	( 99.65,177.42) --
	( 99.65,175.66) --
	(112.63,175.66) --
	(112.63,173.73) --
	(125.62,173.73) --
	(125.62,171.48) --
	(138.60,171.48) --
	(138.60,169.54) --
	(151.59,169.54) --
	(151.59,167.29) --
	(164.57,167.29) --
	(164.57,164.91) --
	(177.56,164.91) --
	(177.56,162.59) --
	(190.55,162.59) --
	(190.55,160.59) --
	(203.53,160.59) --
	(203.53,158.58) --
	(216.52,158.58) --
	(216.52,156.31) --
	(229.50,156.31) --
	(229.50,154.16) --
	(242.49,154.16) --
	(242.49,152.19) --
	(255.47,152.19) --
	(255.47,150.14) --
	(268.46,150.14) --
	(268.46,148.15) --
	(281.44,148.15) --
	(281.44,146.17) --
	(294.43,146.17) --
	(294.43,139.11);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 47.70,178.88) --
	( 60.69,178.88) --
	( 60.69,178.88) --
	( 73.67,178.88) --
	( 73.67,178.88) --
	( 86.66,178.88) --
	( 86.66,176.64) --
	( 99.65,176.64) --
	( 99.65,174.04) --
	(112.63,174.04) --
	(112.63,171.22) --
	(125.62,171.22) --
	(125.62,168.60) --
	(138.60,168.60) --
	(138.60,165.71) --
	(151.59,165.71) --
	(151.59,162.66) --
	(164.57,162.66) --
	(164.57,159.97) --
	(177.56,159.97) --
	(177.56,157.30) --
	(190.55,157.30) --
	(190.55,154.58) --
	(203.53,154.58) --
	(203.53,151.87) --
	(216.52,151.87) --
	(216.52,149.49) --
	(229.50,149.49) --
	(229.50,147.01) --
	(242.49,147.01) --
	(242.49,144.69) --
	(255.47,144.69) --
	(255.47,142.13) --
	(268.46,142.13) --
	(268.46,139.83) --
	(281.44,139.83) --
	(281.44,137.60) --
	(294.43,137.60) --
	(294.43,129.17);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 47.70,178.88) --
	( 60.69,178.88) --
	( 60.69,178.70) --
	( 73.67,178.70) --
	( 73.67,178.41) --
	( 86.66,178.41) --
	( 86.66,177.98) --
	( 99.65,177.98) --
	( 99.65,177.45) --
	(112.63,177.45) --
	(112.63,176.80) --
	(125.62,176.80) --
	(125.62,176.04) --
	(138.60,176.04) --
	(138.60,175.17) --
	(151.59,175.17) --
	(151.59,174.19) --
	(164.57,174.19) --
	(164.57,173.08) --
	(177.56,173.08) --
	(177.56,171.86) --
	(190.55,171.86) --
	(190.55,170.54) --
	(203.53,170.54) --
	(203.53,169.13) --
	(216.52,169.13) --
	(216.52,167.61) --
	(229.50,167.61) --
	(229.50,166.02) --
	(242.49,166.02) --
	(242.49,164.31) --
	(255.47,164.31) --
	(255.47,162.50) --
	(268.46,162.50) --
	(268.46,160.61) --
	(281.44,160.61) --
	(281.44,158.62) --
	(294.43,158.62) --
	(294.43,156.56);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 47.70,178.88) --
	( 60.69,178.88) --
	( 60.69,178.63) --
	( 73.67,178.63) --
	( 73.67,178.25) --
	( 86.66,178.25) --
	( 86.66,177.71) --
	( 99.65,177.71) --
	( 99.65,177.04) --
	(112.63,177.04) --
	(112.63,176.23) --
	(125.62,176.23) --
	(125.62,175.28) --
	(138.60,175.28) --
	(138.60,174.20) --
	(151.59,174.20) --
	(151.59,172.99) --
	(164.57,172.99) --
	(164.57,171.63) --
	(177.56,171.63) --
	(177.56,170.13) --
	(190.55,170.13) --
	(190.55,168.53) --
	(203.53,168.53) --
	(203.53,166.82) --
	(216.52,166.82) --
	(216.52,165.00) --
	(229.50,165.00) --
	(229.50,163.02) --
	(242.49,163.02) --
	(242.49,160.91) --
	(255.47,160.91) --
	(255.47,158.70) --
	(268.46,158.70) --
	(268.46,156.33) --
	(281.44,156.33) --
	(281.44,153.95) --
	(294.43,153.95) --
	(294.43,151.48);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 47.70,178.88) --
	( 60.69,178.88) --
	( 60.69,177.69) --
	( 73.67,177.69) --
	( 73.67,176.25) --
	( 86.66,176.25) --
	( 86.66,174.54) --
	( 99.65,174.54) --
	( 99.65,172.79) --
	(112.63,172.79) --
	(112.63,170.93) --
	(125.62,170.93) --
	(125.62,169.05) --
	(138.60,169.05) --
	(138.60,167.15) --
	(151.59,167.15) --
	(151.59,165.20) --
	(164.57,165.20) --
	(164.57,163.23) --
	(177.56,163.23) --
	(177.56,161.24) --
	(190.55,161.24) --
	(190.55,159.26) --
	(203.53,159.26) --
	(203.53,157.29) --
	(216.52,157.29) --
	(216.52,155.32) --
	(229.50,155.32) --
	(229.50,153.40) --
	(242.49,153.40) --
	(242.49,151.46) --
	(255.47,151.46) --
	(255.47,149.52) --
	(268.46,149.52) --
	(268.46,147.60) --
	(281.44,147.60) --
	(281.44,145.70) --
	(294.43,145.70) --
	(294.43,143.72);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 47.70,178.88) --
	( 60.69,178.88) --
	( 60.69,177.20) --
	( 73.67,177.20) --
	( 73.67,175.34) --
	( 86.66,175.34) --
	( 86.66,173.17) --
	( 99.65,173.17) --
	( 99.65,170.94) --
	(112.63,170.94) --
	(112.63,168.65) --
	(125.62,168.65) --
	(125.62,166.31) --
	(138.60,166.31) --
	(138.60,163.95) --
	(151.59,163.95) --
	(151.59,161.57) --
	(164.57,161.57) --
	(164.57,159.16) --
	(177.56,159.16) --
	(177.56,156.72) --
	(190.55,156.72) --
	(190.55,154.33) --
	(203.53,154.33) --
	(203.53,151.98) --
	(216.52,151.98) --
	(216.52,149.64) --
	(229.50,149.64) --
	(229.50,147.27) --
	(242.49,147.27) --
	(242.49,144.90) --
	(255.47,144.90) --
	(255.47,142.55) --
	(268.46,142.55) --
	(268.46,140.18) --
	(281.44,140.18) --
	(281.44,137.92) --
	(294.43,137.92) --
	(294.43,135.58);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37,125.79) rectangle (306.77,181.41);
\end{scope}
\begin{scope}
\path[clip] ( 35.37, 64.67) rectangle (306.77,120.29);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37, 64.67) rectangle (306.77,120.29);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 75.12) --
	(306.77, 75.12);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 92.18) --
	(306.77, 92.18);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,109.23) --
	(306.77,109.23);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 67.18, 64.67) --
	( 67.18,120.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (132.11, 64.67) --
	(132.11,120.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (197.04, 64.67) --
	(197.04,120.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.97, 64.67) --
	(261.97,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 66.59) --
	(306.77, 66.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 83.65) --
	(306.77, 83.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,100.70) --
	(306.77,100.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,117.76) --
	(306.77,117.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.65, 64.67) --
	( 99.65,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (164.57, 64.67) --
	(164.57,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.50, 64.67) --
	(229.50,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.43, 64.67) --
	(294.43,120.29);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 47.70,117.76) --
	( 60.69,117.76) --
	( 60.69,117.76) --
	( 73.67,117.76) --
	( 73.67,117.76) --
	( 86.66,117.76) --
	( 86.66,116.20) --
	( 99.65,116.20) --
	( 99.65,114.52) --
	(112.63,114.52) --
	(112.63,112.66) --
	(125.62,112.66) --
	(125.62,110.66) --
	(138.60,110.66) --
	(138.60,108.53) --
	(151.59,108.53) --
	(151.59,106.28) --
	(164.57,106.28) --
	(164.57,104.13) --
	(177.56,104.13) --
	(177.56,101.96) --
	(190.55,101.96) --
	(190.55, 99.89) --
	(203.53, 99.89) --
	(203.53, 97.67) --
	(216.52, 97.67) --
	(216.52, 95.77) --
	(229.50, 95.77) --
	(229.50, 93.71) --
	(242.49, 93.71) --
	(242.49, 91.68) --
	(255.47, 91.68) --
	(255.47, 89.56) --
	(268.46, 89.56) --
	(268.46, 87.59) --
	(281.44, 87.59) --
	(281.44, 85.63) --
	(294.43, 85.63) --
	(294.43, 78.25);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 47.70,117.76) --
	( 60.69,117.76) --
	( 60.69,117.76) --
	( 73.67,117.76) --
	( 73.67,117.76) --
	( 86.66,117.76) --
	( 86.66,115.63) --
	( 99.65,115.63) --
	( 99.65,112.88) --
	(112.63,112.88) --
	(112.63,109.89) --
	(125.62,109.89) --
	(125.62,107.28) --
	(138.60,107.28) --
	(138.60,104.24) --
	(151.59,104.24) --
	(151.59,101.45) --
	(164.57,101.45) --
	(164.57, 98.80) --
	(177.56, 98.80) --
	(177.56, 96.10) --
	(190.55, 96.10) --
	(190.55, 93.39) --
	(203.53, 93.39) --
	(203.53, 90.63) --
	(216.52, 90.63) --
	(216.52, 88.25) --
	(229.50, 88.25) --
	(229.50, 85.89) --
	(242.49, 85.89) --
	(242.49, 83.44) --
	(255.47, 83.44) --
	(255.47, 80.87) --
	(268.46, 80.87) --
	(268.46, 78.53) --
	(281.44, 78.53) --
	(281.44, 76.10) --
	(294.43, 76.10) --
	(294.43, 67.20);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 47.70,117.76) --
	( 60.69,117.76) --
	( 60.69,117.58) --
	( 73.67,117.58) --
	( 73.67,117.28) --
	( 86.66,117.28) --
	( 86.66,116.87) --
	( 99.65,116.87) --
	( 99.65,116.33) --
	(112.63,116.33) --
	(112.63,115.67) --
	(125.62,115.67) --
	(125.62,114.88) --
	(138.60,114.88) --
	(138.60,113.98) --
	(151.59,113.98) --
	(151.59,112.97) --
	(164.57,112.97) --
	(164.57,111.84) --
	(177.56,111.84) --
	(177.56,110.59) --
	(190.55,110.59) --
	(190.55,109.24) --
	(203.53,109.24) --
	(203.53,107.80) --
	(216.52,107.80) --
	(216.52,106.25) --
	(229.50,106.25) --
	(229.50,104.60) --
	(242.49,104.60) --
	(242.49,102.86) --
	(255.47,102.86) --
	(255.47,101.02) --
	(268.46,101.02) --
	(268.46, 99.07) --
	(281.44, 99.07) --
	(281.44, 97.06) --
	(294.43, 97.06) --
	(294.43, 94.98);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 47.70,117.76) --
	( 60.69,117.76) --
	( 60.69,117.54) --
	( 73.67,117.54) --
	( 73.67,117.16) --
	( 86.66,117.16) --
	( 86.66,116.62) --
	( 99.65,116.62) --
	( 99.65,115.92) --
	(112.63,115.92) --
	(112.63,115.11) --
	(125.62,115.11) --
	(125.62,114.17) --
	(138.60,114.17) --
	(138.60,113.06) --
	(151.59,113.06) --
	(151.59,111.82) --
	(164.57,111.82) --
	(164.57,110.39) --
	(177.56,110.39) --
	(177.56,108.82) --
	(190.55,108.82) --
	(190.55,107.21) --
	(203.53,107.21) --
	(203.53,105.45) --
	(216.52,105.45) --
	(216.52,103.59) --
	(229.50,103.59) --
	(229.50,101.50) --
	(242.49,101.50) --
	(242.49, 99.26) --
	(255.47, 99.26) --
	(255.47, 96.93) --
	(268.46, 96.93) --
	(268.46, 94.54) --
	(281.44, 94.54) --
	(281.44, 92.02) --
	(294.43, 92.02) --
	(294.43, 89.54);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 47.70,117.76) --
	( 60.69,117.76) --
	( 60.69,116.56) --
	( 73.67,116.56) --
	( 73.67,115.09) --
	( 86.66,115.09) --
	( 86.66,113.43) --
	( 99.65,113.43) --
	( 99.65,111.65) --
	(112.63,111.65) --
	(112.63,109.79) --
	(125.62,109.79) --
	(125.62,107.83) --
	(138.60,107.83) --
	(138.60,105.87) --
	(151.59,105.87) --
	(151.59,103.89) --
	(164.57,103.89) --
	(164.57,101.88) --
	(177.56,101.88) --
	(177.56, 99.85) --
	(190.55, 99.85) --
	(190.55, 97.85) --
	(203.53, 97.85) --
	(203.53, 95.84) --
	(216.52, 95.84) --
	(216.52, 93.85) --
	(229.50, 93.85) --
	(229.50, 91.86) --
	(242.49, 91.86) --
	(242.49, 89.89) --
	(255.47, 89.89) --
	(255.47, 87.93) --
	(268.46, 87.93) --
	(268.46, 85.96) --
	(281.44, 85.96) --
	(281.44, 84.03) --
	(294.43, 84.03) --
	(294.43, 82.03);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 47.70,117.76) --
	( 60.69,117.76) --
	( 60.69,116.30) --
	( 73.67,116.30) --
	( 73.67,114.43) --
	( 86.66,114.43) --
	( 86.66,112.27) --
	( 99.65,112.27) --
	( 99.65,109.97) --
	(112.63,109.97) --
	(112.63,107.69) --
	(125.62,107.69) --
	(125.62,105.38) --
	(138.60,105.38) --
	(138.60,102.96) --
	(151.59,102.96) --
	(151.59,100.54) --
	(164.57,100.54) --
	(164.57, 98.00) --
	(177.56, 98.00) --
	(177.56, 95.48) --
	(190.55, 95.48) --
	(190.55, 93.09) --
	(203.53, 93.09) --
	(203.53, 90.67) --
	(216.52, 90.67) --
	(216.52, 88.30) --
	(229.50, 88.30) --
	(229.50, 85.80) --
	(242.49, 85.80) --
	(242.49, 83.29) --
	(255.47, 83.29) --
	(255.47, 80.83) --
	(268.46, 80.83) --
	(268.46, 78.44) --
	(281.44, 78.44) --
	(281.44, 76.05) --
	(294.43, 76.05) --
	(294.43, 73.69);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37, 64.67) rectangle (306.77,120.29);
\end{scope}
\begin{scope}
\path[clip] (306.77,125.79) rectangle (319.71,181.41);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,125.79) rectangle (319.71,181.41);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,153.60) {Scenario 1};
\end{scope}
\begin{scope}
\path[clip] (306.77, 64.67) rectangle (319.71,120.29);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77, 64.67) rectangle (319.71,120.29);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14, 92.48) {Scenario 2};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.65, 61.92) --
	( 99.65, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (164.57, 61.92) --
	(164.57, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.50, 61.92) --
	(229.50, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.43, 61.92) --
	(294.43, 64.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 99.65, 54.21) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (164.57, 54.21) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (229.50, 54.21) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (294.43, 54.21) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,124.96) {0.85};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,142.01) {0.90};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,159.07) {0.95};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,176.12) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,127.71) --
	( 35.37,127.71);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,144.77) --
	( 35.37,144.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,161.82) --
	( 35.37,161.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,178.88) --
	( 35.37,178.88);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 63.84) {0.85};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 80.89) {0.90};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 97.95) {0.95};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,115.00) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 66.59) --
	( 35.37, 66.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 83.65) --
	( 35.37, 83.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,100.70) --
	( 35.37,100.70);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,117.76) --
	( 35.37,117.76);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (171.07, 43.70) {Years since hire};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,123.04) {Cancer-free survival};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 21.60,  5.50) rectangle (144.67, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 32.60, 11.00) rectangle ( 47.05, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.05, 18.23) -- ( 45.61, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 88.63, 11.00) rectangle (103.09, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 90.08, 18.23) -- (101.64, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 52.55, 15.13) {Rule $a_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (108.59, 15.13) {Rule $a_1$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (155.67,  5.50) rectangle (320.53, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (166.67, 11.00) rectangle (181.12, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.11, 18.23) -- (179.68, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (214.89, 11.00) rectangle (229.34, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (216.34, 18.23) -- (227.90, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (264.84, 11.00) rectangle (279.29, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (266.28, 18.23) -- (277.85, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (186.62, 15.13) {Truth};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (234.84, 15.13) {WKM};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (284.79, 15.13) {AWKM};
\end{scope}
\end{tikzpicture}

\end{document}