### 2022-03-04 reports/resources/gm_results_3-4.tex

% Created by tikzDevice version 0.12.3.1 on 2022-01-19 23:30:37
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,186.91);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (325.21,186.91);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,125.79) rectangle (306.77,181.41);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,125.79) rectangle (306.77,181.41);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,135.73) --
	(306.77,135.73);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,152.99) --
	(306.77,152.99);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,170.25) --
	(306.77,170.25);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,125.79) --
	( 63.65,181.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,125.79) --
	(129.54,181.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,125.79) --
	(195.42,181.41);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,125.79) --
	(261.31,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,127.10) --
	(306.77,127.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,144.36) --
	(306.77,144.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,161.62) --
	(306.77,161.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,178.88) --
	(306.77,178.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,125.79) --
	( 96.59,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,125.79) --
	(162.48,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,125.79) --
	(228.36,181.41);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,125.79) --
	(294.25,181.41);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,178.88) --
	( 57.06,178.88) --
	( 57.06,178.88) --
	( 70.24,178.88) --
	( 70.24,178.88) --
	( 83.42,178.88) --
	( 83.42,178.20) --
	( 96.59,178.20) --
	( 96.59,177.27) --
	(109.77,177.27) --
	(109.77,176.28) --
	(122.95,176.28) --
	(122.95,175.28) --
	(136.12,175.28) --
	(136.12,174.14) --
	(149.30,174.14) --
	(149.30,173.02) --
	(162.48,173.02) --
	(162.48,172.01) --
	(175.66,172.01) --
	(175.66,170.88) --
	(188.83,170.88) --
	(188.83,169.86) --
	(202.01,169.86) --
	(202.01,168.78) --
	(215.19,168.78) --
	(215.19,167.74) --
	(228.36,167.74) --
	(228.36,166.69) --
	(241.54,166.69) --
	(241.54,165.71) --
	(254.72,165.71) --
	(254.72,164.68) --
	(267.89,164.68) --
	(267.89,163.58) --
	(281.07,163.58) --
	(281.07,162.52) --
	(294.25,162.52) --
	(294.25,158.92);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,178.88) --
	( 57.06,178.88) --
	( 57.06,178.88) --
	( 70.24,178.88) --
	( 70.24,178.88) --
	( 83.42,178.88) --
	( 83.42,177.76) --
	( 96.59,177.76) --
	( 96.59,176.36) --
	(109.77,176.36) --
	(109.77,174.82) --
	(122.95,174.82) --
	(122.95,173.48) --
	(136.12,173.48) --
	(136.12,171.94) --
	(149.30,171.94) --
	(149.30,170.53) --
	(162.48,170.53) --
	(162.48,169.10) --
	(175.66,169.10) --
	(175.66,167.63) --
	(188.83,167.63) --
	(188.83,166.19) --
	(202.01,166.19) --
	(202.01,164.67) --
	(215.19,164.67) --
	(215.19,163.39) --
	(228.36,163.39) --
	(228.36,162.12) --
	(241.54,162.12) --
	(241.54,160.87) --
	(254.72,160.87) --
	(254.72,159.70) --
	(267.89,159.70) --
	(267.89,158.46) --
	(281.07,158.46) --
	(281.07,157.23) --
	(294.25,157.23) --
	(294.25,152.51);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,178.88) --
	( 57.06,178.88) --
	( 57.06,178.79) --
	( 70.24,178.79) --
	( 70.24,178.64) --
	( 83.42,178.64) --
	( 83.42,178.43) --
	( 96.59,178.43) --
	( 96.59,178.16) --
	(109.77,178.16) --
	(109.77,177.83) --
	(122.95,177.83) --
	(122.95,177.45) --
	(136.12,177.45) --
	(136.12,177.01) --
	(149.30,177.01) --
	(149.30,176.51) --
	(162.48,176.51) --
	(162.48,175.96) --
	(175.66,175.96) --
	(175.66,175.35) --
	(188.83,175.35) --
	(188.83,174.68) --
	(202.01,174.68) --
	(202.01,173.96) --
	(215.19,173.96) --
	(215.19,173.19) --
	(228.36,173.19) --
	(228.36,172.39) --
	(241.54,172.39) --
	(241.54,171.52) --
	(254.72,171.52) --
	(254.72,170.61) --
	(267.89,170.61) --
	(267.89,169.65) --
	(281.07,169.65) --
	(281.07,168.65) --
	(294.25,168.65) --
	(294.25,167.61);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,178.88) --
	( 57.06,178.88) --
	( 57.06,178.75) --
	( 70.24,178.75) --
	( 70.24,178.56) --
	( 83.42,178.56) --
	( 83.42,178.28) --
	( 96.59,178.28) --
	( 96.59,177.93) --
	(109.77,177.93) --
	(109.77,177.51) --
	(122.95,177.51) --
	(122.95,177.02) --
	(136.12,177.02) --
	(136.12,176.46) --
	(149.30,176.46) --
	(149.30,175.83) --
	(162.48,175.83) --
	(162.48,175.14) --
	(175.66,175.14) --
	(175.66,174.36) --
	(188.83,174.36) --
	(188.83,173.54) --
	(202.01,173.54) --
	(202.01,172.62) --
	(215.19,172.62) --
	(215.19,171.66) --
	(228.36,171.66) --
	(228.36,170.63) --
	(241.54,170.63) --
	(241.54,169.55) --
	(254.72,169.55) --
	(254.72,168.43) --
	(267.89,168.43) --
	(267.89,167.23) --
	(281.07,167.23) --
	(281.07,165.99) --
	(294.25,165.99) --
	(294.25,164.69);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,178.88) --
	( 57.06,178.88) --
	( 57.06,178.27) --
	( 70.24,178.27) --
	( 70.24,177.54) --
	( 83.42,177.54) --
	( 83.42,176.69) --
	( 96.59,176.69) --
	( 96.59,175.81) --
	(109.77,175.81) --
	(109.77,174.87) --
	(122.95,174.87) --
	(122.95,173.92) --
	(136.12,173.92) --
	(136.12,172.97) --
	(149.30,172.97) --
	(149.30,171.98) --
	(162.48,171.98) --
	(162.48,170.99) --
	(175.66,170.99) --
	(175.66,169.99) --
	(188.83,169.99) --
	(188.83,168.99) --
	(202.01,168.99) --
	(202.01,168.00) --
	(215.19,168.00) --
	(215.19,167.00) --
	(228.36,167.00) --
	(228.36,166.02) --
	(241.54,166.02) --
	(241.54,165.04) --
	(254.72,165.04) --
	(254.72,164.06) --
	(267.89,164.06) --
	(267.89,163.09) --
	(281.07,163.09) --
	(281.07,162.13) --
	(294.25,162.13) --
	(294.25,161.13);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,178.88) --
	( 57.06,178.88) --
	( 57.06,178.06) --
	( 70.24,178.06) --
	( 70.24,177.09) --
	( 83.42,177.09) --
	( 83.42,175.98) --
	( 96.59,175.98) --
	( 96.59,174.82) --
	(109.77,174.82) --
	(109.77,173.63) --
	(122.95,173.63) --
	(122.95,172.40) --
	(136.12,172.40) --
	(136.12,171.20) --
	(149.30,171.20) --
	(149.30,169.95) --
	(162.48,169.95) --
	(162.48,168.72) --
	(175.66,168.72) --
	(175.66,167.46) --
	(188.83,167.46) --
	(188.83,166.23) --
	(202.01,166.23) --
	(202.01,164.97) --
	(215.19,164.97) --
	(215.19,163.73) --
	(228.36,163.73) --
	(228.36,162.50) --
	(241.54,162.50) --
	(241.54,161.29) --
	(254.72,161.29) --
	(254.72,160.10) --
	(267.89,160.10) --
	(267.89,158.90) --
	(281.07,158.90) --
	(281.07,157.72) --
	(294.25,157.72) --
	(294.25,156.49);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,125.79) rectangle (306.77,181.41);
\end{scope}
\begin{scope}
\path[clip] ( 31.37, 64.67) rectangle (306.77,120.29);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37, 64.67) rectangle (306.77,120.29);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37, 74.61) --
	(306.77, 74.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37, 91.87) --
	(306.77, 91.87);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,109.13) --
	(306.77,109.13);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65, 64.67) --
	( 63.65,120.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54, 64.67) --
	(129.54,120.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42, 64.67) --
	(195.42,120.29);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31, 64.67) --
	(261.31,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 65.98) --
	(306.77, 65.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 83.24) --
	(306.77, 83.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,100.50) --
	(306.77,100.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,117.76) --
	(306.77,117.76);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 64.67) --
	( 96.59,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 64.67) --
	(162.48,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 64.67) --
	(228.36,120.29);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 64.67) --
	(294.25,120.29);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,117.76) --
	( 57.06,117.76) --
	( 57.06,117.76) --
	( 70.24,117.76) --
	( 70.24,117.76) --
	( 83.42,117.76) --
	( 83.42,116.34) --
	( 96.59,116.34) --
	( 96.59,114.46) --
	(109.77,114.46) --
	(109.77,112.33) --
	(122.95,112.33) --
	(122.95,109.82) --
	(136.12,109.82) --
	(136.12,107.36) --
	(149.30,107.36) --
	(149.30,104.92) --
	(162.48,104.92) --
	(162.48,102.44) --
	(175.66,102.44) --
	(175.66,100.04) --
	(188.83,100.04) --
	(188.83, 97.63) --
	(202.01, 97.63) --
	(202.01, 95.21) --
	(215.19, 95.21) --
	(215.19, 92.93) --
	(228.36, 92.93) --
	(228.36, 90.77) --
	(241.54, 90.77) --
	(241.54, 88.50) --
	(254.72, 88.50) --
	(254.72, 86.37) --
	(267.89, 86.37) --
	(267.89, 84.35) --
	(281.07, 84.35) --
	(281.07, 82.26) --
	(294.25, 82.26) --
	(294.25, 75.17);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,117.76) --
	( 57.06,117.76) --
	( 57.06,117.76) --
	( 70.24,117.76) --
	( 70.24,117.76) --
	( 83.42,117.76) --
	( 83.42,115.96) --
	( 96.59,115.96) --
	( 96.59,113.38) --
	(109.77,113.38) --
	(109.77,110.41) --
	(122.95,110.41) --
	(122.95,107.23) --
	(136.12,107.23) --
	(136.12,104.13) --
	(149.30,104.13) --
	(149.30,101.03) --
	(162.48,101.03) --
	(162.48, 97.99) --
	(175.66, 97.99) --
	(175.66, 94.91) --
	(188.83, 94.91) --
	(188.83, 92.06) --
	(202.01, 92.06) --
	(202.01, 89.37) --
	(215.19, 89.37) --
	(215.19, 86.81) --
	(228.36, 86.81) --
	(228.36, 84.24) --
	(241.54, 84.24) --
	(241.54, 81.76) --
	(254.72, 81.76) --
	(254.72, 79.43) --
	(267.89, 79.43) --
	(267.89, 77.13) --
	(281.07, 77.13) --
	(281.07, 74.99) --
	(294.25, 74.99) --
	(294.25, 67.20);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,117.76) --
	( 57.06,117.76) --
	( 57.06,117.57) --
	( 70.24,117.57) --
	( 70.24,117.24) --
	( 83.42,117.24) --
	( 83.42,116.77) --
	( 96.59,116.77) --
	( 96.59,116.16) --
	(109.77,116.16) --
	(109.77,115.39) --
	(122.95,115.39) --
	(122.95,114.50) --
	(136.12,114.50) --
	(136.12,113.48) --
	(149.30,113.48) --
	(149.30,112.33) --
	(162.48,112.33) --
	(162.48,111.06) --
	(175.66,111.06) --
	(175.66,109.63) --
	(188.83,109.63) --
	(188.83,108.11) --
	(202.01,108.11) --
	(202.01,106.47) --
	(215.19,106.47) --
	(215.19,104.73) --
	(228.36,104.73) --
	(228.36,102.88) --
	(241.54,102.88) --
	(241.54,100.92) --
	(254.72,100.92) --
	(254.72, 98.90) --
	(267.89, 98.90) --
	(267.89, 96.78) --
	(281.07, 96.78) --
	(281.07, 94.61) --
	(294.25, 94.61) --
	(294.25, 92.33);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,117.76) --
	( 57.06,117.76) --
	( 57.06,117.48) --
	( 70.24,117.48) --
	( 70.24,117.04) --
	( 83.42,117.04) --
	( 83.42,116.43) --
	( 96.59,116.43) --
	( 96.59,115.66) --
	(109.77,115.66) --
	(109.77,114.74) --
	(122.95,114.74) --
	(122.95,113.65) --
	(136.12,113.65) --
	(136.12,112.42) --
	(149.30,112.42) --
	(149.30,111.04) --
	(162.48,111.04) --
	(162.48,109.53) --
	(175.66,109.53) --
	(175.66,107.85) --
	(188.83,107.85) --
	(188.83,106.01) --
	(202.01,106.01) --
	(202.01,104.07) --
	(215.19,104.07) --
	(215.19,102.02) --
	(228.36,102.02) --
	(228.36, 99.79) --
	(241.54, 99.79) --
	(241.54, 97.43) --
	(254.72, 97.43) --
	(254.72, 94.96) --
	(267.89, 94.96) --
	(267.89, 92.49) --
	(281.07, 92.49) --
	(281.07, 89.90) --
	(294.25, 89.90) --
	(294.25, 87.36);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,117.76) --
	( 57.06,117.76) --
	( 57.06,116.48) --
	( 70.24,116.48) --
	( 70.24,114.87) --
	( 83.42,114.87) --
	( 83.42,112.98) --
	( 96.59,112.98) --
	( 96.59,110.97) --
	(109.77,110.97) --
	(109.77,108.81) --
	(122.95,108.81) --
	(122.95,106.61) --
	(136.12,106.61) --
	(136.12,104.40) --
	(149.30,104.40) --
	(149.30,102.19) --
	(162.48,102.19) --
	(162.48, 99.96) --
	(175.66, 99.96) --
	(175.66, 97.69) --
	(188.83, 97.69) --
	(188.83, 95.46) --
	(202.01, 95.46) --
	(202.01, 93.26) --
	(215.19, 93.26) --
	(215.19, 91.09) --
	(228.36, 91.09) --
	(228.36, 88.95) --
	(241.54, 88.95) --
	(241.54, 86.82) --
	(254.72, 86.82) --
	(254.72, 84.77) --
	(267.89, 84.77) --
	(267.89, 82.73) --
	(281.07, 82.73) --
	(281.07, 80.77) --
	(294.25, 80.77) --
	(294.25, 78.69);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,117.76) --
	( 57.06,117.76) --
	( 57.06,115.90) --
	( 70.24,115.90) --
	( 70.24,113.71) --
	( 83.42,113.71) --
	( 83.42,111.29) --
	( 96.59,111.29) --
	( 96.59,108.77) --
	(109.77,108.77) --
	(109.77,106.19) --
	(122.95,106.19) --
	(122.95,103.53) --
	(136.12,103.53) --
	(136.12,100.90) --
	(149.30,100.90) --
	(149.30, 98.25) --
	(162.48, 98.25) --
	(162.48, 95.64) --
	(175.66, 95.64) --
	(175.66, 93.00) --
	(188.83, 93.00) --
	(188.83, 90.37) --
	(202.01, 90.37) --
	(202.01, 87.81) --
	(215.19, 87.81) --
	(215.19, 85.31) --
	(228.36, 85.31) --
	(228.36, 82.77) --
	(241.54, 82.77) --
	(241.54, 80.27) --
	(254.72, 80.27) --
	(254.72, 77.82) --
	(267.89, 77.82) --
	(267.89, 75.51) --
	(281.07, 75.51) --
	(281.07, 73.21) --
	(294.25, 73.21) --
	(294.25, 70.96);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37, 64.67) rectangle (306.77,120.29);
\end{scope}
\begin{scope}
\path[clip] (306.77,125.79) rectangle (319.71,181.41);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,125.79) rectangle (319.71,181.41);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,153.60) {Scenario 3};
\end{scope}
\begin{scope}
\path[clip] (306.77, 64.67) rectangle (319.71,120.29);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77, 64.67) rectangle (319.71,120.29);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14, 92.48) {Scenario 4};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 61.92) --
	( 96.59, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 61.92) --
	(162.48, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 61.92) --
	(228.36, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 61.92) --
	(294.25, 64.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 96.59, 54.21) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (162.48, 54.21) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (228.36, 54.21) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (294.25, 54.21) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,124.34) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,141.60) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,158.86) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,176.12) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,127.10) --
	( 31.37,127.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,144.36) --
	( 31.37,144.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,161.62) --
	( 31.37,161.62);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,178.88) --
	( 31.37,178.88);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 63.23) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 80.48) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 97.74) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,115.00) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 65.98) --
	( 31.37, 65.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 83.24) --
	( 31.37, 83.24);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,100.50) --
	( 31.37,100.50);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,117.76) --
	( 31.37,117.76);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (169.07, 43.70) {Years since hire};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,123.04) {Cancer-free survival};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 19.60,  5.50) rectangle (142.67, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 30.60, 11.00) rectangle ( 45.05, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.05, 18.23) -- ( 43.61, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 86.63, 11.00) rectangle (101.09, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.08, 18.23) -- ( 99.64, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 50.55, 15.13) {Rule $a_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (106.59, 15.13) {Rule $a_1$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (153.67,  5.50) rectangle (318.53, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (164.67, 11.00) rectangle (179.12, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.11, 18.23) -- (177.68, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.89, 11.00) rectangle (227.34, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (214.34, 18.23) -- (225.90, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (262.84, 11.00) rectangle (277.29, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (264.28, 18.23) -- (275.85, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (184.62, 15.13) {Truth};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (232.84, 15.13) {WKM};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,186.91);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (282.79, 15.13) {AWKM};
\end{scope}
\end{tikzpicture}

\end{document}