### 2022-03-04 reports/resources/gm_results_5.tex

% Created by tikzDevice version 0.12.3.1 on 2022-01-19 23:32:12
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,135.51);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (325.21,135.51);
\end{scope}
\begin{scope}
\path[clip] ( 31.37, 64.67) rectangle (306.77,130.01);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37, 64.67) rectangle (306.77,130.01);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37, 78.57) --
	(306.77, 78.57);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37, 97.96) --
	(306.77, 97.96);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,117.34) --
	(306.77,117.34);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65, 64.67) --
	( 63.65,130.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54, 64.67) --
	(129.54,130.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42, 64.67) --
	(195.42,130.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31, 64.67) --
	(261.31,130.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 68.87) --
	(306.77, 68.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 88.26) --
	(306.77, 88.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,107.65) --
	(306.77,107.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,127.04) --
	(306.77,127.04);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 64.67) --
	( 96.59,130.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 64.67) --
	(162.48,130.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 64.67) --
	(228.36,130.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 64.67) --
	(294.25,130.01);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,127.04) --
	( 57.06,127.04) --
	( 57.06,127.04) --
	( 70.24,127.04) --
	( 70.24,127.04) --
	( 83.42,127.04) --
	( 83.42,124.66) --
	( 96.59,124.66) --
	( 96.59,122.06) --
	(109.77,122.06) --
	(109.77,119.04) --
	(122.95,119.04) --
	(122.95,116.10) --
	(136.12,116.10) --
	(136.12,113.15) --
	(149.30,113.15) --
	(149.30,110.13) --
	(162.48,110.13) --
	(162.48,107.32) --
	(175.66,107.32) --
	(175.66,104.56) --
	(188.83,104.56) --
	(188.83,101.93) --
	(202.01,101.93) --
	(202.01, 99.23) --
	(215.19, 99.23) --
	(215.19, 96.51) --
	(228.36, 96.51) --
	(228.36, 94.06) --
	(241.54, 94.06) --
	(241.54, 91.66) --
	(254.72, 91.66) --
	(254.72, 89.32) --
	(267.89, 89.32) --
	(267.89, 87.14) --
	(281.07, 87.14) --
	(281.07, 85.07) --
	(294.25, 85.07) --
	(294.25, 76.99);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,127.04) --
	( 57.06,127.04) --
	( 57.06,127.04) --
	( 70.24,127.04) --
	( 70.24,127.04) --
	( 83.42,127.04) --
	( 83.42,124.02) --
	( 96.59,124.02) --
	( 96.59,120.04) --
	(109.77,120.04) --
	(109.77,115.99) --
	(122.95,115.99) --
	(122.95,112.23) --
	(136.12,112.23) --
	(136.12,108.44) --
	(149.30,108.44) --
	(149.30,105.00) --
	(162.48,105.00) --
	(162.48,101.60) --
	(175.66,101.60) --
	(175.66, 98.02) --
	(188.83, 98.02) --
	(188.83, 94.76) --
	(202.01, 94.76) --
	(202.01, 91.83) --
	(215.19, 91.83) --
	(215.19, 88.97) --
	(228.36, 88.97) --
	(228.36, 86.33) --
	(241.54, 86.33) --
	(241.54, 83.71) --
	(254.72, 83.71) --
	(254.72, 81.09) --
	(267.89, 81.09) --
	(267.89, 78.47) --
	(281.07, 78.47) --
	(281.07, 76.21) --
	(294.25, 76.21) --
	(294.25, 67.64);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,127.04) --
	( 57.06,127.04) --
	( 57.06,126.77) --
	( 70.24,126.77) --
	( 70.24,126.34) --
	( 83.42,126.34) --
	( 83.42,125.76) --
	( 96.59,125.76) --
	( 96.59,125.03) --
	(109.77,125.03) --
	(109.77,124.17) --
	(122.95,124.17) --
	(122.95,123.15) --
	(136.12,123.15) --
	(136.12,122.00) --
	(149.30,122.00) --
	(149.30,120.71) --
	(162.48,120.71) --
	(162.48,119.27) --
	(175.66,119.27) --
	(175.66,117.70) --
	(188.83,117.70) --
	(188.83,116.01) --
	(202.01,116.01) --
	(202.01,114.18) --
	(215.19,114.18) --
	(215.19,112.23) --
	(228.36,112.23) --
	(228.36,110.16) --
	(241.54,110.16) --
	(241.54,108.00) --
	(254.72,108.00) --
	(254.72,105.73) --
	(267.89,105.73) --
	(267.89,103.36) --
	(281.07,103.36) --
	(281.07,100.91) --
	(294.25,100.91) --
	(294.25, 98.40);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,127.04) --
	( 57.06,127.04) --
	( 57.06,126.69) --
	( 70.24,126.69) --
	( 70.24,126.14) --
	( 83.42,126.14) --
	( 83.42,125.40) --
	( 96.59,125.40) --
	( 96.59,124.49) --
	(109.77,124.49) --
	(109.77,123.37) --
	(122.95,123.37) --
	(122.95,122.15) --
	(136.12,122.15) --
	(136.12,120.78) --
	(149.30,120.78) --
	(149.30,119.19) --
	(162.48,119.19) --
	(162.48,117.49) --
	(175.66,117.49) --
	(175.66,115.61) --
	(188.83,115.61) --
	(188.83,113.58) --
	(202.01,113.58) --
	(202.01,111.40) --
	(215.19,111.40) --
	(215.19,109.11) --
	(228.36,109.11) --
	(228.36,106.72) --
	(241.54,106.72) --
	(241.54,104.23) --
	(254.72,104.23) --
	(254.72,101.60) --
	(267.89,101.60) --
	(267.89, 98.83) --
	(281.07, 98.83) --
	(281.07, 95.98) --
	(294.25, 95.98) --
	(294.25, 93.00);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,127.04) --
	( 57.06,127.04) --
	( 57.06,125.25) --
	( 70.24,125.25) --
	( 70.24,123.13) --
	( 83.42,123.13) --
	( 83.42,120.82) --
	( 96.59,120.82) --
	( 96.59,118.41) --
	(109.77,118.41) --
	(109.77,115.98) --
	(122.95,115.98) --
	(122.95,113.50) --
	(136.12,113.50) --
	(136.12,111.00) --
	(149.30,111.00) --
	(149.30,108.51) --
	(162.48,108.51) --
	(162.48,106.02) --
	(175.66,106.02) --
	(175.66,103.53) --
	(188.83,103.53) --
	(188.83,101.08) --
	(202.01,101.08) --
	(202.01, 98.62) --
	(215.19, 98.62) --
	(215.19, 96.21) --
	(228.36, 96.21) --
	(228.36, 93.81) --
	(241.54, 93.81) --
	(241.54, 91.48) --
	(254.72, 91.48) --
	(254.72, 89.18) --
	(267.89, 89.18) --
	(267.89, 86.91) --
	(281.07, 86.91) --
	(281.07, 84.70) --
	(294.25, 84.70) --
	(294.25, 82.43);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,127.04) --
	( 57.06,127.04) --
	( 57.06,124.69) --
	( 70.24,124.69) --
	( 70.24,121.98) --
	( 83.42,121.98) --
	( 83.42,119.04) --
	( 96.59,119.04) --
	( 96.59,116.05) --
	(109.77,116.05) --
	(109.77,112.90) --
	(122.95,112.90) --
	(122.95,109.96) --
	(136.12,109.96) --
	(136.12,107.02) --
	(149.30,107.02) --
	(149.30,103.98) --
	(162.48,103.98) --
	(162.48,101.05) --
	(175.66,101.05) --
	(175.66, 98.12) --
	(188.83, 98.12) --
	(188.83, 95.23) --
	(202.01, 95.23) --
	(202.01, 92.36) --
	(215.19, 92.36) --
	(215.19, 89.58) --
	(228.36, 89.58) --
	(228.36, 86.88) --
	(241.54, 86.88) --
	(241.54, 84.25) --
	(254.72, 84.25) --
	(254.72, 81.64) --
	(267.89, 81.64) --
	(267.89, 79.06) --
	(281.07, 79.06) --
	(281.07, 76.54) --
	(294.25, 76.54) --
	(294.25, 73.92);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37, 64.67) rectangle (306.77,130.01);
\end{scope}
\begin{scope}
\path[clip] (306.77, 64.67) rectangle (319.71,130.01);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77, 64.67) rectangle (319.71,130.01);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14, 97.34) {Scenario 5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 61.92) --
	( 96.59, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 61.92) --
	(162.48, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 61.92) --
	(228.36, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 61.92) --
	(294.25, 64.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 96.59, 54.21) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (162.48, 54.21) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (228.36, 54.21) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (294.25, 54.21) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 66.12) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 85.51) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,104.89) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,124.28) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 68.87) --
	( 31.37, 68.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 88.26) --
	( 31.37, 88.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,107.65) --
	( 31.37,107.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,127.04) --
	( 31.37,127.04);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (169.07, 43.70) {Years since hire};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70, 97.34) {Survival};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 19.60,  5.50) rectangle (142.67, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 30.60, 11.00) rectangle ( 45.05, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.05, 18.23) -- ( 43.61, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 86.63, 11.00) rectangle (101.09, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.08, 18.23) -- ( 99.64, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 50.55, 15.13) {Rule $a_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (106.59, 15.13) {Rule $a_1$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (153.67,  5.50) rectangle (318.53, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (164.67, 11.00) rectangle (179.12, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.11, 18.23) -- (177.68, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.89, 11.00) rectangle (227.34, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (214.34, 18.23) -- (225.90, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (262.84, 11.00) rectangle (277.29, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (264.28, 18.23) -- (275.85, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (184.62, 15.13) {Truth};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (232.84, 15.13) {WKM};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,135.51);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (282.79, 15.13) {AWKM};
\end{scope}
\end{tikzpicture}

\end{document}