### 2022-03-04 reports/resources/results.tex

% Created by tikzDevice version 0.12.3.1 on 2022-03-04 00:12:39
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,560.09);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (325.21,560.09);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,463.90) rectangle (306.77,554.59);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,463.90) rectangle (306.77,554.59);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,483.19) --
	(306.77,483.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,510.10) --
	(306.77,510.10);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,537.01) --
	(306.77,537.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,463.90) --
	( 63.65,554.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,463.90) --
	(129.54,554.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,463.90) --
	(195.42,554.59);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,463.90) --
	(261.31,554.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,469.74) --
	(306.77,469.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,496.65) --
	(306.77,496.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,523.56) --
	(306.77,523.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,550.47) --
	(306.77,550.47);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,463.90) --
	( 96.59,554.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,463.90) --
	(162.48,554.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,463.90) --
	(228.36,554.59);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,463.90) --
	(294.25,554.59);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,550.47) --
	( 57.06,550.47) --
	( 57.06,550.47) --
	( 70.24,550.47) --
	( 70.24,550.47) --
	( 83.42,550.47) --
	( 83.42,548.70) --
	( 96.59,548.70) --
	( 96.59,546.65) --
	(109.77,546.65) --
	(109.77,544.43) --
	(122.95,544.43) --
	(122.95,542.36) --
	(136.12,542.36) --
	(136.12,540.08) --
	(149.30,540.08) --
	(149.30,537.68) --
	(162.48,537.68) --
	(162.48,535.56) --
	(175.66,535.56) --
	(175.66,533.45) --
	(188.83,533.45) --
	(188.83,531.30) --
	(202.01,531.30) --
	(202.01,529.16) --
	(215.19,529.16) --
	(215.19,527.28) --
	(228.36,527.28) --
	(228.36,525.33) --
	(241.54,525.33) --
	(241.54,523.49) --
	(254.72,523.49) --
	(254.72,521.48) --
	(267.89,521.48) --
	(267.89,519.66) --
	(281.07,519.66) --
	(281.07,517.90) --
	(294.25,517.90) --
	(294.25,511.26);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,550.47) --
	( 57.06,550.47) --
	( 57.06,550.47) --
	( 70.24,550.47) --
	( 70.24,550.47) --
	( 83.42,550.47) --
	( 83.42,549.32) --
	( 96.59,549.32) --
	( 96.59,547.94) --
	(109.77,547.94) --
	(109.77,546.41) --
	(122.95,546.41) --
	(122.95,544.63) --
	(136.12,544.63) --
	(136.12,543.11) --
	(149.30,543.11) --
	(149.30,541.33) --
	(162.48,541.33) --
	(162.48,539.45) --
	(175.66,539.45) --
	(175.66,537.62) --
	(188.83,537.62) --
	(188.83,536.04) --
	(202.01,536.04) --
	(202.01,534.45) --
	(215.19,534.45) --
	(215.19,532.67) --
	(228.36,532.67) --
	(228.36,530.97) --
	(241.54,530.97) --
	(241.54,529.42) --
	(254.72,529.42) --
	(254.72,527.80) --
	(267.89,527.80) --
	(267.89,526.23) --
	(281.07,526.23) --
	(281.07,524.66) --
	(294.25,524.66) --
	(294.25,519.09);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,550.47) --
	( 57.06,550.47) --
	( 57.06,550.37) --
	( 70.24,550.37) --
	( 70.24,550.22) --
	( 83.42,550.22) --
	( 83.42,550.01) --
	( 96.59,550.01) --
	( 96.59,549.74) --
	(109.77,549.74) --
	(109.77,549.42) --
	(122.95,549.42) --
	(122.95,549.05) --
	(136.12,549.05) --
	(136.12,548.63) --
	(149.30,548.63) --
	(149.30,548.14) --
	(162.48,548.14) --
	(162.48,547.61) --
	(175.66,547.61) --
	(175.66,547.02) --
	(188.83,547.02) --
	(188.83,546.38) --
	(202.01,546.38) --
	(202.01,545.71) --
	(215.19,545.71) --
	(215.19,544.99) --
	(228.36,544.99) --
	(228.36,544.21) --
	(241.54,544.21) --
	(241.54,543.38) --
	(254.72,543.38) --
	(254.72,542.50) --
	(267.89,542.50) --
	(267.89,541.57) --
	(281.07,541.57) --
	(281.07,540.63) --
	(294.25,540.63) --
	(294.25,539.65);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,550.47) --
	( 57.06,550.47) --
	( 57.06,550.40) --
	( 70.24,550.40) --
	( 70.24,550.29) --
	( 83.42,550.29) --
	( 83.42,550.12) --
	( 96.59,550.12) --
	( 96.59,549.91) --
	(109.77,549.91) --
	(109.77,549.65) --
	(122.95,549.65) --
	(122.95,549.35) --
	(136.12,549.35) --
	(136.12,549.01) --
	(149.30,549.01) --
	(149.30,548.62) --
	(162.48,548.62) --
	(162.48,548.18) --
	(175.66,548.18) --
	(175.66,547.70) --
	(188.83,547.70) --
	(188.83,547.18) --
	(202.01,547.18) --
	(202.01,546.62) --
	(215.19,546.62) --
	(215.19,546.02) --
	(228.36,546.02) --
	(228.36,545.39) --
	(241.54,545.39) --
	(241.54,544.72) --
	(254.72,544.72) --
	(254.72,544.00) --
	(267.89,544.00) --
	(267.89,543.26) --
	(281.07,543.26) --
	(281.07,542.47) --
	(294.25,542.47) --
	(294.25,541.66);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,550.47) --
	( 57.06,550.47) --
	( 57.06,549.81) --
	( 70.24,549.81) --
	( 70.24,549.07) --
	( 83.42,549.07) --
	( 83.42,548.22) --
	( 96.59,548.22) --
	( 96.59,547.34) --
	(109.77,547.34) --
	(109.77,546.43) --
	(122.95,546.43) --
	(122.95,545.51) --
	(136.12,545.51) --
	(136.12,544.58) --
	(149.30,544.58) --
	(149.30,543.64) --
	(162.48,543.64) --
	(162.48,542.68) --
	(175.66,542.68) --
	(175.66,541.72) --
	(188.83,541.72) --
	(188.83,540.78) --
	(202.01,540.78) --
	(202.01,539.85) --
	(215.19,539.85) --
	(215.19,538.93) --
	(228.36,538.93) --
	(228.36,537.99) --
	(241.54,537.99) --
	(241.54,537.05) --
	(254.72,537.05) --
	(254.72,536.13) --
	(267.89,536.13) --
	(267.89,535.19) --
	(281.07,535.19) --
	(281.07,534.30) --
	(294.25,534.30) --
	(294.25,533.37);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,550.47) --
	( 57.06,550.47) --
	( 57.06,550.00) --
	( 70.24,550.00) --
	( 70.24,549.43) --
	( 83.42,549.43) --
	( 83.42,548.76) --
	( 96.59,548.76) --
	( 96.59,548.07) --
	(109.77,548.07) --
	(109.77,547.34) --
	(122.95,547.34) --
	(122.95,546.59) --
	(136.12,546.59) --
	(136.12,545.84) --
	(149.30,545.84) --
	(149.30,545.07) --
	(162.48,545.07) --
	(162.48,544.29) --
	(175.66,544.29) --
	(175.66,543.51) --
	(188.83,543.51) --
	(188.83,542.72) --
	(202.01,542.72) --
	(202.01,541.95) --
	(215.19,541.95) --
	(215.19,541.17) --
	(228.36,541.17) --
	(228.36,540.41) --
	(241.54,540.41) --
	(241.54,539.65) --
	(254.72,539.65) --
	(254.72,538.88) --
	(267.89,538.88) --
	(267.89,538.12) --
	(281.07,538.12) --
	(281.07,537.37) --
	(294.25,537.37) --
	(294.25,536.59);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,463.90) rectangle (306.77,554.59);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,367.70) rectangle (306.77,458.40);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,367.70) rectangle (306.77,458.40);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,387.00) --
	(306.77,387.00);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,413.91) --
	(306.77,413.91);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,440.82) --
	(306.77,440.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,367.70) --
	( 63.65,458.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,367.70) --
	(129.54,458.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,367.70) --
	(195.42,458.40);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,367.70) --
	(261.31,458.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,373.54) --
	(306.77,373.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,400.45) --
	(306.77,400.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,427.36) --
	(306.77,427.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,454.28) --
	(306.77,454.28);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,367.70) --
	( 96.59,458.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,367.70) --
	(162.48,458.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,367.70) --
	(228.36,458.40);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,367.70) --
	(294.25,458.40);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,454.28) --
	( 57.06,454.28) --
	( 57.06,454.28) --
	( 70.24,454.28) --
	( 70.24,454.28) --
	( 83.42,454.28) --
	( 83.42,452.60) --
	( 96.59,452.60) --
	( 96.59,450.43) --
	(109.77,450.43) --
	(109.77,448.06) --
	(122.95,448.06) --
	(122.95,446.01) --
	(136.12,446.01) --
	(136.12,443.61) --
	(149.30,443.61) --
	(149.30,441.41) --
	(162.48,441.41) --
	(162.48,439.32) --
	(175.66,439.32) --
	(175.66,437.19) --
	(188.83,437.19) --
	(188.83,435.05) --
	(202.01,435.05) --
	(202.01,432.88) --
	(215.19,432.88) --
	(215.19,431.00) --
	(228.36,431.00) --
	(228.36,429.13) --
	(241.54,429.13) --
	(241.54,427.20) --
	(254.72,427.20) --
	(254.72,425.17) --
	(267.89,425.17) --
	(267.89,423.32) --
	(281.07,423.32) --
	(281.07,421.41) --
	(294.25,421.41) --
	(294.25,414.38);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,454.28) --
	( 57.06,454.28) --
	( 57.06,454.28) --
	( 70.24,454.28) --
	( 70.24,454.28) --
	( 83.42,454.28) --
	( 83.42,453.04) --
	( 96.59,453.04) --
	( 96.59,451.72) --
	(109.77,451.72) --
	(109.77,450.26) --
	(122.95,450.26) --
	(122.95,448.67) --
	(136.12,448.67) --
	(136.12,447.00) --
	(149.30,447.00) --
	(149.30,445.22) --
	(162.48,445.22) --
	(162.48,443.53) --
	(175.66,443.53) --
	(175.66,441.81) --
	(188.83,441.81) --
	(188.83,440.18) --
	(202.01,440.18) --
	(202.01,438.43) --
	(215.19,438.43) --
	(215.19,436.93) --
	(228.36,436.93) --
	(228.36,435.30) --
	(241.54,435.30) --
	(241.54,433.70) --
	(254.72,433.70) --
	(254.72,432.03) --
	(267.89,432.03) --
	(267.89,430.48) --
	(281.07,430.48) --
	(281.07,428.93) --
	(294.25,428.93) --
	(294.25,423.11);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,454.28) --
	( 57.06,454.28) --
	( 57.06,454.19) --
	( 70.24,454.19) --
	( 70.24,454.04) --
	( 83.42,454.04) --
	( 83.42,453.83) --
	( 96.59,453.83) --
	( 96.59,453.55) --
	(109.77,453.55) --
	(109.77,453.23) --
	(122.95,453.23) --
	(122.95,452.86) --
	(136.12,452.86) --
	(136.12,452.42) --
	(149.30,452.42) --
	(149.30,451.93) --
	(162.48,451.93) --
	(162.48,451.37) --
	(175.66,451.37) --
	(175.66,450.75) --
	(188.83,450.75) --
	(188.83,450.11) --
	(202.01,450.11) --
	(202.01,449.41) --
	(215.19,449.41) --
	(215.19,448.68) --
	(228.36,448.68) --
	(228.36,447.85) --
	(241.54,447.85) --
	(241.54,446.97) --
	(254.72,446.97) --
	(254.72,446.05) --
	(267.89,446.05) --
	(267.89,445.11) --
	(281.07,445.11) --
	(281.07,444.11) --
	(294.25,444.11) --
	(294.25,443.13);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,454.28) --
	( 57.06,454.28) --
	( 57.06,454.20) --
	( 70.24,454.20) --
	( 70.24,454.09) --
	( 83.42,454.09) --
	( 83.42,453.92) --
	( 96.59,453.92) --
	( 96.59,453.71) --
	(109.77,453.71) --
	(109.77,453.45) --
	(122.95,453.45) --
	(122.95,453.14) --
	(136.12,453.14) --
	(136.12,452.78) --
	(149.30,452.78) --
	(149.30,452.39) --
	(162.48,452.39) --
	(162.48,451.94) --
	(175.66,451.94) --
	(175.66,451.45) --
	(188.83,451.45) --
	(188.83,450.91) --
	(202.01,450.91) --
	(202.01,450.34) --
	(215.19,450.34) --
	(215.19,449.73) --
	(228.36,449.73) --
	(228.36,449.08) --
	(241.54,449.08) --
	(241.54,448.39) --
	(254.72,448.39) --
	(254.72,447.66) --
	(267.89,447.66) --
	(267.89,446.90) --
	(281.07,446.90) --
	(281.07,446.10) --
	(294.25,446.10) --
	(294.25,445.28);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,454.28) --
	( 57.06,454.28) --
	( 57.06,453.70) --
	( 70.24,453.70) --
	( 70.24,452.96) --
	( 83.42,452.96) --
	( 83.42,452.11) --
	( 96.59,452.11) --
	( 96.59,451.20) --
	(109.77,451.20) --
	(109.77,450.30) --
	(122.95,450.30) --
	(122.95,449.39) --
	(136.12,449.39) --
	(136.12,448.43) --
	(149.30,448.43) --
	(149.30,447.48) --
	(162.48,447.48) --
	(162.48,446.47) --
	(175.66,446.47) --
	(175.66,445.48) --
	(188.83,445.48) --
	(188.83,444.53) --
	(202.01,444.53) --
	(202.01,443.58) --
	(215.19,443.58) --
	(215.19,442.64) --
	(228.36,442.64) --
	(228.36,441.65) --
	(241.54,441.65) --
	(241.54,440.66) --
	(254.72,440.66) --
	(254.72,439.69) --
	(267.89,439.69) --
	(267.89,438.75) --
	(281.07,438.75) --
	(281.07,437.80) --
	(294.25,437.80) --
	(294.25,436.87);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,454.28) --
	( 57.06,454.28) --
	( 57.06,453.80) --
	( 70.24,453.80) --
	( 70.24,453.22) --
	( 83.42,453.22) --
	( 83.42,452.57) --
	( 96.59,452.57) --
	( 96.59,451.87) --
	(109.77,451.87) --
	(109.77,451.13) --
	(122.95,451.13) --
	(122.95,450.36) --
	(136.12,450.36) --
	(136.12,449.58) --
	(149.30,449.58) --
	(149.30,448.80) --
	(162.48,448.80) --
	(162.48,448.01) --
	(175.66,448.01) --
	(175.66,447.21) --
	(188.83,447.21) --
	(188.83,446.41) --
	(202.01,446.41) --
	(202.01,445.62) --
	(215.19,445.62) --
	(215.19,444.84) --
	(228.36,444.84) --
	(228.36,444.05) --
	(241.54,444.05) --
	(241.54,443.27) --
	(254.72,443.27) --
	(254.72,442.50) --
	(267.89,442.50) --
	(267.89,441.72) --
	(281.07,441.72) --
	(281.07,440.96) --
	(294.25,440.96) --
	(294.25,440.17);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,367.70) rectangle (306.77,458.40);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,271.51) rectangle (306.77,362.20);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,271.51) rectangle (306.77,362.20);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,290.80) --
	(306.77,290.80);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,317.72) --
	(306.77,317.72);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,344.63) --
	(306.77,344.63);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,271.51) --
	( 63.65,362.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,271.51) --
	(129.54,362.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,271.51) --
	(195.42,362.20);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,271.51) --
	(261.31,362.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,277.35) --
	(306.77,277.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,304.26) --
	(306.77,304.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,331.17) --
	(306.77,331.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,358.08) --
	(306.77,358.08);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,271.51) --
	( 96.59,362.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,271.51) --
	(162.48,362.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,271.51) --
	(228.36,362.20);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,271.51) --
	(294.25,362.20);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,358.08) --
	( 57.06,358.08) --
	( 57.06,358.08) --
	( 70.24,358.08) --
	( 70.24,358.08) --
	( 83.42,358.08) --
	( 83.42,356.34) --
	( 96.59,356.34) --
	( 96.59,354.16) --
	(109.77,354.16) --
	(109.77,351.76) --
	(122.95,351.76) --
	(122.95,349.67) --
	(136.12,349.67) --
	(136.12,347.27) --
	(149.30,347.27) --
	(149.30,345.06) --
	(162.48,345.06) --
	(162.48,342.84) --
	(175.66,342.84) --
	(175.66,340.54) --
	(188.83,340.54) --
	(188.83,338.31) --
	(202.01,338.31) --
	(202.01,335.93) --
	(215.19,335.93) --
	(215.19,333.94) --
	(228.36,333.94) --
	(228.36,331.95) --
	(241.54,331.95) --
	(241.54,330.01) --
	(254.72,330.01) --
	(254.72,328.17) --
	(267.89,328.17) --
	(267.89,326.25) --
	(281.07,326.25) --
	(281.07,324.34) --
	(294.25,324.34) --
	(294.25,316.97);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,358.08) --
	( 57.06,358.08) --
	( 57.06,358.08) --
	( 70.24,358.08) --
	( 70.24,358.08) --
	( 83.42,358.08) --
	( 83.42,357.03) --
	( 96.59,357.03) --
	( 96.59,355.57) --
	(109.77,355.57) --
	(109.77,354.03) --
	(122.95,354.03) --
	(122.95,352.47) --
	(136.12,352.47) --
	(136.12,350.70) --
	(149.30,350.70) --
	(149.30,348.94) --
	(162.48,348.94) --
	(162.48,347.38) --
	(175.66,347.38) --
	(175.66,345.61) --
	(188.83,345.61) --
	(188.83,344.02) --
	(202.01,344.02) --
	(202.01,342.34) --
	(215.19,342.34) --
	(215.19,340.71) --
	(228.36,340.71) --
	(228.36,339.07) --
	(241.54,339.07) --
	(241.54,337.55) --
	(254.72,337.55) --
	(254.72,335.94) --
	(267.89,335.94) --
	(267.89,334.23) --
	(281.07,334.23) --
	(281.07,332.57) --
	(294.25,332.57) --
	(294.25,326.97);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,358.08) --
	( 57.06,358.08) --
	( 57.06,357.99) --
	( 70.24,357.99) --
	( 70.24,357.83) --
	( 83.42,357.83) --
	( 83.42,357.62) --
	( 96.59,357.62) --
	( 96.59,357.34) --
	(109.77,357.34) --
	(109.77,357.02) --
	(122.95,357.02) --
	(122.95,356.63) --
	(136.12,356.63) --
	(136.12,356.20) --
	(149.30,356.20) --
	(149.30,355.70) --
	(162.48,355.70) --
	(162.48,355.16) --
	(175.66,355.16) --
	(175.66,354.56) --
	(188.83,354.56) --
	(188.83,353.91) --
	(202.01,353.91) --
	(202.01,353.20) --
	(215.19,353.20) --
	(215.19,352.45) --
	(228.36,352.45) --
	(228.36,351.64) --
	(241.54,351.64) --
	(241.54,350.80) --
	(254.72,350.80) --
	(254.72,349.93) --
	(267.89,349.93) --
	(267.89,348.99) --
	(281.07,348.99) --
	(281.07,348.03) --
	(294.25,348.03) --
	(294.25,347.02);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,358.08) --
	( 57.06,358.08) --
	( 57.06,358.01) --
	( 70.24,358.01) --
	( 70.24,357.90) --
	( 83.42,357.90) --
	( 83.42,357.73) --
	( 96.59,357.73) --
	( 96.59,357.52) --
	(109.77,357.52) --
	(109.77,357.27) --
	(122.95,357.27) --
	(122.95,356.96) --
	(136.12,356.96) --
	(136.12,356.63) --
	(149.30,356.63) --
	(149.30,356.24) --
	(162.48,356.24) --
	(162.48,355.81) --
	(175.66,355.81) --
	(175.66,355.33) --
	(188.83,355.33) --
	(188.83,354.81) --
	(202.01,354.81) --
	(202.01,354.25) --
	(215.19,354.25) --
	(215.19,353.65) --
	(228.36,353.65) --
	(228.36,353.02) --
	(241.54,353.02) --
	(241.54,352.34) --
	(254.72,352.34) --
	(254.72,351.63) --
	(267.89,351.63) --
	(267.89,350.89) --
	(281.07,350.89) --
	(281.07,350.10) --
	(294.25,350.10) --
	(294.25,349.29);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,358.08) --
	( 57.06,358.08) --
	( 57.06,357.44) --
	( 70.24,357.44) --
	( 70.24,356.69) --
	( 83.42,356.69) --
	( 83.42,355.82) --
	( 96.59,355.82) --
	( 96.59,354.92) --
	(109.77,354.92) --
	(109.77,353.99) --
	(122.95,353.99) --
	(122.95,353.03) --
	(136.12,353.03) --
	(136.12,352.09) --
	(149.30,352.09) --
	(149.30,351.11) --
	(162.48,351.11) --
	(162.48,350.15) --
	(175.66,350.15) --
	(175.66,349.17) --
	(188.83,349.17) --
	(188.83,348.21) --
	(202.01,348.21) --
	(202.01,347.23) --
	(215.19,347.23) --
	(215.19,346.26) --
	(228.36,346.26) --
	(228.36,345.30) --
	(241.54,345.30) --
	(241.54,344.36) --
	(254.72,344.36) --
	(254.72,343.43) --
	(267.89,343.43) --
	(267.89,342.49) --
	(281.07,342.49) --
	(281.07,341.58) --
	(294.25,341.58) --
	(294.25,340.62);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,358.08) --
	( 57.06,358.08) --
	( 57.06,357.61) --
	( 70.24,357.61) --
	( 70.24,357.04) --
	( 83.42,357.04) --
	( 83.42,356.38) --
	( 96.59,356.38) --
	( 96.59,355.69) --
	(109.77,355.69) --
	(109.77,354.96) --
	(122.95,354.96) --
	(122.95,354.21) --
	(136.12,354.21) --
	(136.12,353.47) --
	(149.30,353.47) --
	(149.30,352.71) --
	(162.48,352.71) --
	(162.48,351.93) --
	(175.66,351.93) --
	(175.66,351.15) --
	(188.83,351.15) --
	(188.83,350.37) --
	(202.01,350.37) --
	(202.01,349.60) --
	(215.19,349.60) --
	(215.19,348.82) --
	(228.36,348.82) --
	(228.36,348.05) --
	(241.54,348.05) --
	(241.54,347.29) --
	(254.72,347.29) --
	(254.72,346.53) --
	(267.89,346.53) --
	(267.89,345.77) --
	(281.07,345.77) --
	(281.07,345.02) --
	(294.25,345.02) --
	(294.25,344.24);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,271.51) rectangle (306.77,362.20);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,175.32) rectangle (306.77,266.01);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,175.32) rectangle (306.77,266.01);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,194.61) --
	(306.77,194.61);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,221.52) --
	(306.77,221.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,248.43) --
	(306.77,248.43);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,175.32) --
	( 63.65,266.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,175.32) --
	(129.54,266.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,175.32) --
	(195.42,266.01);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,175.32) --
	(261.31,266.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,181.15) --
	(306.77,181.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,208.07) --
	(306.77,208.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,234.98) --
	(306.77,234.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,261.89) --
	(306.77,261.89);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,175.32) --
	( 96.59,266.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,175.32) --
	(162.48,266.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,175.32) --
	(228.36,266.01);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,175.32) --
	(294.25,266.01);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,261.89) --
	( 57.06,261.89) --
	( 57.06,261.89) --
	( 70.24,261.89) --
	( 70.24,261.89) --
	( 83.42,261.89) --
	( 83.42,259.08) --
	( 96.59,259.08) --
	( 96.59,255.06) --
	(109.77,255.06) --
	(109.77,250.42) --
	(122.95,250.42) --
	(122.95,245.48) --
	(136.12,245.48) --
	(136.12,240.64) --
	(149.30,240.64) --
	(149.30,235.81) --
	(162.48,235.81) --
	(162.48,231.06) --
	(175.66,231.06) --
	(175.66,226.27) --
	(188.83,226.27) --
	(188.83,221.82) --
	(202.01,221.82) --
	(202.01,217.62) --
	(215.19,217.62) --
	(215.19,213.63) --
	(228.36,213.63) --
	(228.36,209.63) --
	(241.54,209.63) --
	(241.54,205.76) --
	(254.72,205.76) --
	(254.72,202.12) --
	(267.89,202.12) --
	(267.89,198.54) --
	(281.07,198.54) --
	(281.07,195.21) --
	(294.25,195.21) --
	(294.25,183.05);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,261.89) --
	( 57.06,261.89) --
	( 57.06,261.89) --
	( 70.24,261.89) --
	( 70.24,261.89) --
	( 83.42,261.89) --
	( 83.42,259.68) --
	( 96.59,259.68) --
	( 96.59,256.75) --
	(109.77,256.75) --
	(109.77,253.43) --
	(122.95,253.43) --
	(122.95,249.51) --
	(136.12,249.51) --
	(136.12,245.67) --
	(149.30,245.67) --
	(149.30,241.87) --
	(162.48,241.87) --
	(162.48,238.00) --
	(175.66,238.00) --
	(175.66,234.26) --
	(188.83,234.26) --
	(188.83,230.51) --
	(202.01,230.51) --
	(202.01,226.73) --
	(215.19,226.73) --
	(215.19,223.17) --
	(228.36,223.17) --
	(228.36,219.80) --
	(241.54,219.80) --
	(241.54,216.27) --
	(254.72,216.27) --
	(254.72,212.95) --
	(267.89,212.95) --
	(267.89,209.79) --
	(281.07,209.79) --
	(281.07,206.54) --
	(294.25,206.54) --
	(294.25,195.49);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,261.89) --
	( 57.06,261.89) --
	( 57.06,261.67) --
	( 70.24,261.67) --
	( 70.24,261.33) --
	( 83.42,261.33) --
	( 83.42,260.85) --
	( 96.59,260.85) --
	( 96.59,260.25) --
	(109.77,260.25) --
	(109.77,259.53) --
	(122.95,259.53) --
	(122.95,258.68) --
	(136.12,258.68) --
	(136.12,257.72) --
	(149.30,257.72) --
	(149.30,256.64) --
	(162.48,256.64) --
	(162.48,255.46) --
	(175.66,255.46) --
	(175.66,254.15) --
	(188.83,254.15) --
	(188.83,252.72) --
	(202.01,252.72) --
	(202.01,251.20) --
	(215.19,251.20) --
	(215.19,249.60) --
	(228.36,249.60) --
	(228.36,247.86) --
	(241.54,247.86) --
	(241.54,246.02) --
	(254.72,246.02) --
	(254.72,244.09) --
	(267.89,244.09) --
	(267.89,242.17) --
	(281.07,242.17) --
	(281.07,240.14) --
	(294.25,240.14) --
	(294.25,238.16);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,261.89) --
	( 57.06,261.89) --
	( 57.06,261.74) --
	( 70.24,261.74) --
	( 70.24,261.48) --
	( 83.42,261.48) --
	( 83.42,261.11) --
	( 96.59,261.11) --
	( 96.59,260.64) --
	(109.77,260.64) --
	(109.77,260.04) --
	(122.95,260.04) --
	(122.95,259.34) --
	(136.12,259.34) --
	(136.12,258.54) --
	(149.30,258.54) --
	(149.30,257.65) --
	(162.48,257.65) --
	(162.48,256.66) --
	(175.66,256.66) --
	(175.66,255.54) --
	(188.83,255.54) --
	(188.83,254.35) --
	(202.01,254.35) --
	(202.01,253.08) --
	(215.19,253.08) --
	(215.19,251.72) --
	(228.36,251.72) --
	(228.36,250.28) --
	(241.54,250.28) --
	(241.54,248.74) --
	(254.72,248.74) --
	(254.72,247.17) --
	(267.89,247.17) --
	(267.89,245.51) --
	(281.07,245.51) --
	(281.07,243.82) --
	(294.25,243.82) --
	(294.25,242.04);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,261.89) --
	( 57.06,261.89) --
	( 57.06,260.44) --
	( 70.24,260.44) --
	( 70.24,258.73) --
	( 83.42,258.73) --
	( 83.42,256.84) --
	( 96.59,256.84) --
	( 96.59,254.87) --
	(109.77,254.87) --
	(109.77,252.86) --
	(122.95,252.86) --
	(122.95,250.78) --
	(136.12,250.78) --
	(136.12,248.73) --
	(149.30,248.73) --
	(149.30,246.66) --
	(162.48,246.66) --
	(162.48,244.62) --
	(175.66,244.62) --
	(175.66,242.57) --
	(188.83,242.57) --
	(188.83,240.51) --
	(202.01,240.51) --
	(202.01,238.51) --
	(215.19,238.51) --
	(215.19,236.56) --
	(228.36,236.56) --
	(228.36,234.58) --
	(241.54,234.58) --
	(241.54,232.63) --
	(254.72,232.63) --
	(254.72,230.71) --
	(267.89,230.71) --
	(267.89,228.91) --
	(281.07,228.91) --
	(281.07,227.12) --
	(294.25,227.12) --
	(294.25,225.36);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,261.89) --
	( 57.06,261.89) --
	( 57.06,260.89) --
	( 70.24,260.89) --
	( 70.24,259.63) --
	( 83.42,259.63) --
	( 83.42,258.16) --
	( 96.59,258.16) --
	( 96.59,256.59) --
	(109.77,256.59) --
	(109.77,254.90) --
	(122.95,254.90) --
	(122.95,253.19) --
	(136.12,253.19) --
	(136.12,251.46) --
	(149.30,251.46) --
	(149.30,249.74) --
	(162.48,249.74) --
	(162.48,248.00) --
	(175.66,248.00) --
	(175.66,246.22) --
	(188.83,246.22) --
	(188.83,244.49) --
	(202.01,244.49) --
	(202.01,242.77) --
	(215.19,242.77) --
	(215.19,241.08) --
	(228.36,241.08) --
	(228.36,239.40) --
	(241.54,239.40) --
	(241.54,237.74) --
	(254.72,237.74) --
	(254.72,236.14) --
	(267.89,236.14) --
	(267.89,234.55) --
	(281.07,234.55) --
	(281.07,233.01) --
	(294.25,233.01) --
	(294.25,231.39);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,175.32) rectangle (306.77,266.01);
\end{scope}
\begin{scope}
\path[clip] ( 31.37, 79.12) rectangle (306.77,169.82);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37, 79.12) rectangle (306.77,169.82);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37, 98.42) --
	(306.77, 98.42);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,125.33) --
	(306.77,125.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,152.24) --
	(306.77,152.24);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65, 79.12) --
	( 63.65,169.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54, 79.12) --
	(129.54,169.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42, 79.12) --
	(195.42,169.82);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31, 79.12) --
	(261.31,169.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 84.96) --
	(306.77, 84.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,111.87) --
	(306.77,111.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,138.78) --
	(306.77,138.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,165.69) --
	(306.77,165.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 79.12) --
	( 96.59,169.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 79.12) --
	(162.48,169.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 79.12) --
	(228.36,169.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 79.12) --
	(294.25,169.82);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,165.69) --
	( 57.06,165.69) --
	( 57.06,165.69) --
	( 70.24,165.69) --
	( 70.24,165.69) --
	( 83.42,165.69) --
	( 83.42,161.51) --
	( 96.59,161.51) --
	( 96.59,155.98) --
	(109.77,155.98) --
	(109.77,150.36) --
	(122.95,150.36) --
	(122.95,145.14) --
	(136.12,145.14) --
	(136.12,139.89) --
	(149.30,139.89) --
	(149.30,135.11) --
	(162.48,135.11) --
	(162.48,130.39) --
	(175.66,130.39) --
	(175.66,125.41) --
	(188.83,125.41) --
	(188.83,120.90) --
	(202.01,120.90) --
	(202.01,116.83) --
	(215.19,116.83) --
	(215.19,112.86) --
	(228.36,112.86) --
	(228.36,109.19) --
	(241.54,109.19) --
	(241.54,105.56) --
	(254.72,105.56) --
	(254.72,101.91) --
	(267.89,101.91) --
	(267.89, 98.28) --
	(281.07, 98.28) --
	(281.07, 95.14) --
	(294.25, 95.14) --
	(294.25, 83.24);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,165.69) --
	( 57.06,165.69) --
	( 57.06,165.69) --
	( 70.24,165.69) --
	( 70.24,165.69) --
	( 83.42,165.69) --
	( 83.42,162.40) --
	( 96.59,162.40) --
	( 96.59,158.79) --
	(109.77,158.79) --
	(109.77,154.59) --
	(122.95,154.59) --
	(122.95,150.52) --
	(136.12,150.52) --
	(136.12,146.42) --
	(149.30,146.42) --
	(149.30,142.23) --
	(162.48,142.23) --
	(162.48,138.32) --
	(175.66,138.32) --
	(175.66,134.49) --
	(188.83,134.49) --
	(188.83,130.84) --
	(202.01,130.84) --
	(202.01,127.10) --
	(215.19,127.10) --
	(215.19,123.32) --
	(228.36,123.32) --
	(228.36,119.92) --
	(241.54,119.92) --
	(241.54,116.59) --
	(254.72,116.59) --
	(254.72,113.35) --
	(267.89,113.35) --
	(267.89,110.31) --
	(281.07,110.31) --
	(281.07,107.44) --
	(294.25,107.44) --
	(294.25, 96.23);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,165.69) --
	( 57.06,165.69) --
	( 57.06,165.35) --
	( 70.24,165.35) --
	( 70.24,164.83) --
	( 83.42,164.83) --
	( 83.42,164.11) --
	( 96.59,164.11) --
	( 96.59,163.21) --
	(109.77,163.21) --
	(109.77,162.12) --
	(122.95,162.12) --
	(122.95,160.90) --
	(136.12,160.90) --
	(136.12,159.52) --
	(149.30,159.52) --
	(149.30,157.94) --
	(162.48,157.94) --
	(162.48,156.22) --
	(175.66,156.22) --
	(175.66,154.34) --
	(188.83,154.34) --
	(188.83,152.30) --
	(202.01,152.30) --
	(202.01,150.10) --
	(215.19,150.10) --
	(215.19,147.78) --
	(228.36,147.78) --
	(228.36,145.35) --
	(241.54,145.35) --
	(241.54,142.80) --
	(254.72,142.80) --
	(254.72,140.11) --
	(267.89,140.11) --
	(267.89,137.29) --
	(281.07,137.29) --
	(281.07,134.36) --
	(294.25,134.36) --
	(294.25,131.32);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,165.69) --
	( 57.06,165.69) --
	( 57.06,165.44) --
	( 70.24,165.44) --
	( 70.24,165.03) --
	( 83.42,165.03) --
	( 83.42,164.45) --
	( 96.59,164.45) --
	( 96.59,163.74) --
	(109.77,163.74) --
	(109.77,162.88) --
	(122.95,162.88) --
	(122.95,161.88) --
	(136.12,161.88) --
	(136.12,160.73) --
	(149.30,160.73) --
	(149.30,159.45) --
	(162.48,159.45) --
	(162.48,158.02) --
	(175.66,158.02) --
	(175.66,156.45) --
	(188.83,156.45) --
	(188.83,154.75) --
	(202.01,154.75) --
	(202.01,152.92) --
	(215.19,152.92) --
	(215.19,150.98) --
	(228.36,150.98) --
	(228.36,148.91) --
	(241.54,148.91) --
	(241.54,146.73) --
	(254.72,146.73) --
	(254.72,144.45) --
	(267.89,144.45) --
	(267.89,142.05) --
	(281.07,142.05) --
	(281.07,139.57) --
	(294.25,139.57) --
	(294.25,137.02);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,165.69) --
	( 57.06,165.69) --
	( 57.06,163.41) --
	( 70.24,163.41) --
	( 70.24,160.81) --
	( 83.42,160.81) --
	( 83.42,157.93) --
	( 96.59,157.93) --
	( 96.59,154.99) --
	(109.77,154.99) --
	(109.77,151.91) --
	(122.95,151.91) --
	(122.95,148.94) --
	(136.12,148.94) --
	(136.12,145.97) --
	(149.30,145.97) --
	(149.30,142.91) --
	(162.48,142.91) --
	(162.48,139.94) --
	(175.66,139.94) --
	(175.66,136.96) --
	(188.83,136.96) --
	(188.83,134.01) --
	(202.01,134.01) --
	(202.01,131.07) --
	(215.19,131.07) --
	(215.19,128.21) --
	(228.36,128.21) --
	(228.36,125.40) --
	(241.54,125.40) --
	(241.54,122.65) --
	(254.72,122.65) --
	(254.72,119.93) --
	(267.89,119.93) --
	(267.89,117.23) --
	(281.07,117.23) --
	(281.07,114.59) --
	(294.25,114.59) --
	(294.25,111.85);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,165.69) --
	( 57.06,165.69) --
	( 57.06,163.98) --
	( 70.24,163.98) --
	( 70.24,161.94) --
	( 83.42,161.94) --
	( 83.42,159.66) --
	( 96.59,159.66) --
	( 96.59,157.30) --
	(109.77,157.30) --
	(109.77,154.88) --
	(122.95,154.88) --
	(122.95,152.41) --
	(136.12,152.41) --
	(136.12,149.93) --
	(149.30,149.93) --
	(149.30,147.44) --
	(162.48,147.44) --
	(162.48,144.92) --
	(175.66,144.92) --
	(175.66,142.41) --
	(188.83,142.41) --
	(188.83,139.93) --
	(202.01,139.93) --
	(202.01,137.45) --
	(215.19,137.45) --
	(215.19,135.00) --
	(228.36,135.00) --
	(228.36,132.58) --
	(241.54,132.58) --
	(241.54,130.19) --
	(254.72,130.19) --
	(254.72,127.83) --
	(267.89,127.83) --
	(267.89,125.50) --
	(281.07,125.50) --
	(281.07,123.22) --
	(294.25,123.22) --
	(294.25,120.87);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37, 79.12) rectangle (306.77,169.82);
\end{scope}
\begin{scope}
\path[clip] (306.77,463.90) rectangle (319.71,554.59);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,463.90) rectangle (319.71,554.59);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,509.25) {Scenario 1};
\end{scope}
\begin{scope}
\path[clip] (306.77,367.70) rectangle (319.71,458.40);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,367.70) rectangle (319.71,458.40);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,413.05) {Scenario 2};
\end{scope}
\begin{scope}
\path[clip] (306.77,271.51) rectangle (319.71,362.20);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,271.51) rectangle (319.71,362.20);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,316.86) {Scenario 3};
\end{scope}
\begin{scope}
\path[clip] (306.77,175.32) rectangle (319.71,266.01);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,175.32) rectangle (319.71,266.01);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,220.66) {Scenario 4};
\end{scope}
\begin{scope}
\path[clip] (306.77, 79.12) rectangle (319.71,169.82);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77, 79.12) rectangle (319.71,169.82);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,124.47) {Scenario 5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 76.37) --
	( 96.59, 79.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 76.37) --
	(162.48, 79.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 76.37) --
	(228.36, 79.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 76.37) --
	(294.25, 79.12);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 96.59, 68.66) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (162.48, 68.66) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (228.36, 68.66) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (294.25, 68.66) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,466.98) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,493.89) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,520.80) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,547.72) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,469.74) --
	( 31.37,469.74);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,496.65) --
	( 31.37,496.65);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,523.56) --
	( 31.37,523.56);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,550.47) --
	( 31.37,550.47);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,370.79) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,397.70) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,424.61) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,451.52) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,373.54) --
	( 31.37,373.54);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,400.45) --
	( 31.37,400.45);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,427.36) --
	( 31.37,427.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,454.28) --
	( 31.37,454.28);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,274.59) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,301.50) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,328.42) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,355.33) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,277.35) --
	( 31.37,277.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,304.26) --
	( 31.37,304.26);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,331.17) --
	( 31.37,331.17);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,358.08) --
	( 31.37,358.08);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,178.40) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,205.31) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,232.22) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,259.13) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,181.15) --
	( 31.37,181.15);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,208.07) --
	( 31.37,208.07);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,234.98) --
	( 31.37,234.98);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,261.89) --
	( 31.37,261.89);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 82.21) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,109.12) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,136.03) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,162.94) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 84.96) --
	( 31.37, 84.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,111.87) --
	( 31.37,111.87);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,138.78) --
	( 31.37,138.78);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,165.69) --
	( 31.37,165.69);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (169.07, 58.16) {Years since hire};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,316.86) {Cancer-free survival};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 55.91, 30.95) rectangle (282.22, 45.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 61.41, 30.95) rectangle ( 75.86, 45.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.86, 38.18) -- ( 74.42, 38.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (205.46, 30.95) rectangle (219.91, 45.41);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] (206.91, 38.18) -- (218.47, 38.18);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 81.36, 35.08) {Always exposed while at work};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (225.41, 35.08) {Never exposed};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 92.13,  5.50) rectangle (246.00, 19.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 97.63,  5.50) rectangle (112.09, 19.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.08, 12.73) -- (110.64, 12.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (145.86,  5.50) rectangle (160.31, 19.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (147.30, 12.73) -- (158.87, 12.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (195.80,  5.50) rectangle (210.26, 19.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (197.25, 12.73) -- (208.81, 12.73);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (117.59,  9.63) {Truth};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (165.81,  9.63) {WKM};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,560.09);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (215.76,  9.63) {AWKM};
\end{scope}
\end{tikzpicture}

\end{document}