### 2022-03-04 reports/resources/sensitivity/results.tex

% Created by tikzDevice version 0.12.3.1 on 2022-01-19 11:19:01
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,542.02);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (325.21,542.02);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,446.55) rectangle (306.77,536.52);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,446.55) rectangle (306.77,536.52);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,466.60) --
	(306.77,466.60);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,492.93) --
	(306.77,492.93);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,519.27) --
	(306.77,519.27);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,446.55) --
	( 63.65,536.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,446.55) --
	(129.54,536.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,446.55) --
	(195.42,536.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,446.55) --
	(261.31,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,453.43) --
	(306.77,453.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,479.77) --
	(306.77,479.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,506.10) --
	(306.77,506.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,532.44) --
	(306.77,532.44);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,446.55) --
	( 96.59,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,446.55) --
	(162.48,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,446.55) --
	(228.36,536.52);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,446.55) --
	(294.25,536.52);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,532.44) --
	( 70.24,532.44) --
	( 70.24,532.44) --
	( 83.42,532.44) --
	( 83.42,531.33) --
	( 96.59,531.33) --
	( 96.59,530.06) --
	(109.77,530.06) --
	(109.77,528.69) --
	(122.95,528.69) --
	(122.95,527.08) --
	(136.12,527.08) --
	(136.12,525.39) --
	(149.30,525.39) --
	(149.30,523.81) --
	(162.48,523.81) --
	(162.48,522.05) --
	(175.66,522.05) --
	(175.66,520.52) --
	(188.83,520.52) --
	(188.83,518.91) --
	(202.01,518.91) --
	(202.01,517.21) --
	(215.19,517.21) --
	(215.19,515.61) --
	(228.36,515.61) --
	(228.36,514.24) --
	(241.54,514.24) --
	(241.54,512.91) --
	(254.72,512.91) --
	(254.72,511.38) --
	(267.89,511.38) --
	(267.89,509.61) --
	(281.07,509.61) --
	(281.07,508.11) --
	(294.25,508.11) --
	(294.25,502.32);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,532.44) --
	( 70.24,532.44) --
	( 70.24,532.44) --
	( 83.42,532.44) --
	( 83.42,530.87) --
	( 96.59,530.87) --
	( 96.59,528.70) --
	(109.77,528.70) --
	(109.77,526.39) --
	(122.95,526.39) --
	(122.95,524.01) --
	(136.12,524.01) --
	(136.12,521.83) --
	(149.30,521.83) --
	(149.30,519.56) --
	(162.48,519.56) --
	(162.48,517.37) --
	(175.66,517.37) --
	(175.66,515.30) --
	(188.83,515.30) --
	(188.83,513.21) --
	(202.01,513.21) --
	(202.01,511.26) --
	(215.19,511.26) --
	(215.19,509.28) --
	(228.36,509.28) --
	(228.36,507.48) --
	(241.54,507.48) --
	(241.54,505.65) --
	(254.72,505.65) --
	(254.72,503.82) --
	(267.89,503.82) --
	(267.89,502.08) --
	(281.07,502.08) --
	(281.07,500.50) --
	(294.25,500.50) --
	(294.25,494.02);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,531.97) --
	( 70.24,531.97) --
	( 70.24,531.46) --
	( 83.42,531.46) --
	( 83.42,530.89) --
	( 96.59,530.89) --
	( 96.59,530.26) --
	(109.77,530.26) --
	(109.77,529.57) --
	(122.95,529.57) --
	(122.95,528.86) --
	(136.12,528.86) --
	(136.12,528.08) --
	(149.30,528.08) --
	(149.30,527.26) --
	(162.48,527.26) --
	(162.48,526.43) --
	(175.66,526.43) --
	(175.66,525.58) --
	(188.83,525.58) --
	(188.83,524.70) --
	(202.01,524.70) --
	(202.01,523.81) --
	(215.19,523.81) --
	(215.19,522.90) --
	(228.36,522.90) --
	(228.36,521.97) --
	(241.54,521.97) --
	(241.54,521.04) --
	(254.72,521.04) --
	(254.72,520.13) --
	(267.89,520.13) --
	(267.89,519.20) --
	(281.07,519.20) --
	(281.07,518.27) --
	(294.25,518.27) --
	(294.25,517.22);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,531.83) --
	( 70.24,531.83) --
	( 70.24,531.17) --
	( 83.42,531.17) --
	( 83.42,530.44) --
	( 96.59,530.44) --
	( 96.59,529.64) --
	(109.77,529.64) --
	(109.77,528.78) --
	(122.95,528.78) --
	(122.95,527.89) --
	(136.12,527.89) --
	(136.12,526.93) --
	(149.30,526.93) --
	(149.30,525.92) --
	(162.48,525.92) --
	(162.48,524.89) --
	(175.66,524.89) --
	(175.66,523.85) --
	(188.83,523.85) --
	(188.83,522.78) --
	(202.01,522.78) --
	(202.01,521.69) --
	(215.19,521.69) --
	(215.19,520.57) --
	(228.36,520.57) --
	(228.36,519.46) --
	(241.54,519.46) --
	(241.54,518.35) --
	(254.72,518.35) --
	(254.72,517.24) --
	(267.89,517.24) --
	(267.89,516.10) --
	(281.07,516.10) --
	(281.07,515.00) --
	(294.25,515.00) --
	(294.25,513.74);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,531.06) --
	( 70.24,531.06) --
	( 70.24,529.66) --
	( 83.42,529.66) --
	( 83.42,528.18) --
	( 96.59,528.18) --
	( 96.59,526.67) --
	(109.77,526.67) --
	(109.77,525.07) --
	(122.95,525.07) --
	(122.95,523.54) --
	(136.12,523.54) --
	(136.12,521.93) --
	(149.30,521.93) --
	(149.30,520.31) --
	(162.48,520.31) --
	(162.48,518.72) --
	(175.66,518.72) --
	(175.66,517.11) --
	(188.83,517.11) --
	(188.83,515.48) --
	(202.01,515.48) --
	(202.01,513.92) --
	(215.19,513.92) --
	(215.19,512.32) --
	(228.36,512.32) --
	(228.36,510.72) --
	(241.54,510.72) --
	(241.54,509.14) --
	(254.72,509.14) --
	(254.72,507.62) --
	(267.89,507.62) --
	(267.89,506.09) --
	(281.07,506.09) --
	(281.07,504.74) --
	(294.25,504.74) --
	(294.25,503.36);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,532.44) --
	( 57.06,532.44) --
	( 57.06,530.66) --
	( 70.24,530.66) --
	( 70.24,528.84) --
	( 83.42,528.84) --
	( 83.42,526.96) --
	( 96.59,526.96) --
	( 96.59,525.04) --
	(109.77,525.04) --
	(109.77,523.05) --
	(122.95,523.05) --
	(122.95,521.14) --
	(136.12,521.14) --
	(136.12,519.17) --
	(149.30,519.17) --
	(149.30,517.17) --
	(162.48,517.17) --
	(162.48,515.19) --
	(175.66,515.19) --
	(175.66,513.24) --
	(188.83,513.24) --
	(188.83,511.28) --
	(202.01,511.28) --
	(202.01,509.36) --
	(215.19,509.36) --
	(215.19,507.43) --
	(228.36,507.43) --
	(228.36,505.52) --
	(241.54,505.52) --
	(241.54,503.65) --
	(254.72,503.65) --
	(254.72,501.82) --
	(267.89,501.82) --
	(267.89,499.96) --
	(281.07,499.96) --
	(281.07,498.38) --
	(294.25,498.38) --
	(294.25,496.75);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,446.55) rectangle (306.77,536.52);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,351.08) rectangle (306.77,441.05);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,351.08) rectangle (306.77,441.05);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,371.13) --
	(306.77,371.13);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,397.46) --
	(306.77,397.46);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,423.80) --
	(306.77,423.80);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,351.08) --
	( 63.65,441.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,351.08) --
	(129.54,441.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,351.08) --
	(195.42,441.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,351.08) --
	(261.31,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,357.96) --
	(306.77,357.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,384.30) --
	(306.77,384.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,410.63) --
	(306.77,410.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,436.96) --
	(306.77,436.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,351.08) --
	( 96.59,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,351.08) --
	(162.48,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,351.08) --
	(228.36,441.05);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,351.08) --
	(294.25,441.05);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.96) --
	( 70.24,436.96) --
	( 70.24,436.96) --
	( 83.42,436.96) --
	( 83.42,435.99) --
	( 96.59,435.99) --
	( 96.59,434.54) --
	(109.77,434.54) --
	(109.77,433.15) --
	(122.95,433.15) --
	(122.95,431.60) --
	(136.12,431.60) --
	(136.12,430.03) --
	(149.30,430.03) --
	(149.30,428.31) --
	(162.48,428.31) --
	(162.48,426.57) --
	(175.66,426.57) --
	(175.66,424.90) --
	(188.83,424.90) --
	(188.83,423.22) --
	(202.01,423.22) --
	(202.01,421.65) --
	(215.19,421.65) --
	(215.19,420.13) --
	(228.36,420.13) --
	(228.36,418.54) --
	(241.54,418.54) --
	(241.54,417.01) --
	(254.72,417.01) --
	(254.72,415.42) --
	(267.89,415.42) --
	(267.89,413.97) --
	(281.07,413.97) --
	(281.07,412.39) --
	(294.25,412.39) --
	(294.25,406.12);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.96) --
	( 70.24,436.96) --
	( 70.24,436.96) --
	( 83.42,436.96) --
	( 83.42,435.30) --
	( 96.59,435.30) --
	( 96.59,433.42) --
	(109.77,433.42) --
	(109.77,431.17) --
	(122.95,431.17) --
	(122.95,428.91) --
	(136.12,428.91) --
	(136.12,426.60) --
	(149.30,426.60) --
	(149.30,424.44) --
	(162.48,424.44) --
	(162.48,422.41) --
	(175.66,422.41) --
	(175.66,420.27) --
	(188.83,420.27) --
	(188.83,418.05) --
	(202.01,418.05) --
	(202.01,416.08) --
	(215.19,416.08) --
	(215.19,414.23) --
	(228.36,414.23) --
	(228.36,412.46) --
	(241.54,412.46) --
	(241.54,410.47) --
	(254.72,410.47) --
	(254.72,408.61) --
	(267.89,408.61) --
	(267.89,406.79) --
	(281.07,406.79) --
	(281.07,405.02) --
	(294.25,405.02) --
	(294.25,398.52);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.49) --
	( 70.24,436.49) --
	( 70.24,435.98) --
	( 83.42,435.98) --
	( 83.42,435.40) --
	( 96.59,435.40) --
	( 96.59,434.74) --
	(109.77,434.74) --
	(109.77,434.03) --
	(122.95,434.03) --
	(122.95,433.29) --
	(136.12,433.29) --
	(136.12,432.49) --
	(149.30,432.49) --
	(149.30,431.67) --
	(162.48,431.67) --
	(162.48,430.81) --
	(175.66,430.81) --
	(175.66,429.93) --
	(188.83,429.93) --
	(188.83,429.01) --
	(202.01,429.01) --
	(202.01,428.11) --
	(215.19,428.11) --
	(215.19,427.14) --
	(228.36,427.14) --
	(228.36,426.18) --
	(241.54,426.18) --
	(241.54,425.20) --
	(254.72,425.20) --
	(254.72,424.22) --
	(267.89,424.22) --
	(267.89,423.26) --
	(281.07,423.26) --
	(281.07,422.29) --
	(294.25,422.29) --
	(294.25,421.27);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,436.35) --
	( 70.24,436.35) --
	( 70.24,435.69) --
	( 83.42,435.69) --
	( 83.42,434.95) --
	( 96.59,434.95) --
	( 96.59,434.13) --
	(109.77,434.13) --
	(109.77,433.26) --
	(122.95,433.26) --
	(122.95,432.33) --
	(136.12,432.33) --
	(136.12,431.33) --
	(149.30,431.33) --
	(149.30,430.29) --
	(162.48,430.29) --
	(162.48,429.21) --
	(175.66,429.21) --
	(175.66,428.12) --
	(188.83,428.12) --
	(188.83,427.01) --
	(202.01,427.01) --
	(202.01,425.90) --
	(215.19,425.90) --
	(215.19,424.74) --
	(228.36,424.74) --
	(228.36,423.59) --
	(241.54,423.59) --
	(241.54,422.43) --
	(254.72,422.43) --
	(254.72,421.32) --
	(267.89,421.32) --
	(267.89,420.16) --
	(281.07,420.16) --
	(281.07,419.01) --
	(294.25,419.01) --
	(294.25,417.72);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,435.58) --
	( 70.24,435.58) --
	( 70.24,434.16) --
	( 83.42,434.16) --
	( 83.42,432.67) --
	( 96.59,432.67) --
	( 96.59,431.09) --
	(109.77,431.09) --
	(109.77,429.46) --
	(122.95,429.46) --
	(122.95,427.88) --
	(136.12,427.88) --
	(136.12,426.24) --
	(149.30,426.24) --
	(149.30,424.64) --
	(162.48,424.64) --
	(162.48,423.00) --
	(175.66,423.00) --
	(175.66,421.35) --
	(188.83,421.35) --
	(188.83,419.68) --
	(202.01,419.68) --
	(202.01,418.10) --
	(215.19,418.10) --
	(215.19,416.43) --
	(228.36,416.43) --
	(228.36,414.79) --
	(241.54,414.79) --
	(241.54,413.14) --
	(254.72,413.14) --
	(254.72,411.54) --
	(267.89,411.54) --
	(267.89,409.96) --
	(281.07,409.96) --
	(281.07,408.56) --
	(294.25,408.56) --
	(294.25,407.24);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,436.96) --
	( 57.06,436.96) --
	( 57.06,435.17) --
	( 70.24,435.17) --
	( 70.24,433.34) --
	( 83.42,433.34) --
	( 83.42,431.44) --
	( 96.59,431.44) --
	( 96.59,429.47) --
	(109.77,429.47) --
	(109.77,427.46) --
	(122.95,427.46) --
	(122.95,425.48) --
	(136.12,425.48) --
	(136.12,423.46) --
	(149.30,423.46) --
	(149.30,421.43) --
	(162.48,421.43) --
	(162.48,419.38) --
	(175.66,419.38) --
	(175.66,417.35) --
	(188.83,417.35) --
	(188.83,415.36) --
	(202.01,415.36) --
	(202.01,413.43) --
	(215.19,413.43) --
	(215.19,411.45) --
	(228.36,411.45) --
	(228.36,409.51) --
	(241.54,409.51) --
	(241.54,407.57) --
	(254.72,407.57) --
	(254.72,405.77) --
	(267.89,405.77) --
	(267.89,403.91) --
	(281.07,403.91) --
	(281.07,402.27) --
	(294.25,402.27) --
	(294.25,400.62);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,351.08) rectangle (306.77,441.05);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,255.61) rectangle (306.77,345.58);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,255.61) rectangle (306.77,345.58);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,275.66) --
	(306.77,275.66);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,301.99) --
	(306.77,301.99);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,328.33) --
	(306.77,328.33);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,255.61) --
	( 63.65,345.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,255.61) --
	(129.54,345.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,255.61) --
	(195.42,345.58);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,255.61) --
	(261.31,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,262.49) --
	(306.77,262.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,288.82) --
	(306.77,288.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,315.16) --
	(306.77,315.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,341.49) --
	(306.77,341.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,255.61) --
	( 96.59,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,255.61) --
	(162.48,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,255.61) --
	(228.36,345.58);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,255.61) --
	(294.25,345.58);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,341.49) --
	( 70.24,341.49) --
	( 70.24,341.49) --
	( 83.42,341.49) --
	( 83.42,340.32) --
	( 96.59,340.32) --
	( 96.59,339.14) --
	(109.77,339.14) --
	(109.77,337.75) --
	(122.95,337.75) --
	(122.95,336.29) --
	(136.12,336.29) --
	(136.12,334.60) --
	(149.30,334.60) --
	(149.30,333.00) --
	(162.48,333.00) --
	(162.48,331.43) --
	(175.66,331.43) --
	(175.66,329.77) --
	(188.83,329.77) --
	(188.83,328.09) --
	(202.01,328.09) --
	(202.01,326.46) --
	(215.19,326.46) --
	(215.19,324.88) --
	(228.36,324.88) --
	(228.36,323.40) --
	(241.54,323.40) --
	(241.54,321.86) --
	(254.72,321.86) --
	(254.72,320.41) --
	(267.89,320.41) --
	(267.89,318.82) --
	(281.07,318.82) --
	(281.07,317.38) --
	(294.25,317.38) --
	(294.25,311.45);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,341.49) --
	( 70.24,341.49) --
	( 70.24,341.49) --
	( 83.42,341.49) --
	( 83.42,339.84) --
	( 96.59,339.84) --
	( 96.59,337.72) --
	(109.77,337.72) --
	(109.77,335.45) --
	(122.95,335.45) --
	(122.95,333.29) --
	(136.12,333.29) --
	(136.12,331.07) --
	(149.30,331.07) --
	(149.30,328.94) --
	(162.48,328.94) --
	(162.48,326.63) --
	(175.66,326.63) --
	(175.66,324.55) --
	(188.83,324.55) --
	(188.83,322.42) --
	(202.01,322.42) --
	(202.01,320.51) --
	(215.19,320.51) --
	(215.19,318.52) --
	(228.36,318.52) --
	(228.36,316.54) --
	(241.54,316.54) --
	(241.54,314.60) --
	(254.72,314.60) --
	(254.72,312.76) --
	(267.89,312.76) --
	(267.89,310.77) --
	(281.07,310.77) --
	(281.07,309.01) --
	(294.25,309.01) --
	(294.25,302.19);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,341.04) --
	( 70.24,341.04) --
	( 70.24,340.53) --
	( 83.42,340.53) --
	( 83.42,339.96) --
	( 96.59,339.96) --
	( 96.59,339.34) --
	(109.77,339.34) --
	(109.77,338.65) --
	(122.95,338.65) --
	(122.95,337.93) --
	(136.12,337.93) --
	(136.12,337.15) --
	(149.30,337.15) --
	(149.30,336.35) --
	(162.48,336.35) --
	(162.48,335.52) --
	(175.66,335.52) --
	(175.66,334.66) --
	(188.83,334.66) --
	(188.83,333.79) --
	(202.01,333.79) --
	(202.01,332.90) --
	(215.19,332.90) --
	(215.19,332.01) --
	(228.36,332.01) --
	(228.36,331.09) --
	(241.54,331.09) --
	(241.54,330.18) --
	(254.72,330.18) --
	(254.72,329.26) --
	(267.89,329.26) --
	(267.89,328.30) --
	(281.07,328.30) --
	(281.07,327.37) --
	(294.25,327.37) --
	(294.25,326.33);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,340.87) --
	( 70.24,340.87) --
	( 70.24,340.16) --
	( 83.42,340.16) --
	( 83.42,339.43) --
	( 96.59,339.43) --
	( 96.59,338.62) --
	(109.77,338.62) --
	(109.77,337.75) --
	(122.95,337.75) --
	(122.95,336.84) --
	(136.12,336.84) --
	(136.12,335.86) --
	(149.30,335.86) --
	(149.30,334.83) --
	(162.48,334.83) --
	(162.48,333.81) --
	(175.66,333.81) --
	(175.66,332.76) --
	(188.83,332.76) --
	(188.83,331.67) --
	(202.01,331.67) --
	(202.01,330.55) --
	(215.19,330.55) --
	(215.19,329.40) --
	(228.36,329.40) --
	(228.36,328.22) --
	(241.54,328.22) --
	(241.54,327.10) --
	(254.72,327.10) --
	(254.72,325.95) --
	(267.89,325.95) --
	(267.89,324.77) --
	(281.07,324.77) --
	(281.07,323.67) --
	(294.25,323.67) --
	(294.25,322.38);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,340.16) --
	( 70.24,340.16) --
	( 70.24,338.75) --
	( 83.42,338.75) --
	( 83.42,337.29) --
	( 96.59,337.29) --
	( 96.59,335.78) --
	(109.77,335.78) --
	(109.77,334.19) --
	(122.95,334.19) --
	(122.95,332.65) --
	(136.12,332.65) --
	(136.12,331.05) --
	(149.30,331.05) --
	(149.30,329.45) --
	(162.48,329.45) --
	(162.48,327.86) --
	(175.66,327.86) --
	(175.66,326.23) --
	(188.83,326.23) --
	(188.83,324.62) --
	(202.01,324.62) --
	(202.01,323.06) --
	(215.19,323.06) --
	(215.19,321.50) --
	(228.36,321.50) --
	(228.36,319.93) --
	(241.54,319.93) --
	(241.54,318.36) --
	(254.72,318.36) --
	(254.72,316.84) --
	(267.89,316.84) --
	(267.89,315.25) --
	(281.07,315.25) --
	(281.07,313.90) --
	(294.25,313.90) --
	(294.25,312.54);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,341.49) --
	( 57.06,341.49) --
	( 57.06,339.67) --
	( 70.24,339.67) --
	( 70.24,337.70) --
	( 83.42,337.70) --
	( 83.42,335.81) --
	( 96.59,335.81) --
	( 96.59,333.86) --
	(109.77,333.86) --
	(109.77,331.84) --
	(122.95,331.84) --
	(122.95,329.88) --
	(136.12,329.88) --
	(136.12,327.89) --
	(149.30,327.89) --
	(149.30,325.84) --
	(162.48,325.84) --
	(162.48,323.89) --
	(175.66,323.89) --
	(175.66,321.90) --
	(188.83,321.90) --
	(188.83,319.92) --
	(202.01,319.92) --
	(202.01,317.95) --
	(215.19,317.95) --
	(215.19,315.96) --
	(228.36,315.96) --
	(228.36,313.96) --
	(241.54,313.96) --
	(241.54,312.05) --
	(254.72,312.05) --
	(254.72,310.18) --
	(267.89,310.18) --
	(267.89,308.25) --
	(281.07,308.25) --
	(281.07,306.67) --
	(294.25,306.67) --
	(294.25,305.01);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,255.61) rectangle (306.77,345.58);
\end{scope}
\begin{scope}
\path[clip] ( 31.37,160.14) rectangle (306.77,250.11);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37,160.14) rectangle (306.77,250.11);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,180.19) --
	(306.77,180.19);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,206.52) --
	(306.77,206.52);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,232.85) --
	(306.77,232.85);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65,160.14) --
	( 63.65,250.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54,160.14) --
	(129.54,250.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42,160.14) --
	(195.42,250.11);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31,160.14) --
	(261.31,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,167.02) --
	(306.77,167.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,193.35) --
	(306.77,193.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,219.69) --
	(306.77,219.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,246.02) --
	(306.77,246.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59,160.14) --
	( 96.59,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48,160.14) --
	(162.48,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36,160.14) --
	(228.36,250.11);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25,160.14) --
	(294.25,250.11);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,246.02) --
	( 70.24,246.02) --
	( 70.24,246.02) --
	( 83.42,246.02) --
	( 83.42,243.99) --
	( 96.59,243.99) --
	( 96.59,241.05) --
	(109.77,241.05) --
	(109.77,237.70) --
	(122.95,237.70) --
	(122.95,233.72) --
	(136.12,233.72) --
	(136.12,229.94) --
	(149.30,229.94) --
	(149.30,225.93) --
	(162.48,225.93) --
	(162.48,221.98) --
	(175.66,221.98) --
	(175.66,218.17) --
	(188.83,218.17) --
	(188.83,214.60) --
	(202.01,214.60) --
	(202.01,211.01) --
	(215.19,211.01) --
	(215.19,207.49) --
	(228.36,207.49) --
	(228.36,204.27) --
	(241.54,204.27) --
	(241.54,201.08) --
	(254.72,201.08) --
	(254.72,197.91) --
	(267.89,197.91) --
	(267.89,194.81) --
	(281.07,194.81) --
	(281.07,191.65) --
	(294.25,191.65) --
	(294.25,180.29);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,246.02) --
	( 70.24,246.02) --
	( 70.24,246.02) --
	( 83.42,246.02) --
	( 83.42,243.12) --
	( 96.59,243.12) --
	( 96.59,238.79) --
	(109.77,238.79) --
	(109.77,234.27) --
	(122.95,234.27) --
	(122.95,229.56) --
	(136.12,229.56) --
	(136.12,224.88) --
	(149.30,224.88) --
	(149.30,220.23) --
	(162.48,220.23) --
	(162.48,215.62) --
	(175.66,215.62) --
	(175.66,211.52) --
	(188.83,211.52) --
	(188.83,206.92) --
	(202.01,206.92) --
	(202.01,202.81) --
	(215.19,202.81) --
	(215.19,198.66) --
	(228.36,198.66) --
	(228.36,194.73) --
	(241.54,194.73) --
	(241.54,190.98) --
	(254.72,190.98) --
	(254.72,187.55) --
	(267.89,187.55) --
	(267.89,184.05) --
	(281.07,184.05) --
	(281.07,180.84) --
	(294.25,180.84) --
	(294.25,168.65);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,245.06) --
	( 70.24,245.06) --
	( 70.24,243.91) --
	( 83.42,243.91) --
	( 83.42,242.59) --
	( 96.59,242.59) --
	( 96.59,241.11) --
	(109.77,241.11) --
	(109.77,239.52) --
	(122.95,239.52) --
	(122.95,237.84) --
	(136.12,237.84) --
	(136.12,236.04) --
	(149.30,236.04) --
	(149.30,234.18) --
	(162.48,234.18) --
	(162.48,232.27) --
	(175.66,232.27) --
	(175.66,230.33) --
	(188.83,230.33) --
	(188.83,228.36) --
	(202.01,228.36) --
	(202.01,226.37) --
	(215.19,226.37) --
	(215.19,224.33) --
	(228.36,224.33) --
	(228.36,222.29) --
	(241.54,222.29) --
	(241.54,220.26) --
	(254.72,220.26) --
	(254.72,218.27) --
	(267.89,218.27) --
	(267.89,216.24) --
	(281.07,216.24) --
	(281.07,214.24) --
	(294.25,214.24) --
	(294.25,212.02);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,244.62) --
	( 70.24,244.62) --
	( 70.24,243.07) --
	( 83.42,243.07) --
	( 83.42,241.41) --
	( 96.59,241.41) --
	( 96.59,239.56) --
	(109.77,239.56) --
	(109.77,237.62) --
	(122.95,237.62) --
	(122.95,235.60) --
	(136.12,235.60) --
	(136.12,233.47) --
	(149.30,233.47) --
	(149.30,231.24) --
	(162.48,231.24) --
	(162.48,228.97) --
	(175.66,228.97) --
	(175.66,226.62) --
	(188.83,226.62) --
	(188.83,224.30) --
	(202.01,224.30) --
	(202.01,221.92) --
	(215.19,221.92) --
	(215.19,219.51) --
	(228.36,219.51) --
	(228.36,217.09) --
	(241.54,217.09) --
	(241.54,214.63) --
	(254.72,214.63) --
	(254.72,212.15) --
	(267.89,212.15) --
	(267.89,209.73) --
	(281.07,209.73) --
	(281.07,207.33) --
	(294.25,207.33) --
	(294.25,204.79);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,243.20) --
	( 70.24,243.20) --
	( 70.24,240.04) --
	( 83.42,240.04) --
	( 83.42,236.63) --
	( 96.59,236.63) --
	( 96.59,233.06) --
	(109.77,233.06) --
	(109.77,229.39) --
	(122.95,229.39) --
	(122.95,225.81) --
	(136.12,225.81) --
	(136.12,222.15) --
	(149.30,222.15) --
	(149.30,218.49) --
	(162.48,218.49) --
	(162.48,214.87) --
	(175.66,214.87) --
	(175.66,211.28) --
	(188.83,211.28) --
	(188.83,207.72) --
	(202.01,207.72) --
	(202.01,204.32) --
	(215.19,204.32) --
	(215.19,200.86) --
	(228.36,200.86) --
	(228.36,197.47) --
	(241.54,197.47) --
	(241.54,194.12) --
	(254.72,194.12) --
	(254.72,190.96) --
	(267.89,190.96) --
	(267.89,187.78) --
	(281.07,187.78) --
	(281.07,185.03) --
	(294.25,185.03) --
	(294.25,182.31);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,246.02) --
	( 57.06,246.02) --
	( 57.06,241.91) --
	( 70.24,241.91) --
	( 70.24,237.65) --
	( 83.42,237.65) --
	( 83.42,233.36) --
	( 96.59,233.36) --
	( 96.59,228.90) --
	(109.77,228.90) --
	(109.77,224.45) --
	(122.95,224.45) --
	(122.95,220.18) --
	(136.12,220.18) --
	(136.12,215.86) --
	(149.30,215.86) --
	(149.30,211.53) --
	(162.48,211.53) --
	(162.48,207.29) --
	(175.66,207.29) --
	(175.66,203.01) --
	(188.83,203.01) --
	(188.83,198.88) --
	(202.01,198.88) --
	(202.01,194.88) --
	(215.19,194.88) --
	(215.19,190.88) --
	(228.36,190.88) --
	(228.36,186.95) --
	(241.54,186.95) --
	(241.54,182.99) --
	(254.72,182.99) --
	(254.72,179.15) --
	(267.89,179.15) --
	(267.89,175.46) --
	(281.07,175.46) --
	(281.07,172.26) --
	(294.25,172.26) --
	(294.25,169.25);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37,160.14) rectangle (306.77,250.11);
\end{scope}
\begin{scope}
\path[clip] ( 31.37, 64.67) rectangle (306.77,154.64);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 31.37, 64.67) rectangle (306.77,154.64);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37, 84.71) --
	(306.77, 84.71);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,111.05) --
	(306.77,111.05);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 31.37,137.38) --
	(306.77,137.38);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 63.65, 64.67) --
	( 63.65,154.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (129.54, 64.67) --
	(129.54,154.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (195.42, 64.67) --
	(195.42,154.64);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.31, 64.67) --
	(261.31,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 71.55) --
	(306.77, 71.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37, 97.88) --
	(306.77, 97.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,124.22) --
	(306.77,124.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 31.37,150.55) --
	(306.77,150.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 64.67) --
	( 96.59,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 64.67) --
	(162.48,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 64.67) --
	(228.36,154.64);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 64.67) --
	(294.25,154.64);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,150.55) --
	( 70.24,150.55) --
	( 70.24,150.55) --
	( 83.42,150.55) --
	( 83.42,147.34) --
	( 96.59,147.34) --
	( 96.59,143.70) --
	(109.77,143.70) --
	(109.77,139.88) --
	(122.95,139.88) --
	(122.95,135.65) --
	(136.12,135.65) --
	(136.12,131.81) --
	(149.30,131.81) --
	(149.30,127.92) --
	(162.48,127.92) --
	(162.48,124.01) --
	(175.66,124.01) --
	(175.66,120.10) --
	(188.83,120.10) --
	(188.83,116.29) --
	(202.01,116.29) --
	(202.01,112.73) --
	(215.19,112.73) --
	(215.19,109.26) --
	(228.36,109.26) --
	(228.36,106.08) --
	(241.54,106.08) --
	(241.54,102.93) --
	(254.72,102.93) --
	(254.72, 99.83) --
	(267.89, 99.83) --
	(267.89, 96.92) --
	(281.07, 96.92) --
	(281.07, 94.25) --
	(294.25, 94.25) --
	(294.25, 83.61);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,150.55) --
	( 70.24,150.55) --
	( 70.24,150.55) --
	( 83.42,150.55) --
	( 83.42,146.65) --
	( 96.59,146.65) --
	( 96.59,141.13) --
	(109.77,141.13) --
	(109.77,135.61) --
	(122.95,135.61) --
	(122.95,130.52) --
	(136.12,130.52) --
	(136.12,125.32) --
	(149.30,125.32) --
	(149.30,120.35) --
	(162.48,120.35) --
	(162.48,115.63) --
	(175.66,115.63) --
	(175.66,111.00) --
	(188.83,111.00) --
	(188.83,106.51) --
	(202.01,106.51) --
	(202.01,102.52) --
	(215.19,102.52) --
	(215.19, 98.47) --
	(228.36, 98.47) --
	(228.36, 94.60) --
	(241.54, 94.60) --
	(241.54, 91.02) --
	(254.72, 91.02) --
	(254.72, 87.48) --
	(267.89, 87.48) --
	(267.89, 83.96) --
	(281.07, 83.96) --
	(281.07, 80.93) --
	(294.25, 80.93) --
	(294.25, 68.76);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,149.34) --
	( 70.24,149.34) --
	( 70.24,148.04) --
	( 83.42,148.04) --
	( 83.42,146.59) --
	( 96.59,146.59) --
	( 96.59,145.04) --
	(109.77,145.04) --
	(109.77,143.41) --
	(122.95,143.41) --
	(122.95,141.71) --
	(136.12,141.71) --
	(136.12,139.90) --
	(149.30,139.90) --
	(149.30,138.03) --
	(162.48,138.03) --
	(162.48,136.10) --
	(175.66,136.10) --
	(175.66,134.16) --
	(188.83,134.16) --
	(188.83,132.18) --
	(202.01,132.18) --
	(202.01,130.20) --
	(215.19,130.20) --
	(215.19,128.16) --
	(228.36,128.16) --
	(228.36,126.14) --
	(241.54,126.14) --
	(241.54,124.12) --
	(254.72,124.12) --
	(254.72,122.14) --
	(267.89,122.14) --
	(267.89,120.11) --
	(281.07,120.11) --
	(281.07,118.12) --
	(294.25,118.12) --
	(294.25,115.91);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,148.97) --
	( 70.24,148.97) --
	( 70.24,147.24) --
	( 83.42,147.24) --
	( 83.42,145.46) --
	( 96.59,145.46) --
	( 96.59,143.56) --
	(109.77,143.56) --
	(109.77,141.56) --
	(122.95,141.56) --
	(122.95,139.50) --
	(136.12,139.50) --
	(136.12,137.33) --
	(149.30,137.33) --
	(149.30,135.09) --
	(162.48,135.09) --
	(162.48,132.82) --
	(175.66,132.82) --
	(175.66,130.50) --
	(188.83,130.50) --
	(188.83,128.20) --
	(202.01,128.20) --
	(202.01,125.90) --
	(215.19,125.90) --
	(215.19,123.50) --
	(228.36,123.50) --
	(228.36,121.13) --
	(241.54,121.13) --
	(241.54,118.80) --
	(254.72,118.80) --
	(254.72,116.52) --
	(267.89,116.52) --
	(267.89,114.21) --
	(281.07,114.21) --
	(281.07,111.97) --
	(294.25,111.97) --
	(294.25,109.50);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,147.00) --
	( 70.24,147.00) --
	( 70.24,143.42) --
	( 83.42,143.42) --
	( 83.42,139.69) --
	( 96.59,139.69) --
	( 96.59,135.94) --
	(109.77,135.94) --
	(109.77,132.19) --
	(122.95,132.19) --
	(122.95,128.56) --
	(136.12,128.56) --
	(136.12,124.88) --
	(149.30,124.88) --
	(149.30,121.23) --
	(162.48,121.23) --
	(162.48,117.58) --
	(175.66,117.58) --
	(175.66,113.99) --
	(188.83,113.99) --
	(188.83,110.43) --
	(202.01,110.43) --
	(202.01,107.04) --
	(215.19,107.04) --
	(215.19,103.62) --
	(228.36,103.62) --
	(228.36,100.27) --
	(241.54,100.27) --
	(241.54, 96.95) --
	(254.72, 96.95) --
	(254.72, 93.82) --
	(267.89, 93.82) --
	(267.89, 90.65) --
	(281.07, 90.65) --
	(281.07, 87.93) --
	(294.25, 87.93) --
	(294.25, 85.24);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 43.89,150.55) --
	( 57.06,150.55) --
	( 57.06,145.92) --
	( 70.24,145.92) --
	( 70.24,141.15) --
	( 83.42,141.15) --
	( 83.42,136.56) --
	( 96.59,136.56) --
	( 96.59,131.97) --
	(109.77,131.97) --
	(109.77,127.38) --
	(122.95,127.38) --
	(122.95,123.00) --
	(136.12,123.00) --
	(136.12,118.62) --
	(149.30,118.62) --
	(149.30,114.27) --
	(162.48,114.27) --
	(162.48,110.02) --
	(175.66,110.02) --
	(175.66,105.79) --
	(188.83,105.79) --
	(188.83,101.71) --
	(202.01,101.71) --
	(202.01, 97.83) --
	(215.19, 97.83) --
	(215.19, 93.87) --
	(228.36, 93.87) --
	(228.36, 90.03) --
	(241.54, 90.03) --
	(241.54, 86.28) --
	(254.72, 86.28) --
	(254.72, 82.77) --
	(267.89, 82.77) --
	(267.89, 79.27) --
	(281.07, 79.27) --
	(281.07, 76.29) --
	(294.25, 76.29) --
	(294.25, 73.36);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 31.37, 64.67) rectangle (306.77,154.64);
\end{scope}
\begin{scope}
\path[clip] (306.77,446.55) rectangle (319.71,536.52);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,446.55) rectangle (319.71,536.52);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,491.54) {Scenario 1};
\end{scope}
\begin{scope}
\path[clip] (306.77,351.08) rectangle (319.71,441.05);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,351.08) rectangle (319.71,441.05);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,396.07) {Scenario 2};
\end{scope}
\begin{scope}
\path[clip] (306.77,255.61) rectangle (319.71,345.58);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,255.61) rectangle (319.71,345.58);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,300.60) {Scenario 3};
\end{scope}
\begin{scope}
\path[clip] (306.77,160.14) rectangle (319.71,250.11);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77,160.14) rectangle (319.71,250.11);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,205.12) {Scenario 4};
\end{scope}
\begin{scope}
\path[clip] (306.77, 64.67) rectangle (319.71,154.64);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77, 64.67) rectangle (319.71,154.64);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,109.65) {Scenario 5};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 96.59, 61.92) --
	( 96.59, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (162.48, 61.92) --
	(162.48, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (228.36, 61.92) --
	(228.36, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.25, 61.92) --
	(294.25, 64.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 96.59, 54.21) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (162.48, 54.21) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (228.36, 54.21) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (294.25, 54.21) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,450.68) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,477.01) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,503.35) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,529.68) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,453.43) --
	( 31.37,453.43);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,479.77) --
	( 31.37,479.77);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,506.10) --
	( 31.37,506.10);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,532.44) --
	( 31.37,532.44);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,355.21) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,381.54) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,407.88) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,434.21) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,357.96) --
	( 31.37,357.96);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,384.30) --
	( 31.37,384.30);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,410.63) --
	( 31.37,410.63);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,436.96) --
	( 31.37,436.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,259.74) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,286.07) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,312.40) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,338.74) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,262.49) --
	( 31.37,262.49);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,288.82) --
	( 31.37,288.82);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,315.16) --
	( 31.37,315.16);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,341.49) --
	( 31.37,341.49);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,164.26) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,190.60) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,216.93) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,243.27) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,167.02) --
	( 31.37,167.02);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,193.35) --
	( 31.37,193.35);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,219.69) --
	( 31.37,219.69);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,246.02) --
	( 31.37,246.02);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 68.79) {0.7};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42, 95.13) {0.8};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,121.46) {0.9};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 26.42,147.79) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 71.55) --
	( 31.37, 71.55);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62, 97.88) --
	( 31.37, 97.88);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,124.22) --
	( 31.37,124.22);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 28.62,150.55) --
	( 31.37,150.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (169.07, 43.70) {Years since hire};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,300.60) {Cancer-free survival};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 19.60,  5.50) rectangle (142.67, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 30.60, 11.00) rectangle ( 45.05, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.05, 18.23) -- ( 43.61, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 86.63, 11.00) rectangle (101.09, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 88.08, 18.23) -- ( 99.64, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 50.55, 15.13) {Rule $a_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (106.59, 15.13) {Rule $a_1$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (153.67,  5.50) rectangle (318.53, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (164.67, 11.00) rectangle (179.12, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (166.11, 18.23) -- (177.68, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (212.89, 11.00) rectangle (227.34, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (214.34, 18.23) -- (225.90, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (262.84, 11.00) rectangle (277.29, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (264.28, 18.23) -- (275.85, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (184.62, 15.13) {Truth};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (232.84, 15.13) {WKM};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,542.02);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (282.79, 15.13) {AWKM};
\end{scope}
\end{tikzpicture}

\end{document}