### 2022-03-04 reports/resources/sensitivity/results_1.tex

% Created by tikzDevice version 0.12.3.1 on 2022-01-19 23:23:08
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{amssymb}
\usepackage[no-math]{fontspec}

\setmainfont{Arial}

\usepackage[italic]{mathastext}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (325.21,154.86);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (325.21,154.86);
\end{scope}
\begin{scope}
\path[clip] ( 35.37, 64.67) rectangle (306.77,149.36);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 35.37, 64.67) rectangle (306.77,149.36);
\definecolor{drawColor}{gray}{0.92}

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37, 79.54) --
	(306.77, 79.54);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,105.93) --
	(306.77,105.93);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 35.37,132.32) --
	(306.77,132.32);

\path[draw=drawColor,line width= 0.3pt,line join=round] ( 67.18, 64.67) --
	( 67.18,149.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (132.11, 64.67) --
	(132.11,149.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (197.04, 64.67) --
	(197.04,149.36);

\path[draw=drawColor,line width= 0.3pt,line join=round] (261.97, 64.67) --
	(261.97,149.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 66.34) --
	(306.77, 66.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37, 92.73) --
	(306.77, 92.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,119.12) --
	(306.77,119.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.37,145.51) --
	(306.77,145.51);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.65, 64.67) --
	( 99.65,149.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (164.57, 64.67) --
	(164.57,149.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.50, 64.67) --
	(229.50,149.36);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.43, 64.67) --
	(294.43,149.36);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 47.70,145.51) --
	( 60.69,145.51) --
	( 73.67,145.51) --
	( 86.66,143.30) --
	( 99.65,140.75) --
	(112.63,138.01) --
	(125.62,134.78) --
	(138.60,131.40) --
	(151.59,128.22) --
	(164.57,124.71) --
	(177.56,121.64) --
	(190.55,118.40) --
	(203.53,115.01) --
	(216.52,111.80) --
	(229.50,109.04) --
	(242.49,106.38) --
	(255.47,103.32) --
	(268.46, 99.77) --
	(281.44, 96.75) --
	(294.43, 85.15);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 47.70,145.51) --
	( 60.69,145.51) --
	( 73.67,145.51) --
	( 86.66,142.38) --
	( 99.65,138.02) --
	(112.63,133.41) --
	(125.62,128.63) --
	(138.60,124.26) --
	(151.59,119.71) --
	(164.57,115.31) --
	(177.56,111.16) --
	(190.55,106.98) --
	(203.53,103.08) --
	(216.52, 99.10) --
	(229.50, 95.50) --
	(242.49, 91.83) --
	(255.47, 88.16) --
	(268.46, 84.67) --
	(281.44, 81.50) --
	(294.43, 68.52);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 47.70,145.51) --
	( 60.69,144.58) --
	( 73.67,143.55) --
	( 86.66,142.41) --
	( 99.65,141.16) --
	(112.63,139.78) --
	(125.62,138.35) --
	(138.60,136.78) --
	(151.59,135.15) --
	(164.57,133.48) --
	(177.56,131.78) --
	(190.55,130.01) --
	(203.53,128.23) --
	(216.52,126.40) --
	(229.50,124.54) --
	(242.49,122.68) --
	(255.47,120.85) --
	(268.46,118.98) --
	(281.44,117.13) --
	(294.43,115.03);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 47.70,145.51) --
	( 60.69,144.30) --
	( 73.67,142.98) --
	( 86.66,141.51) --
	( 99.65,139.92) --
	(112.63,138.20) --
	(125.62,136.41) --
	(138.60,134.48) --
	(151.59,132.46) --
	(164.57,130.39) --
	(177.56,128.31) --
	(190.55,126.17) --
	(203.53,123.97) --
	(216.52,121.74) --
	(229.50,119.50) --
	(242.49,117.29) --
	(255.47,115.06) --
	(268.46,112.77) --
	(281.44,110.57) --
	(294.43,108.03);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 47.70,145.51) --
	( 60.69,142.77) --
	( 73.67,139.94) --
	( 86.66,136.99) --
	( 99.65,133.96) --
	(112.63,130.76) --
	(125.62,127.69) --
	(138.60,124.46) --
	(151.59,121.22) --
	(164.57,118.02) --
	(177.56,114.80) --
	(190.55,111.53) --
	(203.53,108.40) --
	(216.52,105.20) --
	(229.50,102.00) --
	(242.49, 98.82) --
	(255.47, 95.78) --
	(268.46, 92.70) --
	(281.44, 90.00) --
	(294.43, 87.25);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 47.70,145.51) --
	( 60.69,141.95) --
	( 73.67,138.32) --
	( 86.66,134.53) --
	( 99.65,130.68) --
	(112.63,126.70) --
	(125.62,122.87) --
	(138.60,118.92) --
	(151.59,114.92) --
	(164.57,110.94) --
	(177.56,107.05) --
	(190.55,103.11) --
	(203.53, 99.27) --
	(216.52, 95.40) --
	(229.50, 91.58) --
	(242.49, 87.82) --
	(255.47, 84.15) --
	(268.46, 80.43) --
	(281.44, 77.26) --
	(294.43, 73.99);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 35.37, 64.67) rectangle (306.77,149.36);
\end{scope}
\begin{scope}
\path[clip] (306.77, 64.67) rectangle (319.71,149.36);
\definecolor{drawColor}{gray}{0.20}
\definecolor{fillColor}{gray}{0.85}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round,fill=fillColor] (306.77, 64.67) rectangle (319.71,149.36);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,rotate=-90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (310.14,107.02) {Scenario 1};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 99.65, 61.92) --
	( 99.65, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (164.57, 61.92) --
	(164.57, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (229.50, 61.92) --
	(229.50, 64.67);

\path[draw=drawColor,line width= 0.6pt,line join=round] (294.43, 61.92) --
	(294.43, 64.67);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 99.65, 54.21) {5};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (164.57, 54.21) {10};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (229.50, 54.21) {15};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.80] at (294.43, 54.21) {20};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 63.59) {0.85};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42, 89.98) {0.90};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,116.37) {0.95};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.80] at ( 30.42,142.76) {1.00};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 66.34) --
	( 35.37, 66.34);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62, 92.73) --
	( 35.37, 92.73);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,119.12) --
	( 35.37,119.12);

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 32.62,145.51) --
	( 35.37,145.51);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (171.07, 43.70) {Years since hire};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 11.70,107.02) {Cancer-free survival};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 21.60,  5.50) rectangle (144.67, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 32.60, 11.00) rectangle ( 47.05, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{248,118,109}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 34.05, 18.23) -- ( 45.61, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 88.63, 11.00) rectangle (103.09, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,191,196}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 90.08, 18.23) -- (101.64, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 52.55, 15.13) {Rule $a_0$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (108.59, 15.13) {Rule $a_1$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (155.67,  5.50) rectangle (320.53, 30.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (166.67, 11.00) rectangle (181.12, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,line join=round] (168.11, 18.23) -- (179.68, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (214.89, 11.00) rectangle (229.34, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (216.34, 18.23) -- (227.90, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (264.84, 11.00) rectangle (279.29, 25.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (266.28, 18.23) -- (277.85, 18.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (186.62, 15.13) {Truth};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (234.84, 15.13) {WKM};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (325.21,154.86);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.90] at (284.79, 15.13) {AWKM};
\end{scope}
\end{tikzpicture}

\end{document}