### 2022-03-04 slides.Rmd

---
title: "Estimating survival in left filtered data"
subtitle: "Replication and application of Izano's estimator"
author: "Kevin Chen"
fontsize: 11pt
# mainfont: arial
output:
  beamer_presentation:
    toc: false
    slide_level: 1
    # latex_engine: lualatex
    includes:
      in_header: "~/HeadRs/BeamerHead.sty"
bibliography: "`r path.expand('~/Box Sync/papers/EnvEpi.bib')`"
---

```{r, include=F}
knitr::opts_chunk$set(echo = F, results = "asis", fig.align = "center")
if (!"mytheme" %in% ls()) {
  source("~/headRs/00-my-theme.R")
  library(data.table)
  options(tikzDefaultEngine = 'pdftex')
}
library(tidyverse); library(data.table)
library(here)
```

# Chapter 3 of Izano (2017)

## Drivers of Biased Effect Estimates in Left Filtered Data

- Specified a SEM for "left filtering" in the presence of HWSE
- Simulated large-sample ($n = 10^6$) data once for each of 5 scenarios
- Simulated finite-sample ($n = 50,000$) data 500 times for each scenario
- Applied two estimators:
  - Adjusted Kaplan-Meier
  - Delayed-entry adjusted Kaplan-Meier

# Description of variables

\begin{table}
\begin{center}
\begin{adjustbox}{width=0.85\linewidth}{
\begin{tabular}{lp{0.65\linewidth}}
\toprule
 Variable  &  Description                                 \\ \midrule
 $R$       & Time until start of registry \\
 $W$       & Baseline covariates \\
 $S$       & Susceptibility to effects of metalworking fluid exposure \\
 $H(t)$    & Adverse health status at time $t$ \\
 $N(t)$    & Employment status at time $t$ \\
 $A(t)$    & Metalworking fluid exposure at time $t$ \\
 $D(t)$    & Mortality status at time $t$ \\
 $Y^*(t)$  & Cancer status at time $t$ \\
 $Y(t)$    & Observed Cancer status at time $t$ \\
 $t = \{1, 2, \ldots, 20\}$ & Time, indexed in years after hire \\
 \bottomrule
\end{tabular}
}\end{adjustbox}
\end{center}
\end{table}

# Abbreviated DAG summarizing the causal structure

\begin{figure}
\begin{center}
\begin{adjustbox}{width=\linewidth}{
\begin{tikzpicture}[>= stealth, auto, node distance = 2.25cm, semithick, fill=white, inner sep=0pt]
\tikzstyle{every state}=[shape = circle, align = center, draw = none]
\node[state] (H0) {$H(t - 1)$};
\node[state] (N0) [below right of=H0] {$N(t - 1)$};
\node[state] (A0) [below right of=N0] {$\bar A(t - 1)$};
\node[state] (D0) [below right of=A0] {$D(t - 1)$};
\node[state] (Y0) [below right of=D0] {$Y^*(t - 1)$};
\node[state] (W) [above right of=H0] {$W$};
\node[state] (H) [right of=H0, node distance = 4cm] {$H(t)$};
\node[state] (N) [below right of=H, node distance = 4cm] {$N(t)$};
\node[state] (A) [below right of=N, node distance = 4cm] {$\bar A(t)$};
\node[state] (D) [above right of=N, node distance = 3cm] {$D(t)$};
\node[state] (Y) [above right of=A, node distance = 3cm] {$Y^*(t)$};
\node[state] (S) [above right of=D, node distance = 3cm] {$S$};
%
\path[->] (H0) edge (N0);
\path[->] (N0) edge (A0);
\path[->] (A0) edge [bend right=30] (Y0);
\path[->] (H0) edge (H);
\path[->] (W) edge [bend right=10] (N);
\path[->] (N0) edge (N);
\path[->] (H) edge (N);
\path[->] (A0) edge (N);
\path[->] (W) edge [bend right=15] (A);
\path[->] (A0) edge (A);
\path[->] (N) edge (A);
\path[->] (W) edge (D);
\path[->] (A0) edge [bend left=15] (D);
\path[->] (D0) edge [bend right=15] (D);
\path[->] (Y0) edge [bend right=10] (D);
\path[->] (W) edge [bend left=5] (Y);
\path[->] (S) edge [bend left=30] (Y);
\path[->] (S) edge [bend left=15] (Y0);
\path[->] (H) edge (Y);
\path[->] (A) edge (Y);
\path[->] (D) edge (Y);
\path[->] (Y0) edge (Y);
\end{tikzpicture}
}\end{adjustbox}
\end{center}
\end{figure}

# Scenarios

- Five scenarios represent 5 sets of parameters used to generate data according to the SEM
- Sets 2, 3, 4, and 5 differ from Set 1 by a single parameter each
  - Scenario 2: Greater cancer-related mortality\
  (increase HR from 1.6 to 7.4)
  - Scenario 3: Greater proportion of susceptibles\
  (increase proportion from 10\% to 20\%)
  - Scenario 4: Greater time-varying confounding by history of adverse health\
  (decrease HR of being at work from 0.6 to 0.2 \
  and increase HR of cancer incidence from 2.0 to 5.5)
  - Scenario 5: Greater background incidence\
  (increase baseline log hazard from -7 to -6)

# Target and estimation

- Goal: estimate cancer-free survival indexed by time since hire
  - Under rule where all are exposed while at work
  - Under rule where all are not exposed (ever)
  - Competing risk of death is accounted for by preventing death
- Adjusted KM ie the inverse probability of treatment weighted KM (WKM)
- Delayed-entry adjusted KM ie the Aalen-filtered WKM (AWKM)

# Review of the Kaplan-Meier Estimator

- Let $c(t)$ be the number of cases in the interval $(t - 1, t]$
- Let $r(t)$ be the number of people at risk in interval $(t - 1, t]$
- The standard survival estimator is
$$\widehat {S}(t) =
\begin{cases}
1 & \text{if } t < t_1 \\
\prod_{j \le t} \left(1 - \frac{c(j)}{r(j)}\right) & \text{if } t \ge t_1 \\
\end{cases}$$
where $t_1$ is the first event time
- Restricting to followers of rule $a$, we have \vspace{-0.5\baselineskip}
$$c_a(t), \hspace{1em} r_a(t), \hspace{1em} \widehat{S}_a(t)$$

# Mathematical expression of $c_a(t)$ and $r_a(t)$

$$\begin{aligned}
c_{a} (t)
& = \sum_i^n \underbrace{\Ind{Y_i(t) = 1}}_{\text{Cancer by time } t}
    \times \underbrace{\Ind{Y_i(t - 1) = 0}}_{\text{At-risk at time } t - 1}
    \times \underbrace{\Ind{\bar A_i (t) = \bar a(t)}}_{\text{Followed rule } a} \\
r_{a} (t)
& = \sum_i^n \underbrace{\Ind{Y_i(t - 1) = 0}}_{\text{At-risk at time } t - 1}
    \times \underbrace{\Ind{\bar A_i (t) = \bar a(t)}}_{\text{Followed rule } a}.
\end{aligned}$$

# Inverse probability of treatment weighted KM (WKM)

The $i$th person at time $t$ is weighted by $w_{i, a}(t)$ the predicted probability of following rule $a$ through time $t$:

\begin{adjustbox}{scale=0.8}\parbox{\linewidth}{
$$\begin{aligned}
c^w_{a} (t)
& = \sum_i^n w_{i, a}(t)
    \times \Ind{Y_i(t) = 1}
    \times \Ind{Y_i(t - 1) = 0}
    \times \Ind{\bar A_i (t) = \bar a(t)} \\
r^w_{a} (t)
& = \sum_i^n  w_{i, a}(t)
    \times \Ind{Y_i(t - 1) = 0}
    \times \Ind{\bar A_i (t) = \bar a(t)}.
\end{aligned}$$}
\end{adjustbox}

The $w_{i, a}$ can be estimated by fitting logistic regressions.

# Aalen-filtered WKM (AWKM)

In addition to weighting, we restrict the computation of the discrete hazard to those under observation:

\begin{adjustbox}{scale=0.8}\parbox{\linewidth}{
$$\begin{aligned}
c_{a} (t)
& = \sum_i^n w_{i, a}(t) \times \Ind{Y_i(t) = 1} \times \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)} \times \Ind{t \ge R_i} \\
r_{a} (t)
& = \sum_i^n w_{i, a}(t) \times \Ind{Y_i(t - 1) = 0} \times \Ind{\bar A_i (t) = \bar a(t)} \times \Ind{t \ge R_i}
\end{aligned}$$
}\end{adjustbox}

# Results from Izano's Dissertation (2017)

\includegraphics[width=\textwidth]{private/izano_2017_result_1}
\includegraphics[width=\textwidth]{private/izano_2017_result_2}

\begin{center}
\begin{tabular}{ll}
Red: Always exposed & Blue: Never exposed
\end{tabular}
\begin{tabular}{lll}
Long dashed: WKM & Dot dashed: AWKM & Solid: True KM curve
\end{tabular}
\end{center}

# Results from Izano's Dissertation (2017)

\includegraphics[width=\textwidth]{private/izano_2017_result_3}
\includegraphics[width=\textwidth]{private/izano_2017_result_4}

\begin{center}
\begin{tabular}{ll}
Red: Always exposed & Blue: Never exposed
\end{tabular}
\begin{tabular}{lll}
Long dashed: WKM & Dot dashed: AWKM & Solid: True KM curve
\end{tabular}
\end{center}

# Results from Izano's Dissertation (2017)

\includegraphics[width=\textwidth]{private/izano_2017_result_5}

\begin{center}
\begin{tabular}{ll}
Red: Always exposed & Blue: Never exposed
\end{tabular}
\begin{tabular}{lll}
Long dashed: WKM & Dot dashed: AWKM & Solid: True KM curve
\end{tabular}
\end{center}

# Replication results

\includegraphics[width=\textwidth]{reports/resources/results_1-2}

# Replication results

\includegraphics[width=\textwidth]{reports/resources/results_3-4}

# Replication results

\includegraphics[width=\textwidth]{reports/resources/results_5}

# Application to UAW-GM

- Study population: Workers from Plants 1 or 2, hired after 1938
- 55-year follow-up starting at hire
- MWF exposure lagged 15 years
- Employment records end in 1994; workers considered administratively censored if they were still at work in 2010 (15 years after1 995)
- Rules of interest
  - Natural course: ever-exposed above reference level (0 mg/m^3^$\cdot$years for straights and synthetics; 0.05 mg/m^3^$\cdot$years for solubles)
  - Never exposed
- Under both rules, no censoring by death

# Application results

\begin{center}
\includegraphics[width=0.95\textwidth]{reports/resources/gm_results}
\end{center}